---
title: "Tutorial"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


## Topic 1

### Exercise 1

Write the R code required to add 2 plus 2:

```{r two-plus-two, exercise=TRUE, exercise.lines = 1}

```

```{r two-plus-two-solution}
2+2
```

### Exercise 2 

Write the R code required to subtract 3 from 7:

```{r seven-minus-three, exercise=TRUE, exercise.lines = 1}

```

```{r seven-minus-three-solution}
7-3
```


### Exercise 3

Write the R code requried to divide 10 by 2:

```{r ten-divide-two, exercise=TRUE, exercise.lines = 1}

```

```{r ten-divide-two-solution}
10/2
```

### Exercise 4 

Write the R code required to multiple 5 by 5:
```{r five-multiply-five, exercise=TRUE, exercise.lines = 1}

```

```{r five-multiply-five-solution}
5*5
```

## Topic 2 


### Exercise 1

Write a code chunk that shows the first 6 outputs of the data set Iris:

```{r code-chunk, exercise=TRUE, exercise.lines = 1}

```

```{r code-chunk-solution}
head(iris)
```
### Exercise 2

Write a code chunk that shows the last 6 outputs of the data set Iris:

```{r code-chunk2, exercise=TRUE, exercise.lines = 1}

```

```{r code-chunk2-solution}
tail(iris)
```


## Topic 3


