---
title: "Learning LearnR"
author: "Jake Kovaleski"
output: 
  learnr::tutorial:
    theme: united
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
```


## Using ggplot2

### Exercise 1

*Let's preview the dataset you'll be using*

``` {r iris head}
head(iris)
```

Write the R code required make a histogram:

*Remember: the varible needs to be continous to be plotted*

```{r simple_hist, exercise=TRUE}

```

### Exercise 2

*Let's clean up those axes*

Repeat your code, but this time use the `labs()` function to make the axis titles look better:

```{r clean_axis_titles, exercise=TRUE}
  

```

### Exercise 3

*Now let's add some color. Keep in mind the difference between color and fill, and where to insert them in the code*

``` {r adding_color, exercise=TRUE}


```

## Using dplyr

### Exercise 1

Remind youself of the dataset we are using (iris) by only showing the first five rows:

```{r print-limit, exercise=TRUE}

```

```{r print-limit-hint}
'head()' is the function to use
```

### Exercise 2

Using a dyplr function, wrangle the data to only display the versicolor species.

```{r filter-prac,exercise=TRUE }


```

```{r filter-prac-hint}
while select() *selects* columns, what function extracts rows?

```

### Exercise 3

Combine your skills to make a histogram displaying petal length only across the versicolor species.

```{r dpylr_ggplot, exercise=TRUE }


```









