---
title: "Logical Operators"
author: "Jake Kovaleski"
output: 
  learnr::tutorial:
    theme: united
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
```


### Exercise 1

Write the R code required to prove 7 is greater than 5:

*The correct output will be `TRUE`.*

```{r 7_g_5, exercise=TRUE}

```
```{r 7_g_5-solution}
7 > 5 # OR
5 < 7
```

### Exercise 2

Write the R code required to prove 256 is equal to 256

```{r equal, exercise=TRUE}

```
```{r equal-solution}
256 == 256
```

### Exercise 3

Write the R code required to prove 11 is not equal to 23.

```{r not, exercise=TRUE}


```
```{r not-solution}
11 != 23
```

### Exercise 4

Write the R code required to prove 28 is greater than or equal to 13

```{r gret_or_eq, exercise=TRUE}

```
```{r gret_or_eq-solution}
28 >= 13 # OR
13 <= 28
```


### Exercise 5

Assign the value of 33 to `x`, then write the R code necessary to prove `x` is greater than 12 AND `x` is less than 146

```{r and, exercise=TRUE}

```
```{r and-solution}
x = 33
x > 12 & x < 146
```


### Exercise 6

Make `x = 47` and `y = 93`. Use only the varibles, `&`, `<`, `>`, and `!=` to get a `TRUE` output.


```{r many_types, exercise= TRUE}


```
```{r many_types-solution}
x = 47
y = 93
x < y & y > x & x != y

```
### Exercise 7

Assign an array from 1 to 10 to the varible `x`, then visiualize the whole array. 

``` {r array, exercise = TRUE}

```
```{r array-solution}
x = 1:10
print(x)
```

### Exercise 8

Using `x = 1:10` again, prove the number 6 is within the array.

```{r within_array, exercise= TRUE}

```
```{r within_array-solution}
x = 1:10
6 %in% x
```

### Exercise 9

Using `x = 1:10` once more, prove that either 2 OR 6278 is within that array.

```{r or, exercise=TRUE}

```
```{r or-solution}
x = 1:10
1 %in% x | 6278 %in% x
```


### Exercise 10

Say you wanted to know if the `starwars` dataset included any data on Lando Calrissian, how could you quickly see if he was included?

```{r Lando, exercise = TRUE}

```
```{r Lando-solution}
"Lando Calrissian" %in% starwars$name
```

### Exercise 11

`%in` can also be used inside the `filter()` function. Use it to extract the characters with orange skin color.

```{r orange, exercise = TRUE}

```
```{r orange-solution}
starwars %>% filter(skin_color %in% "orange")
```



