---
title: "Logical Operators"
author: "Jake Kovaleski"
output: 
  learnr::tutorial:
    theme: united
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(dslabs)
knitr::opts_chunk$set(echo = FALSE)
```

## Simple Operations

### Exercise 1

Write R code to ask whether 7 is greater than 5.

*The correct output will be `TRUE`.*

```{r simple_logical_1, exercise=TRUE}

```
```{r simple_logical_1-solution}
7 > 5 # OR
5 < 7
```

### Exercise 2

Write R code to ask whether 256 is equal to 256.

```{r simple_logical_2, exercise=TRUE}

```
```{r simple_logical_2-solution}
256 == 256
```

### Exercise 3

Write R code to ask whether 11 is not equal to 23.

```{r simple_logical_3, exercise=TRUE}


```
```{r simple_logical_3-solution}
11 != 23
```

### Exercise 4

Write R code to ask whether 28 is greater than or equal to 13.

*`>=` is the operator that asks this *

```{r simple_logical_4, exercise=TRUE}

```
```{r simple_logical_4-solution}
28 >= 13 # OR
13 <= 28
```

### Exercise 5

Create variables `k` and `f` as 43 and 92, respectively. Ask whether `k` is greater than `f`.

```{r simple_logical_5, exercise=TRUE}


```
```{r simple_logical_5-solution}
f <- 92
k <- 43
k > f

```

### Exercise 6

Use `var1 <- "13 characters"` and `var2 <- "characters = 15"` and ask whether `var1` has more characters than `var2` in its string.

*`nchar()` is a function that asks how many characters are in a string. An example has been included in the answer chunk, run the code to see how it operates.*

```{r simple_logical_6, exercise=TRUE}
example <- "blazing"
nchar(example)
```
```{r simple_logical_6-solution}
var1 <- "13 characters"
var2 <- "characters = 15"
nchar(var1) > nchar(var2)
```

## Compound Operations 

### Exercise 1

Write R code to ask whether 13 is less than 22 *OR* greater than 0.5.

```{r compound_logical_1, exercise=TRUE}

```
```{r compound_logical_1-solution}
13 < 22 | 13 > 0.5
```

### Exercise 2

Create the variable `x` as 33. Ask whether `x` is greater than 12 *AND* less than 146.

```{r compound_logical_2, exercise=TRUE}

```
```{r compound_logical_2-solution}
x <- 33
x > 12 & x < 146
```

### Exercise 3

Create variables `x` and `y` as 77 and 31, respectively. Ask whether `x` is greater than `y` OR `y` is greater than `x`.

```{r compound_logical_3, exercise=TRUE}

```
```{r compound_logical_3-solution}
x <- 77
y <- 31
x > y | y > x
```

### Exercise 4

Assign 47 to `z` and 93 to `w`. Ask whether `z` is less than `w` *AND* if `w` is greater than or equal to `z` *AND* if `z` is not equal to `w`.


```{r compound_logical_4, exercise= TRUE}


```
```{r compound_logical_4-solution}
z <- 47
w <- 93
z < w & w >= z & z != w

```

### Exercise 5

Create variables `a`, `b`, and `c` as 1, 5, and 10, respectively. 

Ask whether the sum of `a` and `b` is less than `c` *OR* the difference of `b` and `c` is greater than `a`.

```{r compound_logical_5, exercise= TRUE}


```
```{r compound_logical_5-solution}
a <- 1
b <- 5
c <- 10
(a + b) < c | (b - c) > a

```

### Exercise 6

Assign "data science" to `string1`, and "for biologists" to `string2`. Ask whether the number of characters in `string1` is equal to 12 *AND* if `string1` is less than or equal to `string2`.

*Reminder: `nchar()` is the function that produces the number of characters*
```{r compound_logical_6, exercise= TRUE}


```
```{r compound_logical_6-solution}
string1 <- "data science"
string2 <- "for biologists"
nchar(string1) == 12 & string1 <= string2

```

## Working with functions and logical

### Exercise 1

Create an array from 1 to 10 to the variable `x`, then print the whole array.

``` {r function_logical_1, exercise = TRUE}

```
```{r function_logical_1-solution}
x <- 1:10
print(x)
```

### Exercise 2

Again, define an array with values 1:10. Write R code to ask whether the number 6 is within the array.

*You must assign the array to `x` again, each chunk is its own environment.*

```{r function_logical_2, exercise= TRUE}

```
```{r function_logical_2-solution}
x <- 1:10
6 %in% x
```

### Exercise 3

Make an array, `ex3_array`, ranging from 1:500, and ask whether that either 2 OR 6278 is within that array.

```{r function_logical_3, exercise=TRUE}

```
```{r function_logical_3-solution}
ex3_array <- 1:500
1 %in% ex3_array | 6278 %in% ex3_array
```

### Exercise 4

Make an array, `ex4_array`, ranging from 47 to 423, then find the length of the array.

*`length()` is the function whose output is the length of an string, array, column, etc.*
*The example pre-populated in the answer chunk will tell you the length of the `Sepal.Length` column within the `iris` data set, run it!*

```{r function_logical_4, exercise=TRUE}
length(iris$Sepal.Length)
```
```{r function_logical_4-solution}
ex4_array <- 47:423
length(ex4_array)
```

### Exercise 5

Make an array, `ex5_array`, ranging from 73 to 632, and ask whether the length is equal to 560.

```{r function_logical_5, exercise=TRUE}

```
```{r function_logical_5-solution}
ex5_array <- 73:632
length(ex5_array) == 559
```

### Exercise 6

Make an array, `ex6_array`, ranging from 31 to 489, and another, `another_array`, ranging from 123 to 589. Ask whether the length of `ex6_array` is less than or equal to `another_array`.

```{r function_logical_6, exercise=TRUE}

```
```{r function_logical_6-solution}
ex6_array <- 31:489
another_array <- 123:589
length(ex6_array) <= length(another_array)
```


## Working with data frames and logical

### Exercise 1

Within the tidyverse, there is a data set titled `starwars` after the legendary movie franchise. 

Run the pre-populated code to see the first five rows as well as each column name to get a sense of what you'll be working with.

```{r dataframe_logical_1, exercise = TRUE}
head(starwars) 
colnames(starwars)
```

### Exercise 2

Confirm that there are 87 different Star Wars characters in the `starwars` data set by asking if its `length()` is equal to 87

*The column title is `name`*

```{r dataframe_logical_2, exercise = TRUE}

```
```{r dataframe_logical_2-solution}
length(starwars$name) == 87
```

### Exercise 3

Say you wanted to know if the `starwars` data set included any data on Lando Calrissian. Without manually scanning the entire data set, how could you quickly see if he was included?

```{r dataframe_logical_3, exercise = TRUE}

```
```{r dataframe_logical_3-solution}
"Lando Calrissian" %in% starwars$name
```

### Exercise 4
Remove any rows that have an `NA` so you'll only have rows with data for each and every column.
This function is important in the event you are working with data where any `NA` may invalidate that row.

*Namely, you'd want to use the `na.omit()` function*

To confirm that some rows have been removed, save your output from above to a variable, named `no_NAs` and write R code to ask if its length is less than 87.

```{r dataframe_logical_4, exercise = TRUE}

```
```{r dataframe_logical_4-solution}
na.omit(starwars) -> "no_NAs"
length(no_NAs) < 87
```

### Exercise 5

Determine what types of columns `hair_color` and `mass` are in the `starwars` data set.
Once determined, ask if they are the same, or equal.

Columns can be *classified* by character, numeric (real numbers), integer, complex, logical (True/False), and are found by using `class()`.

```{r dataframe_logical_5, exercise = TRUE}

```
```{r dataframe_logical_5-solution}
class(starwars$mass)
class(starwars$hair_color)
class(starwars$mass) == class(starwars$hair_color)
```

### Exercise 6

Built into RStudio, there is a data set titled `mtcars`. The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).

Run the pre-populated code to, this time, see the last five rows as well as each column name to see what those 10 aspects of automobile design and performance are.

```{r dataframe_logical_6, exercise = TRUE}
tail(mtcars) 
colnames(mtcars)
```


### Exercise 7

Someone tells you the average quarter mile time (`qsec`) of the cars within the `mtcars` data set is 17.84875 seconds. Ask whether the average is indeed equal to 17.84875 seconds.

*Means are simply found using `mean()`.*

```{r dataframe_logical_7, exercise = TRUE}

```
```{r dataframe_logical_7-solution}
mean(mtcars$qsec) == 17.84875
```

### Exercise 8

Write R code to ask whether the median MPG of the cars within the `mtcars` data set is less than or equal to 20 MPG.

*The function you need to use is similar to how `mean()` is utilized*

```{r dataframe_logical_8, exercise = TRUE}

```
```{r dataframe_logical_8-solution}
median(mtcars$mpg) <= 20
```

### Exercise 9

Write R code to ask what the maximum and minimum values are for horsepower, `hp`, within the `mtcars` data set are. Confirm these values make sense by writing R code to ask if the max is greater than or equal to the minimum.

*You'll need to use `min()` and `max()` to find these values*

```{r dataframe_logical_9, exercise = TRUE}

```
```{r dataframe_logical_9-solution}
max(mtcars$hp)
min(mtcars$hp)
max(mtcars$hp) >= min(mtcars$hp)
```

### Exercise 10

Write R code to ask if the variance of weight (`wt`) is not equal to the standard deviation of weight of the cars within the `mtcars` data set.

*`var()` and `sd()` are the functions that perform the calculations*

```{r dataframe_logical_10, exercise = TRUE}

```
```{r dataframe_logical_10-solution}
var(mtcars$wt) != sd(mtcars$wt)
```


### Exercise 11

Also built into RStudio is a data set titled `iris`. The data was collected for a 1936 paper by R.A. Fisher titled, "The use of multiple measurements in taxonomic problems." The data set includes three iris species with 50 samples each as well as some properties about each flower. 

Run the pre-populated code to see the first five rows and see what properties of each flower are provided.

*Using `?` in front of functions and downloaded data sets provides informative documentation.*

```{r dataframe_logical_11, exercise = TRUE}
?iris
head(iris) 
```

### Exercise 12

Using the `iris` data set, ask if the standard deviation of sepal length is greater than that of sepal width.

```{r dataframe_logical_12, exercise = TRUE}

```
```{r dataframe_logical_12-solution}
sd(iris$Sepal.Length) > sd(iris$Sepal.Width)
```

### Exercise 13

Using the `iris` data set, ask if the sum of all sepal lengths is greater than 900 *OR* 1.2 is less than or equal to the median petal width.


```{r dataframe_logical_13, exercise = TRUE}

```
```{r dataframe_logical_13-solution}
sum(iris$Sepal.Length) > 900 | 1.2 <= median(iris$Petal.Width)
```

### Exercise 14

Using the `iris` data set and `range()`, write R code to find the ranges of each numeric column within `iris`.

Finding the range within the `diff()` function will provide the difference between the values. Now write R code to ask if the difference in range values for petal length is greater or equal to that of sepal width.

```{r dataframe_logical_14, exercise = TRUE}

```
```{r dataframe_logical_14-solution}
range(iris$Sepal.Length)
range(iris$Sepal.Width)
range(iris$Petal.Length)
range(iris$Petal.Width)
diff(range(iris$Petal.Length)) >= diff(range(iris$Sepal.Width))
```

### Exercise 15

The numeric columns within data sets can be made into arrays. Create an array named `ex15_array` from the column `Petal.Width` and write R code to ask for the summary statistics, using `summary()`. 

Write R code to ask if the summary statistics from the array is equal to the range of petal width values. Think about what values from the summary you expect to be true, and which ones false.

*Run the pre-populated code to see what the output of `summary()` looks like.

```{r dataframe_logical_15, exercise = TRUE}
summary(iris$Sepal.Length)
```
```{r dataframe_logical_15-solution}
ex15_array <- iris$Petal.Width
summary(ex15_array) == range(iris$Petal.Width)
```



