---
title: "Logical Operators"
author: "Jake Kovaleski"
output: 
  learnr::tutorial:
    theme: united
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
```

## Simple Operations

### Exercise 1

Write the R code required to ask whether 7 is greater than 5:

*The correct output will be `TRUE`.*

```{r simple_logical_1, exercise=TRUE}

```
```{r simple_logical_1-solution}
7 > 5 # OR
5 < 7
```

### Exercise 2

Write the R code required to ask whether 256 is equal to 256:

```{r simple_logical_2, exercise=TRUE}

```
```{r simple_logical_2-solution}
256 == 256
```

### Exercise 3

Write the R code required to ask whether 11 is not equal to 23:

```{r simple_logical_3, exercise=TRUE}


```
```{r simple_logical_3-solution}
11 != 23
```

### Exercise 4

Write the R code required to ask whether 28 is greater than or equal to 13:

```{r simple_logical_4, exercise=TRUE}

```
```{r simple_logical_4-solution}
28 >= 13 # OR
13 <= 28
```

### Exercise 5

Define the varibles `k` and `f` as 43 and 92, respectively. Ask whether `k` is greater than `f`:

```{r simple_logical_5, exercise=TRUE}


```
```{r simple_logical_5-solution}
f <- 92
k <- 43
k > f

```

### Exercise 6

Use `var1 <- "13 characters"` and `var2 <- "characters = 15"` and ask whether `var1` has more characters than `var2` in its string:

*`nchar()` is a function that asks how many characters are in a string. An example has been included in the answer chunk, run the code to see how it operates.*

```{r simple_logical_6, exercise=TRUE}
example <- "How many characters are in this string?"
nchar(example)
```
```{r simple_logical_6-solution}
var1 <- "13 characters"
var2 <- "characters = 15"
nchar(var1) > nchar(var2)
```

## Compound Operations 

### Exercise 1

Assign the value of 33 to `x`, then ask whether `x` is greater than 12 *AND* less than 146:

```{r compound_logical_1, exercise=TRUE}

```
```{r compound_logical_1-solution}
x = 33
x > 12 & x < 146
```

### Exercise 2

Assign 77 to `x` and 31 to `y`, ask whether `x` is greater than `y` OR `y` is greater than `x`:

```{r compound_logical_2, exercise=TRUE}

```
```{r compound_logical_2-solution}
x <- 77
y <- 31
x > y | y > x
```

### Exercise 3

Assign 47 to `z` and 93 to `w`. Ask whether `z` is less than `w` *AND* `w` is greater than or equal to `z` *AND* `z` is not equal to `w`:


```{r compound_logical_3, exercise= TRUE}


```
```{r compound_logical_3-solution}
z <- 47
w <- 93
z < w & w >= z & z != w

```

## Working with functions and logical

### Exercise 1

Assign an array from 1 to 10 to the varible `x`, then print the whole array:

``` {r function_logical_1, exercise = TRUE}

```
```{r function_logical_1-solution}
x <- 1:10
print(x)
```

### Exercise 2

Using `x = 1:10` again, ask whether the number 6 is within the array:

*You must assign the array to `x` again, each chunk is its own environment.*

```{r function_logical_2, exercise= TRUE}

```
```{r function_logical_2-solution}
x <- 1:10
6 %in% x
```

### Exercise 3

Make an array, `a`, ranging from 1:500, and ask whether that either 2 OR 6278 is within that array:

```{r function_logical_3, exercise=TRUE}

```
```{r function_logical_3-solution}
a <- 1:500
1 %in% a | 6278 %in% a
```

### Exercise 4

Make an array, `ex4_array`, ranging from 47 to 423, then find the length of the array:

*`length()` is the function whose output is the length of an string, array, column, ect.*

```{r function_logical_4, exercise=TRUE}

```
```{r function_logical_4-solution}
ex4_array <- 47:423
length(ex4_array)
```

### Exercise 5

Make an array, `ex4_array`, ranging from 73 to 632, and ask whether the length is equal to 560:

```{r function_logical_5, exercise=TRUE}

```
```{r function_logical_5-solution}
ex5_array <- 73:632
length(ex5_array) == 559
```

## Working with data frames and logical

### Exercsie 1

Confirm that there are 87 different Star Wars characters in the `starwars` dataset:

*The column title is `name`*

```{r dataframe_logical_1, exercise = TRUE}

```
```{r dataframe_logical_1-solution}
length(starwars$name)
```

### Exercise 2

Say you wanted to know if the `starwars` dataset included any data on Lando Calrissian, how could you quickly see if he was included?

```{r dataframe_logical_2, exercise = TRUE}

```
```{r dataframe_logical_2-solution}
"Lando Calrissian" %in% starwars$name
```




