---
title: "Logical Operators"
author: "Jake Kovaleski"
output: 
  learnr::tutorial:
    theme: united
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(dslabs)
knitr::opts_chunk$set(echo = FALSE)
```

## Simple Operations

### Exercise 1

Write the R code required to ask whether 7 is greater than 5.

*The correct output will be `TRUE`.*

```{r simple_logical_1, exercise=TRUE}

```
```{r simple_logical_1-solution}
7 > 5 # OR
5 < 7
```

### Exercise 2

Write the R code required to ask whether 256 is equal to 256.

```{r simple_logical_2, exercise=TRUE}

```
```{r simple_logical_2-solution}
256 == 256
```

### Exercise 3

Write the R code required to ask whether 11 is not equal to 23.

```{r simple_logical_3, exercise=TRUE}


```
```{r simple_logical_3-solution}
11 != 23
```

### Exercise 4

Write the R code required to ask whether 28 is greater than or equal to 13.

```{r simple_logical_4, exercise=TRUE}

```
```{r simple_logical_4-solution}
28 >= 13 # OR
13 <= 28
```

### Exercise 5

Define the varibles `k` and `f` as 43 and 92, respectively. Ask whether `k` is greater than `f`.

```{r simple_logical_5, exercise=TRUE}


```
```{r simple_logical_5-solution}
f <- 92
k <- 43
k > f

```

### Exercise 6

Use `var1 <- "13 characters"` and `var2 <- "characters = 15"` and ask whether `var1` has more characters than `var2` in its string.

*`nchar()` is a function that asks how many characters are in a string. An example has been included in the answer chunk, run the code to see how it operates.*

```{r simple_logical_6, exercise=TRUE}
example <- "How many characters are in this string?"
nchar(example)
```
```{r simple_logical_6-solution}
var1 <- "13 characters"
var2 <- "characters = 15"
nchar(var1) > nchar(var2)
```

## Compound Operations 

### Exercise 1

Ask whether 13 is less than 22 *OR* greater than 0.5.

```{r compound_logical_1, exercise=TRUE}

```
```{r compound_logical_1-solution}
13 < 22 | 13 > 0.5
```

### Exercise 2

Assign the value of 33 to `x`, then ask whether `x` is greater than 12 *AND* less than 146.

```{r compound_logical_2, exercise=TRUE}

```
```{r compound_logical_2-solution}
x <- 33
x > 12 & x < 146
```

### Exercise 3

Assign 77 to `x` and 31 to `y`. Ask whether `x` is greater than `y` OR `y` is greater than `x`.

```{r compound_logical_3, exercise=TRUE}

```
```{r compound_logical_3-solution}
x <- 77
y <- 31
x > y | y > x
```

### Exercise 4

Assign 47 to `z` and 93 to `w`. Ask whether `z` is less than `w` *AND* if `w` is greater than or equal to `z` *AND* if `z` is not equal to `w`.


```{r compound_logical_4, exercise= TRUE}


```
```{r compound_logical_4-solution}
z <- 47
w <- 93
z < w & w >= z & z != w

```

### Exercise 5

Assign `a`, `b`, and `c` as 1, 5, and 10, respectively. Ask whether `a` and `b` together are less than `c` *OR* `b` minus `c` is greater than `a`.

```{r compound_logical_5, exercise= TRUE}


```
```{r compound_logical_5-solution}
a <- 1
b <- 5
c <- 10
(a + b) < c | (b - c) > a

```

### Exercise 6

Assign "data science" to `strng1`, and "for biologists" to `strng2`. Ask whether the number of characters in `strng1` is equal to 12 *AND* if `strng1` is less than or equal to `strng2`.

```{r compound_logical_6, exercise= TRUE}


```
```{r compound_logical_6-solution}
strng1 <- "data science"
strng2 <- "for biologists"
nchar(strng1) == 12 & strng1 <= strng2

```

## Working with functions and logical

### Exercise 1

Assign an array from 1 to 10 to the varible `x`, then print the whole array.

``` {r function_logical_1, exercise = TRUE}

```
```{r function_logical_1-solution}
x <- 1:10
print(x)
```

### Exercise 2

Using `x = 1:10` again, ask whether the number 6 is within the array.

*You must assign the array to `x` again, each chunk is its own environment.*

```{r function_logical_2, exercise= TRUE}

```
```{r function_logical_2-solution}
x <- 1:10
6 %in% x
```

### Exercise 3

Make an array, `a`, ranging from 1:500, and ask whether that either 2 OR 6278 is within that array.

```{r function_logical_3, exercise=TRUE}

```
```{r function_logical_3-solution}
a <- 1:500
1 %in% a | 6278 %in% a
```

### Exercise 4

Make an array, `ex4_array`, ranging from 47 to 423, then find the length of the array.

*`length()` is the function whose output is the length of an string, array, column, ect.*

```{r function_logical_4, exercise=TRUE}

```
```{r function_logical_4-solution}
ex4_array <- 47:423
length(ex4_array)
```

### Exercise 5

Make an array, `ex5_array`, ranging from 73 to 632, and ask whether the length is equal to 560.

```{r function_logical_5, exercise=TRUE}

```
```{r function_logical_5-solution}
ex5_array <- 73:632
length(ex5_array) == 559
```

### Exercise 6

Make an array, `ex6_array`, ranging from 31 to 489, and another, `anthr_array`, ranging from 123 to 589. Ask whether the length of `ex6_array` is less than or equal to `2nd_array`.

```{r function_logical_6, exercise=TRUE}

```
```{r function_logical_6-solution}
ex6_array <- 31:489
anthr_array <- 123:589
length(ex6_array) <= length(anthr_array)
```


## Working with data frames and logical

### Exercsie 1

Confirm that there are 87 different Star Wars characters in the `starwars` dataset.

*The column title is `name`*

```{r dataframe_logical_1, exercise = TRUE}

```
```{r dataframe_logical_1-solution}
length(starwars$name)
```

### Exercise 2

Say you wanted to know if the `starwars` dataset included any data on Lando Calrissian, how could you quickly see if he was included?

```{r dataframe_logical_2, exercise = TRUE}

```
```{r dataframe_logical_2-solution}
"Lando Calrissian" %in% starwars$name
```

### Exercise 3

Determine what types of columns `hair_color` and `mass` are in the `starwars` dataset.

```{r dataframe_logical_3, exercise = TRUE}

```
```{r dataframe_logical_3-solution}
class(starwars$mass)
class(starwars$hair_color)
```

### Exercise 4

Someone tells you the average quater mile time of the cars within the `mtcars` dataset is 17.84875 seconds. Ask whether the average is indeed equal to 17.84875 seconds.

*The code to view the first 5 lines of `mtcars` has been provided*

```{r dataframe_logical_4, exercise = TRUE}
head(mtcars)
```
```{r dataframe_logical_4-solution}
mean(mtcars$qsec) == 17.84875
```

### Exercise 5

Ask whether the median MPG of the cars within the `mtcars` dataset is less than or equal to 20 MPG.

```{r dataframe_logical_5, exercise = TRUE}

```
```{r dataframe_logical_5-solution}
median(mtcars$mpg) <= 20
```

### Exercise 6

Using the iris dataset, ask if the standard deviation of sepal length is greater than the standard deviation of sepal width.

```{r dataframe_logical_6, exercise = TRUE}
head(iris)
```
```{r dataframe_logical_6-solution}
sd(iris$Sepal.Length) > sd(iris$Sepal.Width)
```



