---
title: "Logical Operators"
author: "Jake Kovaleski"
output: 
  learnr::tutorial:
    theme: united
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
```

## Simple Operations

### Exercise 1

Write the R code required to ask whether 7 is greater than 5:

*The correct output will be `TRUE`.*

```{r simple_logical_1, exercise=TRUE}

```
```{r simple_logical_1-solution}
7 > 5 # OR
5 < 7
```

### Exercise 2

Write the R code required to ask whether 256 is equal to 256:

```{r simple_logical_2, exercise=TRUE}

```
```{r simple_logical_2-solution}
256 == 256
```

### Exercise 3

Write the R code required to ask whether 11 is not equal to 23:

```{r simple_logical_3, exercise=TRUE}


```
```{r simple_logical_3-solution}
11 != 23
```

### Exercise 4

Write the R code required to ask whether 28 is greater than or equal to 13:

```{r simple_logical_4, exercise=TRUE}

```
```{r simple_logical_4-solution}
28 >= 13 # OR
13 <= 28
```

### Exercise 5

Ask whether `k` is greater than `f`:

```{r simple_logical_5, exercise=TRUE}
f = 92
k = 43

```
```{r simple_logical_5-solution}
f = 92
k = 43
k > f

```

### Exercise 6

Ask whether `moon` has more characters than `sun` in its string:

```{r simple_logical_6, exercise=TRUE}
sun <- "13 characters"
moon <- "characters = 15"

```
```{r simple_logical_6-solution}
sun <- "13 characters"
moon <- "characters = 15"
nchar(moon) > nchar(sun)
```

## Compound Operations 

### Exercise 1

Assign the value of 33 to `x`, then write the R code necessary to ask whether `x` is greater than 12 AND `x` is less than 146:

```{r compound_logical_1, exercise=TRUE}

```
```{r compound_logical_1-solution}
x = 33
x > 12 & x < 146
```

### Exercise 2

Assign `x = 77` and `y = 31`, ask whether `x` is greater than `y` OR `y` is greater than `x`:

```{r compound_logical_2, exercise=TRUE}

```
```{r compound_logical_2-solution}
x = 77
y = 31
x > y | y > x
```

### Exercise 3

Make `x = 47` and `y = 93`. Use only the varibles, `&`, `<`, `>`, and `!=` to get a `TRUE` output:


```{r compound_logical_3, exercise= TRUE}


```
```{r compound_logical_3-solution}
x = 47
y = 93
x < y & y > x & x != y

```

## Working with functions and logical

### Exercise 1

Assign an array from 1 to 10 to the varible `x`, then visiualize the whole array:

``` {r function_logical_1, exercise = TRUE}

```
```{r function_logical_1-solution}
x = 1:10
print(x)
```

### Exercise 2

Using `x = 1:10` again, ask whether the number 6 is within the array:

```{r function_logical_2, exercise= TRUE}

```
```{r function_logical_2-solution}
x = 1:10
6 %in% x
```

### Exercise 3

Using `x = 1:10` once more, ask whether that either 2 OR 6278 is within that array:

```{r function_logical_3, exercise=TRUE}

```
```{r function_logical_3-solution}
x = 1:10
1 %in% x | 6278 %in% x
```

## Working with data frames and logical

### Exercsie 1

Confirm that there are 87 different Star Wars characters in the `starwars` dataset:

*The column title is `name`*

```{r dataframe_logical_1, exercise = TRUE}

```
```{r dataframe_logical_1-solution}
length(starwars$name)
```

### Exercise 2

Say you wanted to know if the `starwars` dataset included any data on Lando Calrissian, how could you quickly see if he was included?

```{r dataframe_logical_2, exercise = TRUE}

```
```{r dataframe_logical_2-solution}
"Lando Calrissian" %in% starwars$name
```




