---
title: "regex module"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(stringr)
knitr::opts_chunk$set(echo = FALSE)
```

Write a regex that will match four numbers followed by three letters The regex should match all of these:

+ "4562ARE"
+ "1100llk"
+ "0000AbC"


```{r regex, exercise=TRUE}

```

```{r regex-solution}
reg <- "\\d{4}[A-z]{3}"
```

Similar to last question, write same regex making that this is not matched: `7251p_i`

```{r regex2, exercise=TRUE}

```

```{r regex2-solution}
reg <- "\\d{4}[A-z]{1}\\w[A-z]{1}"
```

Write a regex matching all of these numbers "10.45", "11.325", "14.90192", "100.2421", "88.1" but not "90" .

Hint: These are all numbers that have a decimal point with different amounts of digits before and after the decimal.



```{r regex3, exercise=TRUE}

```

```{r regex3-solution}
reg <- "\\d+\\.\\d+"

```

Write a regex that matches all the following phrases:

regex fun
Regex Fun
REGEX FUN
regex Fun

```{r regex4, exercise=TRUE}

```

```{r regex4-solution}
reg <- "[A-z]+ [A-z]+"

reg <- "[A-z]{5} [A-z]{3}"

```



Write a regex that matches any sentence that contains five words and ends with a period, assuming sentences always start with a capital letter.

"This should totally work well."
"this one should not match!"
"this can not match either."
"This should also work well."
"This has way too many words too many"
"Not enough."
"CAPS CAN ALSO MATCH ALSO."

```{r regex5, exercise=TRUE}

```

```{r regex5-solution}
reg <- "[A-z]+ [A-z]+ [A-z]+ [A-z]+ [A-z]+\\."
```


Write a regex that matches any sentence that contains a number, followed by three words, followed by a number, followed by one word and ending with a period.
"8 hot dogs and 9 lemonades."
"5 water bottles and 8 hamburgers."
"7 soup bowls and 7 spoons."

```{r regex6, exercise=TRUE}

```

```{r regex6-solution}
reg <- "\\d{1} [a-z]+ [a-z]+ [a-z]+ \\d{1} [a-z]+\\."
```

Similar to the previous question, write a regex that matches any phrase containing a two or three digit number, followed by two words.

"12 oak trees"
"62 green apples"
"19 red berries"


```{r regex7, exercise=TRUE}

```


```{r regex7-solution}
reg <- "\\d{2,3} [a-z]+ [a-z]+"
```

Write a regex matching both 9.034 and 9.56.

```{r regex8, exercise=TRUE}

```

```{r regex8-solution}
goal <- 
reg <- "\\d\\.\\d+"
  
goal <-
reg <- "^\\d\\.\\d{2,3}$"
```

Write a regex matching both "I like coding." and "Coding is fun."

```{r regex9, exercise=TRUE}

```


```{r regex9-solution}
goal <-
reg <- "[A-z]+ [a-z]+ [a-z]+\\."
  
goal <-
reg <- "[A-z]+ \\w{2,4} \\w{3,6}\\."

goal <-
reg<- "\\w+ \\w+ \\w+\\."

goal <- 
reg <- "\\w{1,6} \\w{2,4} \\w{3,6}\\."
```

Write a regex matching any sentence with a price, followed by a word, followed by a number, followed by a word and ending in a period.

"$12.98 for 7 flowers."
"$20.00 with 3 purchases."


```{r regex10, exercise=TRUE}


```

```{r regex10-solution}
reg <- "\\$\\d+\\.\\d+ [A-z]+ \\d+ [A-z]+\\."

```

Write a regex turning this strand of DNA into RNA.

ATAGCTAGGCTATTCGTA

> Hint: In RNA, the T's in DNA turn to U's. In other words, change the T's in the sequence above to U's.

```{r regex11, exercise=TRUE}

```

```{r regex11-solution}
str_replace("ATAGCTAGGCTATTCGTA", (([A])([T])([A])([G])([C])([T])([A])([G])([G])([C])([T])([A])([T])([T])([C])([G])([T])([A])), "\\2611131416U")
```

Write a regex extracting all lowercase letters a-t in the sentence: "This sentence contains some a's, t's, and everything in between!"

```{r regex12, exercise=TRUE}

```

```{r regex12-solution}
str_extract_all("This sentence contains some a's, t's, and everything in between!", "[a-t]")
```

Write a regex expression extracting all numbers from the phrase "I have 3 dogs and 4 cats".

```{r regex13, exercise=TRUE}

```

```{r regex13-solution}
str_extract_all("I have 3 dogs and 4 cats", "\\d+")
```



Write a regex extracting all numbers from the phrase "I have 12 birds and 10 fish". (12 should print as "12" and not "1" "2")

```{r regex14, exercise=TRUE}

```

```{r regex14-solution}
str_extract_all("I have 12 birds and 10 fish", "\\d\\d")
```

Write a regex extracting all prices from the sentence: "We bought two flowers for $4.99 each and two pots for $2.99 each, so our total was $15.96."

```{r regex15, exercise=TRUE}

```

```{r regex15-solution}
str_extract_all("We bought two flowers for $4.99 each and two pots for $2.99 each, so our total was $15.96.", "\\$\\d+\\.\\d+")

```


Write a regex expression to replace 9.0998 with 9.1.

```{r regex16, exercise=TRUE}

```

```{r regex16-solution}
str_replace("9.0998", "(\\d\\.).+", "\\11")

str_replace("9.0998", "(\\d)\\..+", "\\1.1")
```

Write a regex  searching and replacing the numbers in the sentence "I have 6 cats." with the numbers spelled out. (e.g., 6 = six)

```{r regex17, exercise=TRUE}

```

```{r regex17-solution}
str_replace("I have 6 cats.","[A-Z] [a-z] (\\d) [a-z]\\.", "\\1six" )
```



Write a regex expression searching and replacing the word "cats" in the phrase "I have 2 cats" with the word "fish".

```{r regex18, exercise=TRUE}

```

```{r regex18-solution}
str_replace("I have 2 cats", "[A-Z]+ [a-z]+ \\d+ ([a-z]+)", "\\1fish")
```

Write a regex searching and replacing the lowercase letters are the beginning of each word in the city "las vegas" with their uppercase counterpart.

```{r regex19, exercise=TRUE}
str_replace("las vegas", "([a-z]?)[a-z]+ ([a-z]?)[a-z]+", "\\12[A-Z]")
```

```{r regex19-solution}

```

Write a regex expression to add hyphens in the word "mother in law" where spaces are currently present.

```{r regex20, exercise=TRUE}

```

```{r regex20-solution}
str_replace("mother in law", "[a-z]+( )[a-z]+( ) [a-z]+","\\12\\-")
```



Write a regex expression searching and replacing hyphens in the cities "Las-Vegas" and "San-Juan" with spaces. 

```{r regex21, exercise=TRUE}

```

```{r regex21-solution}
str_replace("Las-Vegas, San-Juan", "[A-z]+(\\-)[A-z]+", "\\1\\ ")
```



Write a regex expression joining the two components of the words "ear ring", "dish washer" and "earth worm" to create their individual compound words.

```{r regex22, exercise=TRUE}

```

```{r regex22-solution}
str_replace("ear ring", "[A-z]+( )[A-z]+", "\\1")
```


Write a regex expression searching and pattern matching the decimals 12.624, 14.83 and 16.2838 to the format of 12.8.

```{r regex23, exercise=TRUE}

```

```{r regex23-solution}
str_replace("12.624, 14.83, 16.2838", "\\d{2}\\.(\\d+)", "\\1\\d?")
```


Write a regex expression searching and pattern matching the cities "las_vegas", "san_antonio", "san_juan" and "san_francisco" to match the format Los Angeles. 

```{r regex24, exercise=TRUE}

```

```{r regex24-solution}
str_replace()
```

Change "Genus species" to "G. species"

Ginko biloba to G. biloba
Equus caballus to E. caballus
Eretmochelys imbricata to E. inbricata

```{r regex25, exercise=TRUE}

```

```{r regex25-solution}

```







