---
title: "regex module"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(stringr)
knitr::opts_chunk$set(echo = FALSE)
```

Write a regex that will match four numbers followed by three letters The regex should match all of these:

+ "4562ARE"
+ "1100llk"
+ "0000AbC"


```{r regex, exercise=TRUE}

```

```{r regex-solution}
reg <- "\\d{4}[A-z]{3}"
```

Similar to last question, write same regex making that this is not matched: `7251p_i`

```{r regex1, exercise=TRUE}

```

```{r regex1-solution}
reg <- "\\d{4}[A-z]{1}\\.[A-z]{1}"
```

Write a regex matching all of these numbers "10.45", "11.325", "14.90192", "100.2421", "88.1" but not "90" .

Hint: These are all numbers that have a decimal point with different amounts of digits before and after the decimal.



```{r regex2, exercise=TRUE}
reg <- "\\d+\\.\\d+"
```

```{r regex2-solution}


```


Write a regex that matches any sentence that contains five words and ends with a period, assuming sentences always start with a capital letter.

"This should totally work well."
"this one should not match!"
"this can not match either."
"This should also work well."
"This has way too many words too many"
"Not enough."
"CAPS CAN ALSO MATCH ALSO."

```{r regex3, exercise=TRUE}

```

```{r regex3-solution}
reg <- "[A-z]+ [A-z]+ [A-z]+ [A-z]+ [A-z]+\\."
```


Write a regex matches any sentence that contains a number, followed by three words, followed by a number, followed by one word and ending with a period.
"8 hot dogs and 9 lemonades."
"5 water bottles and 8 hamburgers."
"7 soup bowls and 7 spoons."

```{r regex4, exercise=TRUE}

```

```{r regex4-solution}
reg <- "\\d{1} [a-z]+ [a-z]+ [a-z]+ \\d{1} [a-z]+\\."
```

Similar to the previous question, write a regex that matches any phrase containing a two or three digit number, followed by two words.

"12 oak trees"
"62 green apples"
"19 red berries"


```{r regex5, exercise=TRUE}

```


```{r regex5-solution}
reg <- "\\d{2,3} [a-z]+ [a-z]+"
```

Write a regex matching both 9.034 and 9.56.

```{r regex6, exercise=TRUE}

```

```{r regex6-solution}
goal <- 
reg <- "\\d\\.\\d+"
  
goal <-
reg <- "^\\d\\.\\d{2,3}$"
```

Write a regex matching both "I like coding." and "Coding is fun."

```{r regex7, exercise=TRUE}

```


```{r regex7-solution}
goal <-
reg <- "[A-z]+ [a-z]+ [a-z]+\\."
  
goal <-
reg <- "[A-z]+ \\w{2,4} \\w{3,6}\\."

goal <-
reg<- "\\w+ \\w+ \\w+\\."

goal <- 
reg <- "\\w{1,6} \\w{2,4} \\w{3,6}\\."
```

Write 

```{r regex8, exercise=TRUE}


```

```{r regex8-solution}


```

Write a regex  extracting all lowercase letters from the city "New York City".

```{r regex9, exercise=TRUE}

```

```{r regex9-solution}

```
Write a regex extracting all uppercase characters A-T and all lowercase letters a-t.

```{r regex8, exercise=TRUE}

```

Write a regex expression extracting all numbers from the phrase "I have 3 dogs and 4 cats".

```{r regex9, exercise=TRUE}

```

Write a regex expression extracting all numbers from the phrase "I have 12 birds and 10 fish". (12 should print as "12" and not "1" "2")

```{r regex10, exercise=TRUE}

```


Write a regex searching and replacing every capital letter with its lowercase counterpart in the phrase "She Has Science Homework".

```{r regex11, exercise=TRUE}

```

```{r regex11-solution}
str_replace("She Has Science Homework", "([A-Z]+)[a-z]+ ([A-Z]+)[a-z]+ ([A-Z]+)[a-z]+ ([A-Z]+)[a-z]+", "\\[A-Z][a-z]")

#str_to_lower("She Has Science Homework")

```


Write a regex expression to replace 9.0998 with 9.1.

```{r regex12, exercise=TRUE}
str_replace("9.0998", "(\\d\\.).+", "\\11")

str_replace("9.0998", "(\\d)\\..+", "\\1.1")
```

Write a regex  searching and replacing the numbers in the phrase "I have 6 cats and 7 dogs" with the numbers spelled out. (e.g., 6 = six)

```{r regex13, exercise=TRUE}

```

Write a regex expression searching and replacing the word "cats" in the phrase "I have 2 cats" with the word "fish".

```{r regex14, exercise=TRUE}
str_replace("I have 2 cats", "[A-Z]+ [a-z]+ \\d+ ([a-z]+)", "\\1fish")
```


Write a regex searching and replacing the lowercase letters are the beginning of each word in the city "las vegas" with their uppercase counterpart.

```{r regex15, exercise=TRUE}
str_replace("lax vegas", "([a-z]?)[a-z]+ ([a-z]?)[a-z]+", "\\12[A-Z]")
```

Write a regex expression to add hyphens in the word "mother in law" where spaces are currently present.

```{r regex16, exercise=TRUE}

```

Write a regex expression searching and replacing hyphens in the cities "Las-Vegas" and "San-Juan" with spaces. 

```{r regex17, exercise=TRUE}

```

Write a regex expression joining the two components of the words "ear ring", "dish washer" and "earth worm" to create their individual compound words.

```{r regex18, exercise=TRUE}

```

Write a regex expression searching and pattern matching the decimals 12.624, 14.83 and 16.2838 to the format of 12.8.

```{r regex19, exercise=TRUE}

```

Write a regex expression searching and pattern matching the cities "las_vegas", "san_antonio", "san_juan" and "san_francisco" to match the format Los Angeles. 

```{r regex20, exercise=TRUE}

```



Change "Genus species" to "G. species"



turn this DNA into RNA 





