---
title: "regex module"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(stringr)
knitr::opts_chunk$set(echo = FALSE)
```


Write a regex expression explaining the phrase "Regex module for beginners".

```{r regex1, exercise=TRUE}

```

Write a regex expression explaining the number 17.993

```{r regex2, exercise=TRUE}

```


Write a regex expression explaining the phrase "8 hot dogs and 9 lemonades".

```{r regex3, exercise=TRUE}

```



Write a regex expression explaining the relationship between 9.034 and 9.56.

```{r regex4, exercise=TRUE}

```


Write a regex expression explaining the relationship between "I like coding" and "Coding is fun".

```{r regex5, exercise=TRUE}

```

Write a regex expression extracting all uppercase letters from the city "New York City".

```{r regex6, exercise=TRUE}

```

Write a regex expression extracting all lowercase letters from the city "New York City".

```{r regex7, exercise=TRUE}

```

Write a regex expression extracting all uppercase characters A-T and all lowercase letters a-t.

```{r regex8, exercise=TRUE}

```

Write a regex expression extracting all numbers from the phrase "I have 3 dogs and 4 cats".

```{r regex9, exercise=TRUE}

```

Write a regex expression extracting all numbers from the phrase "I have 12 birds and 10 fish". (12 should print as "12" and not "1" "2")

```{r regex10, exercise=TRUE}

```


Write a regex expression searching and replacing every capital letter with its lowercase counterpart in the phrase "She Has Science Homework".

```{r regex11, exercise=TRUE}

```

Write a regex expression to replace 9.0998 with 9.1.

```{r regex12, exercise=TRUE}

```

Write a regex expression searching and replacing the numbers in the phrase "I have 6 cats and 7 dogs" with the numbers spelled out. (e.g., 6 = six)

```{r regex13, exercise=TRUE}

```

Write a regex expression searching and replacing the word "cats" in the phrase "I have 2 cats" with the word "fish".

```{r regex14, exercise=TRUE}

```


Write a regex expression searching and replacing the lowercase letters are the beginning of each word in the city "las vegas" with their uppercase counterpart.

```{r regex15, exercise=TRUE}

```

Write a regex expression to add hyphens in the word "mother in law" where spaces are currently present.

```{r regex16, exercise=TRUE}

```

Write a regex expression searching and replacing hyphens in the cities "Las-Vegas" and "San-Juan" with spaces. 

```{r regex17, exercise=TRUE}

```

Write a regex expression joining the two components of the words "ear ring", "dish washer" and "earth worm" to create their individual compound words.

```{r regex18, exercise=TRUE}

```

Write a regex expression searching and pattern matching the decimals 12.624, 14.83 and 16.2838 to the format of 12.8.

```{r regex19, exercise=TRUE}

```

Write a regex expression searching and pattern matching the cities "las_vegas", "san_antonio", "san_juan" and "san_francisco" to match the format Los Angeles. 

```{r regex20, exercise=TRUE}

```










