---
title: "regex module"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(stringr)
knitr::opts_chunk$set(echo = FALSE)
```


Write a regex matching the phrase "Regex module for beginners".

```{r regex1, exercise=TRUE}

```

```{r regex1-solution}

goal <- "Regex module for beginners"
reg <- "^[A-z]+ [a-z]{6,} [a-z]{3,} [a-z]{9,}"


goal <- "Regex module for beginners"
reg <- "[A-Z, a-z]+ [a-z]+ [a-z]+ [a-z]+"


```




Write a regex matching the number 17.993

```{r regex2, exercise=TRUE}

```

```{r regex2-solution}

goal <- "17.993"
reg <- "\\d+\\.\\d+"

goal <- "17.993"
reg <- "\\d{,3}\\.\\d+"

```




Write a regex matching the phrase "8 hot dogs and 9 lemonades".

```{r regex3, exercise=TRUE}

```

```{r regex3-solution}
goal <- "8 hot dogs and 9 lemonades"
reg <- "\\d? [a-z]+ [a-z}+ [a-z]+ \\d? [a-z]+"

goal <- "8 hot dogs and 9 lemonades"
reg <- "^\\d+ \\w+ \\w+ \\w+ \\d+ \\w+$ "
```

Write a regex matching the relationship between 9.034 and 9.56.

```{r regex4, exercise=TRUE}
goal <- 
reg <- "\\d\\.\\d+"
  
goal <-
reg <- "^\\d\\.\\d{2,3}$"
```


Write a regex matching the relationship between "I like coding" and "Coding is fun".

```{r regex5, exercise=TRUE}
goal <-
reg <- "[A-z]+ [a-z]+ [a-z]+"
  
goal <-
reg <- "[A-z]+ \\w{2,4} \\w{3,6}"

goal <-
reg<- "\\w+ \\w+ \\w+"

goal <- 
reg <- "\\w{1,6} \\w{2,4} \\w{3,6}"
```

Write a regex extracting all uppercase letters from the city "New York City".

```{r regex6, exercise=TRUE}


```

```{r regex6-solution}


```

Write a regex  extracting all lowercase letters from the city "New York City".

```{r regex7, exercise=TRUE}

```

```{r regex7-solution}

```
Write a regex extracting all uppercase characters A-T and all lowercase letters a-t.

```{r regex8, exercise=TRUE}

```

Write a regex expression extracting all numbers from the phrase "I have 3 dogs and 4 cats".

```{r regex9, exercise=TRUE}

```

Write a regex expression extracting all numbers from the phrase "I have 12 birds and 10 fish". (12 should print as "12" and not "1" "2")

```{r regex10, exercise=TRUE}

```


Write a regex searching and replacing every capital letter with its lowercase counterpart in the phrase "She Has Science Homework".

```{r regex11, exercise=TRUE}

```

```{r regex11-solution}
str_replace("She Has Science Homework", "([A-Z]+)[a-z]+ ([A-Z]+)[a-z]+ ([A-Z]+)[a-z]+ ([A-Z]+)[a-z]+", "\\[A-Z][a-z]")


```


Write a regex expression to replace 9.0998 with 9.1.

```{r regex12, exercise=TRUE}
str_replace("9.0998", "(\\d\\.).+", "\\11")

str_replace("9.0998", "(\\d)\\..+", "\\1.1")
```

Write a regex  searching and replacing the numbers in the phrase "I have 6 cats and 7 dogs" with the numbers spelled out. (e.g., 6 = six)

```{r regex13, exercise=TRUE}

```

Write a regex expression searching and replacing the word "cats" in the phrase "I have 2 cats" with the word "fish".

```{r regex14, exercise=TRUE}
str_replace("I have 2 cats", "[A-Z]+ [a-z]+ \\d+ ([a-z]+)", "\\1fish")
```


Write a regex searching and replacing the lowercase letters are the beginning of each word in the city "las vegas" with their uppercase counterpart.

```{r regex15, exercise=TRUE}
str_replace("lax vegas", "([a-z]?)[a-z]+ ([a-z]?)[a-z]+", "\\12[A-Z]")
```

Write a regex expression to add hyphens in the word "mother in law" where spaces are currently present.

```{r regex16, exercise=TRUE}

```

Write a regex expression searching and replacing hyphens in the cities "Las-Vegas" and "San-Juan" with spaces. 

```{r regex17, exercise=TRUE}

```

Write a regex expression joining the two components of the words "ear ring", "dish washer" and "earth worm" to create their individual compound words.

```{r regex18, exercise=TRUE}

```

Write a regex expression searching and pattern matching the decimals 12.624, 14.83 and 16.2838 to the format of 12.8.

```{r regex19, exercise=TRUE}

```

Write a regex expression searching and pattern matching the cities "las_vegas", "san_antonio", "san_juan" and "san_francisco" to match the format Los Angeles. 

```{r regex20, exercise=TRUE}

```










