---
title: "Getting help in the `{introverse}`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting help in the `{introverse}`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


Before you get started, make sure the library is loaded into your R session by running `library(introverse)`. Once loaded, you'll see welcome prompt showing some basic usage. 

```{r setup}
library(introverse)
```



### What topics can I get help on?

To see all currently available help topics, run the following command. The function `show_topics()` will print all topic _categories_ and all associated topics within each category. 

```{r, eval=F}
show_topics()
```
Or, all currently available topics can be seen [here](https://spielmanlab.github.io/introverse/articles/available_topics.html).


<br>


### Woah, there's a lot of topics! Can I just see the topic categories to narrow it down?

Yes, you can! Use this function:

```{r, eval=F}
show_categores()
```

<br>

### I know I want help using `{dplyr}`. Can I see only the `{dplyr}` help topics?

Absolutely! You can see _only_ those topics for a given category with `show_topics("category of interest")`. For example, to see the `{dplyr}` topics, you would run...

```{r, eval=F}
show_topics("dplyr") # Don't forget to use quotes!
```

<br>

### I know the topic I want help for! Show me how to get help already!!

Use the function `get_help()` to get help for a particular topic. For example, to get help using the [`filter()`](https://dplyr.tidyverse.org/reference/filter.html) function in `{dplyr}`, you can run:

```{r, eval=F}
get_help("filter")  # Don't forget to use quotes!
```

If you prefer for the help to automatically appear in your browser instead of in the RStudio Viewer pane, you can use the _second argument_ `browser = TRUE` when using `get_help()`. For example:

```{r, eval=F}
get_help("filter", browser = TRUE)  # Don't forget to use quotes!
```

<br>

### Can I learn a little more about the example datasets?

Definitely! We use the dataset `msleep` from the `{ggplot2}` package, as well as a subsetted version of this dataset called `carnivores`, for many of the `{introverse}` docs examples. Learn more about these datasets by getting help for them with:

```{r, eval=F}
get_help("msleep")
# and
get_help("carnivores")
```

Or, directly type `msleep` and/or `carnivores` into the R Console to see the datasets directly.

<br>

### This is all a little confusing. What if I get lost?

Don't worry, the `{introverse}` is here to help! We try to provide helpful feedback as much as possible if you use one of the `{introverse}` functions incorrectly or provide a "bad" argument. Read all of the messages you get carefully, and if you're still confused, let us know how we can improve by filing an [issue](https://github.com/spielmanlab/introverse/issues)!