---
title: "Customizing themes in `ggplot2`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{themes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  fig.width = 4,
  fig.height = 3
)
library(introverse)
library(ggplot2)
library(magrittr)
library(patchwork)
```

We use the term *theme* in the grammar of graphics to refer to all non-data components of a plot, including things like "where is the title placed?", "how large is the axis text?", "does the plot have a grid and how spaced out is the grid?", "is there a background fill to the plot canvas itself?", etc.

Themes are comprised of several different types of *elements*:

+ **`element_text()`: How do certain text elements appear?**
  + For example, `axis.title` (axes titles) is an `element_text()`. 
  + Among other options, you can change their:
    + `color`
    + `size`
    + `face` : "plain", "italic", "bold", or "bold.italic)
    + `angle`
    + `hjust` : "horizontal justification" where 0 is far left, 0.5 is middle, and 1 is far right, and anything in between!
    + `vjust` : "vertical justification" where 0 is bottom, 0.5 is middle, and 1 is top, and anything in between!)
    
<br>  
    
+ **`element_line()`: How do certain line elements appear?**
  + For example, `axis.line` (axis lines themselves) is an `element_line()`. 
  + Among other options, you can change their `color`, `size`, `linetype` ("solid", "dashed", "dotted", "dotdash", "longdash", and "twodash"), and `lineend` ("round", "butt", "square").

<br>  
  
+ **`element_rect()`: How do certain _borders and background_ elements in the plot appear?** 
  + For example, `plot.background` (the background/"canvas" appearance of the plot) is an `element_rect()`. 
  + Among other options, you can change their:
    + `color`
    + `fill`
    + `size` (of the outline)
    + `linetype` ("solid", "dashed", "dotted", "dotdash", "longdash", and "twodash")
  
<br>

+ **`unit()`**: Some theme components are not strictly elements but are defined in terms of size only and will be termed `unit()` or `grid::unit()`.



<br>

All theme components can be fully customized, as is overwhelmingly documented [here](https://ggplot2.tidyverse.org/reference/theme.html). There are also several [existing complete themes]() in `{ggplot2}`, which can be used "out of the box" if you want something different but without further customization, examples for which are shown in this tutorial.

<br><br>

## Customizing a theme


All examples in this tutorial will modify a version of this plot, which uses the default `theme_gray()` with no further customization. The plot is made with the `msleep` dataset. **Note that the examples generally do not make well-styled plots! The modified theme elements are intentionally strong to demonstrate the concepts, not because they look pretty.**

```{r plot}
ggplot(msleep) + 
  aes(x = awake) + 
  geom_histogram()
```

<br>

When customizing a theme, you need to:

+ Search through [the documentation](https://ggplot2.tidyverse.org/reference/theme.html) (literally) to find how `ggplot2` refers to the particular plot component you want to change. 
+ Use [the documentation](https://ggplot2.tidyverse.org/reference/element.html) (sensing a pattern?) to determine whether that component goes with `element_text()`, `element_line()`, or `element_rect()`. 
  + Or, decide if you want to *entirely remove* that particular plot component. Stay tuned...
+ *Add it* to your plot inside the `theme()` function.
  + If you are also specifying a complete theme in your plot code but want to change certain components, make sure to add these components *after* you specify the complete theme. Otherwise, the complete theme settings will *override* your customizations.
+ There is a *very very VERY* useful function you can use as part of the code called `rel()`. This function will help you change the size of different elements *relative to the default baseline*.
  + For example, `rel(1.1)` means 10% larger than default (110% the size of default), and `rel(0.9)` means 10% smaller than the default (90% the size of default). 

### Quick examples of `element_text()`

Let's see some examples using `axis.text`^[As you will notice when perusing the documentation, there is also `axis.text.x` and `axis.text.y` which can be used to separately customize X and Y axis text.], which controls both X and Y axis *text* (aka the labels along the axes, NOT the title). Customizing `axis.text` must be paired with `element_text()`, as it is text!

<br><br>

Make the axis text 50% larger than its default:
```{r}
ggplot(msleep) + 
  aes(x = awake) + 
  geom_histogram() + 
  #theme( plot-thing-to-change = element_<TYPE>(things to customize about it) )
  theme(axis.text = element_text(size = rel(1.5)))
```

<br>
Make the axis text 50% smaller than its default:
```{r}
ggplot(msleep) + 
  aes(x = awake) + 
  geom_histogram() + 
  theme(axis.text = element_text(size = rel(0.5)))
```


<br>
Make the axis text red, *and* 15% larger than default:
```{r}
ggplot(msleep) + 
  aes(x = awake) + 
  geom_histogram() +  
  theme(axis.text = element_text(color = "red", size = rel(1.15)))
```

<br>
Let's change a whole bunch of things! Change baseline theme to `theme_light()`, make BOTH X- and Y- axis have 15% larger font size, make the X-axis text blue, and make the Y-axis text purple.
```{r}
ggplot(msleep) + 
  aes(x = awake) + 
  geom_histogram() + 
  theme_light() +
  theme(axis.text = element_text(size = rel(1.15)), # both axes 
        axis.text.x = element_text(color = "blue"), # only the x!
        axis.text.y = element_text(color = "purple")) # only the y!
```

<br>
You can **REMOVE** theme elements entirely with `element_blank()`. The example below is obviously bad plot design, but shows how to get the job done if you wanted to!
```{r}
ggplot(msleep) + 
  aes(x = awake) + 
  geom_histogram() + 
  theme(axis.text = element_blank()) 
```

<br>
Rotate the text by a 90% angle
```{r}
ggplot(msleep) + 
  aes(x = awake) + 
  geom_histogram() + 
  theme(axis.text = element_text(angle = 90))
```


### Quick examples of `element_rect()`

Change the background fill of the plot:
```{r}
ggplot(msleep) + 
  aes(x = awake) + 
  geom_histogram() + 
  theme(plot.background = element_rect(fill = "pink"))
```

<br>
Change the background fill of the plot *PANEL* with `panel.backgroud`:
```{r}
ggplot(msleep) + 
  aes(x = awake) + 
  geom_histogram() + 
  theme(panel.background = element_rect(fill = "pink"))
```

<br>
Change the background fill and color (and use `size` to increase line thickness) of the *facet (panel) strips*:
```{r, fig.width = 6}
ggplot(msleep) + 
  aes(x = awake) + 
  geom_histogram() + 
  # Add faceting
  facet_wrap(vars(vore), nrow = 1) +
  theme(strip.background = element_rect(fill = "pink", 
                                        color = "blue",
                                        size = rel(4)))
```


### Quick examples of `element_line()`

Change the axis thickness (multiply default size by 2) and color, why not! 
```{r}
ggplot(msleep) + 
  aes(x = awake) + 
  geom_histogram() + 
  theme(axis.line = element_line(size = rel(2), color = "purple"))
```

<br>
Change the background grid to blue! 
```{r}
ggplot(msleep) + 
  aes(x = awake) + 
  geom_histogram() + 
  theme(panel.grid = element_line(color = "blue"))
```

<br>
Remove the *minor grid lines* (still blue to help you see the difference)
```{r}
ggplot(iris, aes(x = Sepal.Length)) + 
  geom_histogram() + 
  theme(panel.grid = element_line(color = "blue"),
        panel.grid.minor = element_blank())
```

### Quick examples of `unit()`

Legend components commonly use `unit()`, which takes two arguments: the target size, and the unit of the target size. There are many units for size, but some popular ones are `"cm"` and `"inches"`!

For example, we might want to change the size of the legend keys to make them smaller:
```{r, fig.width = 8}
ggplot(msleep) + 
  aes(x = awake, fill = vore) + 
  geom_histogram(color = "black") + 
  facet_wrap(vars(vore)) + 
  theme(legend.key.size = unit(0.2, "cm"))
```

Or larger!
```{r,  fig.width = 8}
ggplot(msleep) + 
  aes(x = awake, fill = vore) + 
  geom_histogram(color = "black") + 
  facet_wrap(vars(vore)) + 
  theme(legend.key.size = unit(1, "cm"))
```

## Modifying the theme of an existing plot


You can continue to add features, including theme changes, to a plot once it is created. For example:

```{r}
ggplot(msleep) + 
  aes(x = awake) + 
  geom_histogram() + 
  labs(title = "My awake histogram") -> awake_histogram # save plot to variable

# Add some theme changes to it, without changing the definition of `iris_histogram`
awake_histogram + 
  # moves title to top-MIDDLE of the plot instead of the default top-left
  theme(plot.title = element_text(hjust = 0.5)) 
```
