---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(dplyr)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>ceiling()</code> and <code>floor()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The `ceiling()` and `floor()` functions come with `R` and are part of the Base `R` `{base}` package.
  
We use these functions to round decimals up (`ceiling()`) or down (`floor()`) to the closest integer.

### Conceptual Usage 

```{r usage, eval=FALSE}
ceiling(numeric value to round up)
floor(numeric value to round down)

ceiling(numeric array to round up)
floor(numeric array to round down)
```

### Examples

Some examples below use the `carnivores` dataset. Learn more about this dataset with `get_help("carnivores")`.

```{r carnivores}
# Show the carnivores dataset
carnivores
```

<br>

```{r example1}
# Round the number 5.355266335 up
ceiling(5.355266335)
```

<br>

```{r example2}
# Round the number 5.355266335 down
floor(5.355266335)
```

<br>

```{r example3}
# Round the array down
some_decimals <- c(101.345, 102.298, 103.887)
floor(some_decimals)
```

<br>

```{r example4}
# Round up values in the `carnivores` column `awake` using dplyr::mutate()
carnivores %>%
  dplyr::mutate(awake_rounded_up = ceiling(awake))
```

