---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(dplyr)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>ifelse()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The `unique()` function comes with `R` and is part of the Base `R` `{base}` package.
  
We use this function to extract all _unique_ values from an array, i.e. remove all duplicates from an array.

Note that there is related function in the [`{dplyr}`](https://dplyr.tidyverse.org) package called  [`dplyr::distinct()`](https://dplyr.tidyverse.org/reference/distinct.html) which can be used to obtain distinct (aka, unique) rows from tibbles (data frames).

### Conceptual Usage 

```{r usage, eval=FALSE}
unique(array to keep unique values of)
```

### Examples

Some examples below use the `carnivores` dataset. Learn more about this dataset with `get_help("carnivores")`.

```{r carnivores}
# Show the carnivores dataset
carnivores
```

<br>

```{r example1}
# Keep only unique values in the given array (remove duplicates)
my_array <- c(1, 1, 3, 6, 6, 9, 12, 15, 15)
unique(my_array)
```

<br>

```{r example2}
# Show only the unique values in the carnivores variable `genus`
unique(carnivores$genus)
```
