---
title: "`head()` and `tail()`"
author: "{introverse} quick docs"
output: 
  html_document:
    highlight: pygments
    theme: flatly
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```


## Description

The `head()` and `tail()` functions come with R and are part of the Base R `{utils}` package.
  
We use these functions quickly to obtain the first (`head()`) or last (`tail()`) 6 rows of a given tibble (data frame), or more generally the first/last 6 entries in any array or list. With a second argument, you can change the default from 6 to a different value.



## Conceptual usage 

```{r usage, eval=FALSE}
head(tibble)

head(tibble, a different number of rows)

tail(tibble)
```

## Examples

> Run `get_help("carnivores")` for a refresher on the `carnivores` dataset.

```{r example1}
# See first 6 rows of carnivores
head(carnivores)
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->


```{r example2}
# See last 6 rows of carnivores
tail(carnivores)
```

<br> 


```{r example3, collapse=TRUE}
# See first 3 rows of carnivores
head(carnivores, 3)
```
