---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;<code>head</code> and <code>tail</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The `head()` and `tail()` functions come with R and are part of the Base R `{utils}` package.
  
We use these functions quickly to obtain the first (`head()`) or last (`tail()`) 6 rows of a given tibble (data frame), or more generally the first/last 6 entries in any array or list. With a second argument, you can change the default from 6 to a different value.



### Conceptual Usage 

```{r usage, eval=FALSE}
head(tibble)

head(tibble, a different number of rows)

tail(tibble)
```

### Examples

The examples below use the `carnivores` dataset. Learn more about this dataset with `get_help("carnivores")`.

```{r carnivores}
# Show the carnivores dataset
carnivores
```

<br>

```{r example1}
# See first 6 rows of carnivores
head(carnivores)
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->


```{r example2}
# See last 6 rows of carnivores
tail(carnivores)
```

<br> 


```{r example3}
# See first 3 rows of carnivores
head(carnivores, 3)
```
