---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;<code>as.numeric()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>{introverse}</code> quick docs</b>
</div>
<br>


### Description

The `as.numeric()` function comes with R and is part of the Base R `{base}` package.
  
We use this function to _coerce_ a value or variable to be a numeric, i.e. change the variable's type into a numeric type. 

Not all values can be changed to numeric, and this function will return `NA` if the value could not be coerced.




## Conceptual usage 

```{r usage, eval=FALSE}
as.numeric(value to coerce into numeric type)
```

### Examples


```{r example1}
# Convert the string "10" to the number 10
as.numeric("10")
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->

```{r example2, warning=TRUE}
# This example will return NA since there is no natural way to coerce the argument into a numeric
# It will also reveal a warning message telling you the coercion wasn't possible
as.numeric("you can't make strings into numbers")
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->


```{r example3}
# The logical `TRUE` turns into 1 when coerced to be a number. 
# This is consistent across all programming languages that allow coersion.
as.numeric(TRUE)
```


<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->


```{r example4}
# The logical `FALSE` turns into 0 when coerced to be a number. 
# This is consistent across all programming languages that allow coersion.
as.numeric(FALSE)
```
