---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(dplyr)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>as.factor()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The `as.factor()` function comes with `R` and is part of the Base `R` `{base}` package.
  
We use this function to _coerce_ a value or variable array to be a factor, i.e. change the variable's type into a factor type. 



### Conceptual Usage 

```{r usage, eval=FALSE}
as.factor(array to coerce into factor type)

as.factor(factor array to get factor of all its values)
```

### Examples

Some examples below use the `carnivores` dataset. Learn more about this dataset with `get_help("carnivores")`.

```{r carnivores}
# Show the carnivores dataset
carnivores
```

<br>


```{r example1}
# Convert the character array to a factor array
as.factor( c("a", "b", "b", "c", "c", "c") )
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->


```{r example2, warning=TRUE}
# Coerce the character `genus` column in carnivores into a factor column
carnivores$genus <- as.factor(carnivores$genus)
```

<br>

```{r example3, warning=TRUE}
# Coerce the character `genus` column in carnivores into a factor column using dplyr::mutate()
carnivores %>%
  dplyr::mutate(genus = as.factor(genus))
```
