---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;<code>file.exists()</code> and <code>dir.exists()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>{introverse}</code> quick docs</b>
</div>
<br>

### Description

The `file.exists()` and `dir.exists()` functions come with R and are part of the `{base}` package.
  
We use these functions to determine whether a given file (`file.exists()`) or directory aka folder (`dir.exists()`) exists. These functions return either `TRUE` or `FALSE` and are useful for double-checking if your paths are correct.


## Conceptual usage 

```{r usage, eval=FALSE}
file.exists("path/to/file/you/want/to/know/exists.txt")
dir.exists("path/to/directory/you/want/to/know/exists/")
```
