---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>summary()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The `summary()` function comes with `R` and is part of the Base `R` `{base}` package.
  
We use this function to generally summarize objects and variables. Most commonly, we summarize objects like tibbles/data frames, arrays, and statistical model output.


### Conceptual Usage 

```{r usage, eval=FALSE}
summary(object to summarize)
```

### Examples

The examples below use the `msleep` dataset. Learn more about this dataset with `get_help("msleep")`.

```{r msleep}
# Show the msleep dataset with head()
head(msleep)
```

<br>

```{r example1}
# Summarize the msleep tibble to get information about all columns
summary(msleep)
```

<br>

```{r example2}
# Summarize an array
summary(msleep$awake)
```

<br>

```{r example3}
# Summarize a linear model
fit <- lm(awake ~ bodywt, data = msleep)
summary(fit)
```