---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>paste</code> and <code>paste0</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The `paste()` and `paste0()` functions come with `R` and are part of the Base R `{base}` package.
  
We use these functions to combine multiple strings (or other values that will be coerced into characters) into one. The `paste()` function combines all provided strings separated by a space or a provided other "separator." The `paste0()` function combines all provided strings _without_ any separator.

For a slightly more convenient way to combine strings, the [`glue()`](https://glue.tidyverse.org/reference/glue.html) function from the [`{glue}`](https://glue.tidyverse.org/) package, makes it easier to combine strings with previously-defined variables. Learn more by running `get_help("glue")`.


### Conceptual Usage 

```{r usage, eval=FALSE}
# Combine with spaces
paste(these, strings, will, be, combined, into, a, single, long, string, separated, by, a, space)

# Combine with `!` separating the strings using sep
paste(use, the, sep, argument, to, change, how, these, are, separated, 
      sep = "how those strings should be separated?"

# When combining strings in an array, use `collapse` instead of `sep`
paste( c(these, strings, are, inside, an array), 
       collapse = "how those strings should be separated")

# Combine with nothing separating the strings
paste0(these, strings, will, be, combined, into, a, single, long, string, without, any, separation)

# Setting sep = "" is equivalent to paste0()
paste(equivalent, to, paste0, sep = "")
```

### Examples

<br>


```{r example1}
# Combine the following substrings into the resulting sentence
# A space is placed between each string
favorite_animal <- "tardigrades"
paste("I love", favorite_animal, "the most.")
```

<br>

```{r example2}
# Combine the following substrings into the resulting "sentence," separated by a semi-colon `;`
favorite_animal <- "tardigrades"
paste("I love", favorite_animal, "the most.", sep = ";")
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->

```{r example3}
# Combine the following substrings into the resulting "sentence," not separated at all with paste0()
# Original spaces remain, but nothing was added between "I love"/"tardigrades" and "tardigrades"/"the most."
favorite_animal <- "tardigrades"
paste0("I love", favorite_animal, "the most.")
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->

```{r example4}
# To combine all strings _in an array_ into a single string, use the argument `collapse` instead of `sep` to combine with
favorite_animals <- c("tardigrades", "rotifers", "axolotls")
paste(favorite_animals, collapse="-")
```
