---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;<code>library()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The `library()` function comes with R and is part of the Base R `{base}` package.
  
We use this function to load libraries into the R session. This function will _not install libraries for you_, which is (usually) done with the `{base}` R function `install.packages()`.

Sometimes libraries show messages when they load to tell you about some aspect of its behavior. These are generally _not errors_!

### Examples


```{r example1, eval=F}
# Load the ggplot2 library
library(ggplot2)

# You can also use quotes - it makes no difference
library("ggplot2")
```

<br>

```{r example2, error=T}
# If you try to load a library that isn't installed OR you have a typo, 
# you will see this error message:
library(definitely.a.fake.library)
```

