---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;<code>data.frame()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>{introverse}</code> quick docs</b>
</div>
<br>

### Description

The `data.frame()` function comes with R and is part of the Base R `{base}` package.
  
We use this function to create new data frames from scratch. By default, data frames print in full and show row numbers or names, which are not columns. 

Note that the `{tidyverse}` introduced a variant on the data frame, known as the **tibble**, which prints more visibly and is generally nicer to work with. We can also directly create tibbles with the `{tibble}` package functions `tibble()` or `tribble()`. 

## Conceptual usage 

```{r usage, eval=FALSE}
new_data_frame <- data.frame(
  first_column_name  = values for first column,
  second_column_name = values for second column,
...etc for more columns...)
```

### Examples


```{r example1}
# Define a data frame with two columns `a` and `b`, 
#  making sure the values for a and b columns have the same length. 
new_df <- data.frame(a = 12:17, b = c('a', 'e', 'i', 'o', 'u', 'y') )
new_df
```
