---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;Summary statistics
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>{introverse}</code> quick docs</b>
</div>
<br>

### Description

R comes with several functions (in its `{base}` and `{stats}` packages) that are used to _calculate summary statistics_ from numeric arrays. The greatest hits of R summary statistics functions include the following:

| Function | What it calculates |
|:---------|:-------------------|
| `mean()` | The arithmetic mean (average) of all values |
| `median()` | The median of all values |
| `sd()` | The standard deviation of all values |
| `sum()` | The sum of all values |
| `max()` | The maximum value among all values |
| `min()` | The minimum value among all values |

Note, there is no function to calculate the mode of a numeric array.

<br> 

By default, if the array you want to summarize contains any `NA` values, all functions above will return `NA`; R is very conservative when it encounters missing data. To _ignore `NA` values_ during calculations, provide the second argument `na.rm = TRUE` when using any of the functions.

## Conceptual usage 

```{r usage, eval=FALSE}
mean(array to summarize)

mean(array with NAs to summarize, na.rm = TRUE)
```

### Examples

> Run `get_help("msleep")` for a refresher on the `msleep` dataset.


```{r example1, collapse = TRUE}
# Summarize an array directly
mean( c(100, 125, 145, 167) )
median( c(100, 125, 145, 167) )
sd( c(100, 125, 145, 167) )
sum( c(100, 125, 145, 167) )
min( c(100, 125, 145, 167) )
max( c(100, 125, 145, 167) )
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->

```{r example2, collapse = TRUE}
# Calculate various summary statistics from the `awake` column in `msleep`
# This column does not contain NAs

mean(msleep$awake) 
median(msleep$awake)
sd(msleep$awake)
sum(msleep$awake)
min(msleep$awake)
max(msleep$awake)
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->

```{r example3, collapse = TRUE}
# Calculate various summary statistics from the `brainwt` column in `msleep`
# This column _does_ contain NAs, so we need to use the argument `na.rm=TRUE`.

mean(msleep$brainwt, na.rm = TRUE) 
median(msleep$brainwt, na.rm = TRUE)
sd(msleep$brainwt, na.rm = TRUE)
sum(msleep$brainwt, na.rm = TRUE)
min(msleep$brainwt, na.rm = TRUE)
max(msleep$brainwt, na.rm = TRUE)
```
