---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>levels()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The `levels()` function comes with `R` and is part of the Base `R` `{base}` package.
  
We use this function to find the levels, aka ordered categories, of a factor variable. Any `NA` values will be ignored and will not be considered separate levels. If the provided value is not a factor type, the result will be `NULL`.

### Conceptual Usage 

```{r usage, eval=FALSE}
levels(factor variable)
```

### Examples

The examples below use the `carnivores` dataset. Learn more about this dataset with `get_help("carnivores")`.

```{r carnivores}
# Show the carnivores dataset
carnivores
```


<br>

```{r example1}
# Define a factor for the example from the character variable `carnivores$genus`
my_factor <- as.factor(carnivores$genus)

# Show levels (ordered categories) of the factor variable:
levels(my_factor)
```

<br>

```{r example2}
# Trying to find the levels of a non-factor results in NULL
# carnivores$genus is a character column
levels(carnivores$genus)
```
