---
title: "`levels()`"
author: "{introverse} quick docs"
output: 
  html_document:
    highlight: pygments
    theme: flatly
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```


## Description

The `levels()` function comes with R and is part of the Base R `{base}` package.
  
We use this function to find the levels, aka ordered categories, of a factor variable. Any `NA` values will be ignored and not considered separate levels. If the provided value is not a factor type, the result will be `NULL`.

## Conceptual usage 

```{r usage, eval=FALSE}
levels(factor variable)
```

## Examples

> Run `get_help("msleep")` for a refresher on the `msleep` dataset.


```{r example1}
# Define a factor for the example from the character variable `msleep$vore`
my_factor <- as.factor(msleep$vore)

# Show levels of the factor variable:
levels(my_factor)
```

<br>

```{r example2}
# Trying to find the levels of a non-factor results in NULL
# msleep$vore is a character column
levels(msleep$vore)
```
