---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(dplyr)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>all()</code> and <code>any()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The `all()` and `any()` functions come with `R` and are part of the Base `R` `{base}` package.
  
We use these functions to ask if _all_ values in a given array meet some logical condition (`all()`), or if _any_ values in a given array meet some logical condition (`any()`). In other words...

+ The function `all()` asks: "Is this thing `TRUE` for _all_ of the array values?"
+ The function `any()` asks: "Is this thing `TRUE` for _any_ (at least one) of the array values?"

### Conceptual Usage 

```{r usage, eval=FALSE}
all(logical statement about an array)

any(logical statement about an array)
```

### Examples

<br>

```{r example1}
# Is it TRUE that ALL of the values in this array are greater than 5?
my_array <- c(6, 77, 8, 22)
all(my_array > 5)
```

<br>

```{r example2}
# Is it TRUE that ANY of the values in this array are greater than 5?
my_array <- c(4, 5, 7, 8, 22)
any(my_array > 5)
```

<br>

```{r example3}
# Is it TRUE that ALL of the strings in this array have 3 characters?
my_array <- c("one", "two", "three")
all( nchar(my_array) == 3 )
```

<br>

```{r example4}
# Is it TRUE that ANY of the strings in this array have 3 characters?
my_array <- c("one", "two", "three")
any( nchar(my_array) == 3 )
```

