---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(dplyr)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>str()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The `str()` function comes with `R` and is part of the Base `R` `{utils}` package.
  
We use this function to quickly view the **str**ucture of a given variable, commonly a tibble (data frame). When used on a data frame, this function will show the number rows and columns, column names, columns types, and information about the top values.

Note that there is related function in the [`{dplyr}`](https://dplyr.tidyverse.org) package called  [`dplyr::glimpse()`](https://dplyr.tidyverse.org/reference/distinct.html) which offers an _improved view_ of the structure of tibbles (data frames).

### Conceptual Usage 

```{r usage, eval=FALSE}
str(a data frame or other variable to see structure of)
```

### Examples

Some examples below use the `carnivores` dataset. Learn more about this dataset with `get_help("carnivores")`.

```{r carnivores}
# Show the carnivores dataset
carnivores
```

<br>

```{r example1}
# Show structure of the carnivores tibble
str(carnivores)
```


<br>

```{r example2}
# Show structure of the iris data frame, which comes built in with R
# Note how the output shows this is technically a data frame, not a tibble
str(iris)
```
