---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(dplyr)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>ifelse()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The `ifelse()` function comes with R and is part of the Base R `{base}` package.
  
We use this function to assign a value based on whether a logical statement is `TRUE` or `FALSE`. 

Note that there is a `{dplyr}` package version of this function ([`dplyr::if_else()`](https://dplyr.tidyverse.org/reference/if_else.html), with an underscore) that can be used in exactly the same way, but the `{dplyr}` version enforces that the resulting values must always be the same type.


### Conceptual Usage 

```{r usage, eval=FALSE}
ifelse(logical statement, 
       value to use if statement is `TRUE`, 
       value to useif statement is `FALSE`)
}
```

### Examples

Some examples below use the `carnivores` dataset. Learn more about this dataset with `get_help("carnivores")`.

```{r carnivores}
# Show the carnivores dataset
carnivores
```

<br>

```{r example1}
# Return the value 5 if it is TRUE that 10 < 4. 
# Otherwise, return the value 7.
ifelse(10 < 4, 5, 7)
```

<br>

```{r example2}
# Return the value 5 if it is TRUE that 10 < 100. 
# Otherwise, return the value 'string'. 
# Note: This will *not work* with the {dplyr} version `if_else()` since 5 is a number but 'string' is character (different types!)
ifelse(10 < 100, 5, 'string')
```

<br>

```{r example3}
# Use ifelse to create a new column `sleeps_alot` that will contain
# "yes" if `awake` <= 10, and "no" otherwise
carnivores %>%
  mutate(sleeps_alot = ifelse(awake <= 10, "yes", "no"))
```

