---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>install.packages()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The `install.packages()` function comes with `R` and is part of the Base `R` `{base}` package.
  
We use this function to install packages, aka libraries, onto your computer. Libraries only need to be installed one time. To _load_ an installed library, use the function `library()`. 

If you need to install a package from a non-default database (i.e. not CRAN), this function may or may not work.



### Conceptual Usage 

```{r usage, eval=FALSE}
install.packages("name of package to install in quotes")

install.packages(c("array", "of", "package", "names", "to", "install"))
```

### Examples


```{r example1, eval = FALSE}
# Install the viridis package
install.packages("viridis")
```
