---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>nrow()</code> and <code>ncol()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The `nrow()` and `ncol()` functions come with `R` and are part of the Base `R` `{base}` package.
  
We use these functions to find the number of rows (`nrow()`) or columns (`ncol()`) of a variable, usually a tibble (data frame).

### Conceptual Usage 

```{r usage, eval=FALSE}
nrow(tibble)
ncol(tibble)
```

### Examples

The examples below use the `carnivores` dataset. Learn more about this dataset with `get_help("carnivores")`.

```{r carnivores}
# Show the carnivores dataset
carnivores
```

<br>

```{r example1}
# Find the number of rows in carnivores
nrow(carnivores)
```

<br>

```{r example2}
# Find the number of columns in carnivores
ncol(carnivores)
```
