---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>length()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The `length()` function comes with `R` and is part of the Base `R` `{base}` package.
  
We use this function to determine the length of a value or variable, most commonly of array or list. Note that this function WILL NOT tell you the length of a string (i.e., how many characters it contains); instead, use `nchar()` to ask how long a string is.

### Conceptual Usage 

```{r usage, eval=FALSE}
length(array to find the length of)
```

### Examples


```{r example1}
# Find the length of an array directly
length( c(44, 55, 66) )
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->


```{r example2}
# Find the length of an array which has been defined as a variable
my_array <- c(44, 55, 66)
length(my_array)
```

<br>


```{r example3}
## CAUTION! length() does not work properly to find the number of characters in a string
length("R thinks this string is a single value in an array of length 1")
```
