---
title: "{introverse} quick docs"
output: 
  html_document:
    theme: readable
    syntax: tango
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

<br>

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;The <code>carnivores</code> dataset
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

One of the `{introverse}` example datasets is called **`carnivores`**. This dataset is a tibble with 9 rows and 4 columns and contains information about mammalian species. 

This dataset is a _subset_ of the `msleep` dataset, which comes from the `{ggplot2}` package. To learn more about `msleep`, use `get_help("msleep")`.

You can directly explore the `carinvores` tibble in the Console, or you can interactively explore its contents below.

### Explore `carnivores`

Columns in `carnivores` include the following:

+ `name`: The common name of the mammal
+ `genus`: The Genus classification of the mammal species
+ `awake`: The amount of time spent awake, in hours
+ `brainwt`: The brain weight in kg

Values in the data that are `NA` indicate that the data are _missing and/or unknown._

<br>
```{r echo=FALSE}
DT::datatable(
  dplyr::mutate(carnivores, 
         brainwt = ifelse(is.na(brainwt), "NA", brainwt)),
  rownames=FALSE,
  fillContainer = FALSE)
```


