---
title: "The `msleep` dataset"
author: "`{introverse}` quick docs"
output: 
  html_document:
    theme: flatly
    syntax: pygments
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(DT)
clean_nas <- function(df){
  df %>%
    dplyr::mutate(
      dplyr::across(dplyr::everything(), ~tidyr::replace_na(.x, "NA"))
    )  
}
```

<br>

## Description

One of the `{introverse}` example datasets is called **`msleep`**. This dataset is a tibble with 83 rows and 11 columns and contains information about mammalian species. This dataset comes from the `{ggplot2}` package.

You can directly explore the `msleep` tibble in the Console, or you can interactively explore its contents below.


### Explore `msleep`

Columns in `msleep` include the following:

+ `name`: The common name of the mammal
+ `genus`: The Genus classification of the mammal species
+ `vore`: Whether the species is **carni**vore, **hervi**vore, **omni**vore, or **insecti**vore
+ `order`: The Order classification of the mammal species
+ `conservation`: The conservation status of the mammal species
  + `lc`: "**l**east **c**oncern"
  + `nt`: "**n**ot **t**hreatened"
  + `en`: "**en**dangered"
  + `cd`: "**c**onservation **d**ependent
  + `domesticated`: domesticated
+ `sleep_total`: The total amount of sleep per day, in hours
+ `sleep_rem`: The amount of REM sleep per day, in hours
+ `sleep_cycle`: The length of a sleep cycle, in hours
+ `awake`: The amount of time spent awake, in hours
+ `brainwt`: The brain weight in kg
+ `bodywt`: The body weight in kg

Values in the data that are _NA_ indicate that the data are _missing and/or unknown._

<br>

```{r echo=FALSE}
DT::datatable(clean_nas(msleep), 
              rownames=FALSE)
```

