---
title: "{introverse} quick docs"
output: 
  html_document:
    theme: readable
    syntax: tango
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(DT)
clean_nas <- function(df){
  df %>%
    dplyr::mutate(
      dplyr::across(dplyr::everything(), ~tidyr::replace_na(.x, "NA"))
    )  
}
```

<br>
```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;The <code>msleep</code> dataset
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

One of the `{introverse}` example datasets is called **`msleep`**. This dataset is a tibble with 83 rows and 11 columns and contains information about mammalian species. This dataset comes from the `{ggplot2}` package.

You can directly explore the `msleep` tibble in the Console, or you can interactively explore its contents below.


### Explore `msleep`

Columns in `msleep` include the following:

+ `name`: The common name of the mammal
+ `genus`: The Genus classification of the mammal species
+ `vore`: Whether the species is **carni**vore, **hervi**vore, **omni**vore, or **insecti**vore
+ `order`: The Order classification of the mammal species
+ `conservation`: The conservation status of the mammal species
  + `lc`: **l**east **c**oncern
  + `nt`: **n**ot **t**hreatened
  + `en`: **en**dangered
  + `cd`: **c**onservation **d**ependent
  + `domesticated`: domesticated
+ `sleep_total`: The total amount of sleep per day, in hours
+ `sleep_rem`: The amount of REM sleep per day, in hours
+ `sleep_cycle`: The length of a sleep cycle, in hours
+ `awake`: The amount of time spent awake, in hours
+ `brainwt`: The brain weight in kg
+ `bodywt`: The body weight in kg

Values in the data that are `NA` indicate that the data are _missing and/or unknown._

<br>

```{r echo=FALSE}
DT::datatable(clean_nas(msleep), 
              rownames=FALSE,
              fillContainer = FALSE)
```

