---
title: "The `carnivores` dataset"
author: "`{introverse}` quick docs"
output: 
  html_document:
    theme: flatly
    syntax: pygments
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(DT)
clean_nas <- function(df){
  df %>%
    dplyr::mutate(
      dplyr::across(dplyr::everything(), ~tidyr::replace_na(.x, "NA"))
    )  
}
```

<br>


## Description

One of the `{introverse}` example datasets is called **`carnivores`**. This dataset is a tibble with 9 rows and 4 columns and contains information about mammalian species. 

This dataset is a _subset_ of the `msleep` dataset, which comes from the `{ggplot2}` package. To learn more about `msleep`, use `get_help("msleep")`.

You can directly explore the `carinvores` tibble in the Console, or you can interactively explore its contents below.

### Explore `carnivores`

Columns in `carnivores` include the following:

+ `name`: The common name of the mammal
+ `genus`: The Genus classification of the mammal species
+ `awake`: The amount of time spent awake, in hours
+ `brainwt`: The brain weight in kg

Values in the data that are _NA_ indicate that the data are _missing and/or unknown._

<br>
```{r echo=FALSE}
DT::datatable(clean_nas(carnivores), 
              rownames=FALSE,
              fillContainer = FALSE)
```


