---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(dplyr)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;<code>dplyr::n()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`n()`](https://dplyr.tidyverse.org/reference/n.html) function is part of the [`{dplyr}`](https://dplyr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

The function `n()` is a *helper function* that returns a number representing the total number of rows in a given grouping. It is most useful if you want to add an 'index column' in a tibble (data frame).


<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{dplyr}` library, or always use the function with `dplyr::n()` notation.

```{r description, eval=FALSE}
# Load the library
library(dplyr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
dplyr::n()
```


### Conceptual Usage 

```{r usage, eval=FALSE}
tibble %>% 
  mutate(column_ranging_1_to_n = 1:n())
```

### Examples

The examples below use the `carnivores` dataset. Learn more about this dataset with `get_help("carnivores")`.

```{r carnivores}
# Show the carnivores dataset
carnivores
```

<br>

```{r example1}
# Make a new column `one_to_n` that counts the rows 1-n in carnivores
carnivores %>%
  mutate(one_to_n = 1:n())
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->


```{r example2}
# Make a new column `one_to_n` that counts the rows in carnivores, but
# where the numbering restarts for each `genus` grouping
carnivores %>%
  group_by(genus) %>%
  mutate(one_to_n = 1:n()) %>%
  # arranging by genus to make the output easier to see
  arrange(genus)
```
