---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(dplyr)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>dplyr::case_when()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`case_when()`](https://dplyr.tidyverse.org/reference/case_when.html) function is part of the [`{dplyr}`](https://dplyr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

We use this function to *recode* a variable to different values. The `case_when()` function can be used when there are more than two conditions for recoding, in which case `if_else()` would be appropriate. We often use `case_when()` along with `mutate()` to create new data frame columns.

Be patient with the `case_when()` syntax: It is tricky!


<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{dplyr}` library, or always use the function with `dplyr::case_when()` notation.

```{r description, eval=FALSE}
# Load the library
library(dplyr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
dplyr::case_when()
```


### Conceptual Usage 

```{r usage, eval=FALSE}
case_when(a logical statement ~ return this value if this statement is TRUE,
          a different logical statement ~ return this value if this statement is TRUE,
          yet another logical statement ~ return this value if this statement is TRUE,
          (...as many as you need...)
          # optional
          TRUE ~ value to return if condition was not ever met)

```

### Examples

The example below uses the `carnivores` dataset. Learn more about this dataset with `get_help("carnivores")`.

```{r carnivores}
# Show the carnivores dataset
carnivores
```

<br>
```{r example1}
# Use case_when to help create a new column called `how_awake` in `carnivores` that will be:
## "alert" if awake >= 16
## "medium" if awake < 16 and awake > 10
## "sleepy" if awake <= 10
carnivores %>%
  mutate(how_awake = case_when(
          # Spacing out code like this makes it easier to see, write, and use.
          awake >= 16 ~ 'alert',
          awake < 16 & awake > 10 ~ 'medium',
          awake <= 10 ~ 'sleepy')
    )
```
