---
title: "`dplyr::distinct()`"
author: "{introverse} quick docs"
output: 
  html_document:
    highlight: pygments
    theme: flatly
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(dplyr)
```


## Description

The [`distinct()`](https://dplyr.tidyverse.org/reference/distinct.html) function is part of the [`{dplyr}`](https://dplyr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

It is used to remove duplicate rows from tibbles (data frames). Only distinct rows are retained. No arguments are needed when using this fucntion.

<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{dplyr}` library, or always use the function with `dplyr::distinct()` notation.

```{r description, eval=FALSE}
# Load the library
library(dplyr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
dplyr::distinct()
```


## Conceptual usage 

```{r usage, eval=FALSE}
tibble %>% 
  distinct()
```

## Examples

> Run `get_help("carnivores")` for a refresher on the `carnivores` dataset.

```{r example1}
# Keep only distinct rows in carnivores
## All rows are already distinct, so all rows are kept
carnivores %>% 
  distinct()
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->


```{r example2}
# This example creates a tibble with non-unique rows for more demonstration:

# Make a tibble with 3 rows using `tibble::tribble()` 
new_tibble <- tibble::tribble(
  ~col1, ~col2,
  1, 5, 
  1, 5, 
  2, 6, 
)

# Show new_tibble
new_tibble

# Keep only unique rows from new_tibble
new_tibble %>%
  distinct()
```
