---
title: "`dplyr::summarize()`"
author: "`{introverse}` quick docs"
output: 
  html_document:
    highlight: pygments
    theme: flatly
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(dplyr)
```


## Description

The [`summarize()`](https://dplyr.tidyverse.org/reference/summarize.html) function is part of the [`{dplyr}`](https://dplyr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

It is used to summarize tibbles (data frame), such that the resulting tibble has new column(s) and is smaller, retaining only what was needed for summarizing. This function is often used with `group_by()` to quickly summarize across different groupings at once.


<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{dplyr}` library, or always use the function with `dplyr::summarize()` notation.

```{r description, eval=FALSE}
# Load the library
library(dplyr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
dplyr::summarize()
```


## Conceptual usage 

```{r usage, eval=FALSE}
tibble %>% 
  summarize(name_of_new_column = summary statistic calculation)

tibble %>% 
  group_by(grouping column) %>% 
  summarize(name_of_new_column = summary statistic calculation)
```

## Examples

> Run `get_help("msleep")` for a refresher on the `msleep` dataset.

```{r example1}
# Calculate the mean time spent awake
msleep %>% 
  summarize(mean_awake = mean(awake))
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->


```{r example2}
# Calculate the mean time spent awake for each vore group
msleep %>% 
  group_by(vore) %>%
  summarize(mean_awake = mean(awake))
```

<br> 

```{r example3}
# Calculate the mean time spent awake for each combination of vore and conservation groups
msleep %>% 
  group_by(vore, conservation) %>%
  summarize(mean_awake = mean(awake))
```

<br> 


```{r example4}
# Calculate the median brainwt
## Caution! Because the brainwt column contains NAs, 
## we need to use the _argument_ `na.rm = TRUE` with `median()`
msleep %>% 
  summarize(median_brainwt = median(brainwt, na.rm = TRUE))
```
