---
title: "`dplyr::select()`"
author: "`{introverse}` quick docs"
output: 
  html_document:
    highlight: pygments
    theme: flatly
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(dplyr)
```


## Description

The [`select()`](https://dplyr.tidyverse.org/reference/select.html) function is part of the [`{dplyr}`](https://dplyr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

It is used to keep, remove, and/or reorder columns in tibbles (data frames).


<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{dplyr}` library, or always use the function with `dplyr::select()` notation.

```{r description, eval=FALSE}
# Load the library
library(dplyr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
dplyr::select()
```


## Conceptual usage 

```{r usage, eval=FALSE}
tibble %>%
  select(single column to keep)

tibble %>%
  select(-single column to remove)

tibble %>% 
  select(columns, you, want, to, keep)

tibble %>% 
  select(-columns, -you, -want, -to, -remove)
```

## Examples

> Run `get_help("carnivores")` for a refresher on the `carnivores` dataset.

```{r example1}
# Keep only the columns name and genus from carnivores
carnivores %>% 
  select(name, genus)
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->


```{r example2}
# Remove the column genus from carnivores
carnivores %>% 
  select(-genus)
```

<br> 


```{r example3}
# Remove the columns genus and awake from carnivores
carnivores %>% 
  select(-genus, -awake)
```

<br> 

```{r example4}
# Reorder the columns in order: genus, name, brainwt, awake
carnivores %>%
  select(genus, name, brainwt, awake)
```
