---
title: "`dplyr::case_when()`"
author: "{introverse} quick docs"
output: 
  html_document:
    highlight: pygments
    theme: flatly
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(dplyr)
```


## Description

The [`case_when()`](https://dplyr.tidyverse.org/reference/case_when.html) function is part of the [`{dplyr}`](https://dplyr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

We use the function `case_when()` to *recode* a variable to different values. The `case_when()` function can be used when there are more than two conditions for recoding, in which case `if_else()` would be appropriate. We often use `case_when()` along with `mutate()` to create new data frame columns.

Be patient with the `case_when()` syntax: It is tricky!


<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{dplyr}` library, or always use the function with `dplyr::case_when()` notation.

```{r description, eval=FALSE}
# Load the library
library(dplyr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
dplyr::case_when()
```


## Conceptual usage 

```{r usage, eval=FALSE}
case_when(a logical statement ~ return this value if this statement is TRUE,
          a different logical statement ~ return this value if this statement is TRUE,
          yet another logical statement ~ return this value if this statement is TRUE,
          (...as many as you need...)
          # optional
          TRUE ~ value to return if condition was not ever met)

```

## Examples

> Run `get_help("carnivores")` for a refresher on the `carnivores` dataset.

```{r example1}
# Use case_when to help create a new column called `how_awake` in `carnivores` that will be:
## "alert" if awake >= 16
## "medium" if awake < 16 and awake > 10
## "sleepy" if awake <= 10
carnivores %>%
  mutate(how_awake = case_when(
          # Spacing out code like this makes it easier to see, write, and use.
          awake >= 16 ~ 'alert',
          awake < 16 & awake > 10 ~ 'medium',
          awake <= 10 ~ 'sleepy')
    )
```
