---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(dplyr)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;<code>dplyr::fitler()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>{introverse}</code> quick docs</b>
</div>
<br>

### Description

The [`filter()`](https://dplyr.tidyverse.org/reference/filter.html) function is part of the [`{dplyr}`](https://dplyr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

We use this function to **subset rows** from tibbles (data frames). Rows are kept if the given logical statement (code that gives `TRUE` or `FALSE`) is `TRUE`. Use this function if you only want to keep certain rows, aka certain observations."


<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{dplyr}` library, or always use the function with `dplyr::filter()` notation.

```{r description, eval=FALSE}
# Load the library
library(dplyr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
dplyr::filter()
```


## Conceptual usage 

```{r usage, eval=FALSE}
tibble %>% 
  filter(logical statement)

tibble %>% 
  filter(logical statement,
         another logical statement,
         and more logical statements that you want to all be TRUE)

```

### Examples

> Run `get_help("carnivores")` for a refresher on the `carnivores` dataset.

```{r example1}
# Subset carnivores to keep only rows that are the genus 'Panthera' 
carnivores %>% 
  filter(genus ==  "Panthera")
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->


```{r example2}
# Subset carnivores to keep only rows of carnivores awake for more than 13 hours
carnivores %>%
  filter(awake > 13)
```

<br> 


```{r example3, collapse=TRUE}
# Subset carnivores to keep only rows where the genus is 'Panthera' and they are awake for _less than_ 13 hours
carnivores %>%
  filter(genus == "Panthera", awake < 13)

# Or, use & as "and"
carnivores %>%
  filter(genus == "Panthera" & awake < 13)
```

<br>

```{r example4}
# Subset rows to keep only where genus is 'Panthera' **or** brainwt < 0.05
# Or, use | (vertical bar on \ key) as "or"
carnivores %>%
  filter(genus == "Panthera" | brainwt < 0.05)
```

<br> 

```{r example5, collapse=TRUE}
# Subset carnivores to keep only rows where the genus is _either_ 'Panthera' or 'Canis'
# In other words, it should be TRUE that the genus is %in% the array c("Panthera", "Canis"). 
# Remember: We do _not_ use `==` when asking if a value is in an array.
carnivores %>%
  filter(genus %in% c("Panthera", "Canis"))

# Or, use | (vertical bar on \ key) as "or"
carnivores %>%
  filter(genus == "Panthera" | genus == "Canis")
```


