---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(dplyr)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;<code>dplyr::tally()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`tally()`](https://dplyr.tidyverse.org/reference/tally.html) function is part of the [`{dplyr}`](https://dplyr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

The function `tally()` creates a new column (named `n` by default) in a tibble (data frame) that tallies the number of rows (observations) in specified group.  If there is no grouping, `tally()` will give the total number of rows. This is similar to using base `nrow()` except, like many `{dplyr}` verbs, `tally()` returns a tibble. 

Note that the `{dplyr}` function `count()` provides a shortcut for `group_by() %>% tally() %>% ungroup()`"

<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{dplyr}` library, or always use the function with `dplyr::tally()` notation.

```{r description, eval=FALSE}
# Load the library
library(dplyr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
dplyr::tally()
```


### Conceptual Usage 

```{r usage, eval=FALSE}
# Summarize into tibble indicating total rows
tibble %>% 
  tally()

tibble %>% 
  tally(name = count) # specify a non-default output column name

tibble %>% 
  group_by(grouping column) %>% 
  tally() %>% 
  ungroup()
```

### Examples

The examples below use the `msleep` dataset. Learn more about this dataset with `get_help("msleep")`.

```{r msleep}
# Show the msleep dataset with head()
head(msleep)
```

<br>
```{r example1}
# Count the total number of rows in msleep
msleep %>%
  tally()
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->


```{r example2}
# Count the total number of rows in msleep and specify column name
msleep %>%
  tally(name = 'total')
```

<br> 


```{r example3}
# Count total number of rows for each vore. 
# Note that this can also be done more simply with `dplyr::count()`
msleep %>%
  group_by(vore) %>%
  tally()
```

<br> 

```{r example4}
# Subset carnivores to keep only rows where the genus is _either_ 'Panthera' or 'Canis'
# In other words, it should be TRUE that the genus is %in% the array c("Panthera", "Canis"). 
# Remember: We do _not_ use `==` when asking if a value is in an array.
carnivores %>%
  tally(genus %in% c("Panthera", "Canis"))
```
