---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(dplyr)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;<code>dplyr::select()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>{introverse}</code> quick docs</b>
</div>
<br>

### Description

The [`select()`](https://dplyr.tidyverse.org/reference/select.html) function is part of the [`{dplyr}`](https://dplyr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

It is used to keep, remove, and/or reorder columns in tibbles (data frames).


<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{dplyr}` library, or always use the function with `dplyr::select()` notation.

```{r description, eval=FALSE}
# Load the library
library(dplyr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
dplyr::select()
```


## Conceptual usage 

```{r usage, eval=FALSE}
tibble %>%
  select(single column to keep)

tibble %>%
  select(-single column to remove)

tibble %>% 
  select(columns, you, want, to, keep)

tibble %>% 
  select(-columns, -you, -want, -to, -remove)
```

### Examples

> Run `get_help("carnivores")` for a refresher on the `carnivores` dataset.

```{r example1}
# Keep only the columns name and genus from carnivores
carnivores %>% 
  select(name, genus)
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->


```{r example2}
# Remove the column genus from carnivores
carnivores %>% 
  select(-genus)
```

<br> 


```{r example3}
# Remove the columns genus and awake from carnivores
carnivores %>% 
  select(-genus, -awake)
```

<br> 

```{r example4}
# Reorder the columns in order: genus, name, brainwt, awake
carnivores %>%
  select(genus, name, brainwt, awake)
```
