---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(dplyr)
```

```{css hide-title, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```


<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;<code>dplyr::mutate()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>{introverse}</code> quick docs</b>
</div>
<br>

### Description

The [`mutate()`](https://dplyr.tidyverse.org/reference/mutate.html) function is part of the [`{dplyr}`](https://dplyr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

We use this function to create new columns in tibbles (data frames). Since `mutate()` will also overwrite existing columns, you can also use it to modify existing columns in place.


<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{dplyr}` library, or always use the function with `dplyr::mutate()` notation.

```{r description, eval=FALSE}
# Load the library
library(dplyr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
dplyr::mutate()
```


## Conceptual usage 

```{r usage, eval=FALSE}
tibble %>% 
  mutate(name_of_new_column = what to put in new column)

tibble %>% 
  mutate(name_of_first_new_column = what to put in this new column,
         name_of_second_new_column = what to put in this new column,          
         etc. as needed)

```

### Examples

> Run `get_help("carnivores")` for a refresher on the `carnivores` dataset.

```{r example1}
# Create a new column called `my_new_column` that literally contains the value `5` at all rows
carnivores %>% 
  mutate(my_new_column =  5)
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->


```{r example2}
# Create a new column `brainwt_g` that contains brain weight in g, instead of kg.
carnivores %>%
  mutate(brainwt_g = brainwt * 1000) # multiply by 1000 to convert kg --> g
```

<br> 


```{r example3}
# Create a new column `little_name` that contains the `name` values in lower case.
# We'll use the function stringr::str_to_lower() to convert strings to lower case
carnivores %>%
  mutate(little_name = stringr::str_to_lower(name))
```

<br> 

```{r example4}
# Use mutate() to change the column `genus` into a factor instead of character
carnivores %>%
  mutate(genus = as.factor(genus))
```
