---
title: "`dplyr::filter()`"
author: "{introverse} quick docs"
output: 
  html_document:
    highlight: pygments
    theme: flatly
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      exercise.lines = 6)
library(dplyr)
```

<br>

### Description

The [`filter()`](https://dplyr.tidyverse.org/reference/filter.html) function is part of the [`{dplyr}`](https://dplyr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

It is used to **subset rows** from tibbles (data frames). Rows are kept if the given logical statement (code that gives `TRUE` or `FALSE`) is `TRUE`. Use this function if you only want to keep certain rows, aka certain observations."


<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{dplyr}` library, or always use the function with `dplyr::filter()` notation.

```{r description, eval=FALSE}
# Load the library
library(dplyr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
dplyr::filter()
```




### Conceptual usage 

```{r usage, eval=FALSE}
tibble %>% 
  filter(logical statement)

tibble %>% 
  filter(logical statement,
         another logical statement,
         and more logical statements that you want to all be TRUE)

```

### Examples

> Run `get_help("carnivores")` for a refresher on the `carnivores` dataset.

```{r q1, exercise = TRUE}
# Subset carnivores to keep only rows that are the genus 'Panthera' 
carnivores %>% 
  filter(genus ==  "Panthera")
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->


```{r q2, exercise = TRUE}
# Subset carnivores to keep only rows of carnivores awake for more than 13 hours
carnivores %>%
  filter(awake > 13)
```

<br> 


```{r q3, exercise = TRUE}
# Subset carnivores to keep only rows where the genus is 'Panthera' and they are awake for _less than_ 13 hours
carnivores %>%
  filter(genus == "Panthera", awake < 13)
```

<br> 

```{r q4, exercise = TRUE}
# Subset carnivores to keep only rows where the genus is _either_ 'Panthera' or 'Canis'
# In other words, it should be TRUE that the genus is %in% the array c("Panthera", "Canis"). 
# Remember: We do _not_ use `==` when asking if a value is in an array.
carnivores %>%
  filter(genus %in% c("Panthera", "Canis"))
```
