---
title: "`dplyr::bind_cols()`"
author: "`{introverse}` quick docs"
output: 
  html_document:
    theme: flatly
    syntax: pygments
---

<!-- Welcome to the template! This comment is here to help you get started. The HTML comments will help you fill it out. Please make sure to do the following:
- make the YAML `title` the topic name. For example, the `filter()` docs are titled `dplyr::filter()` (with backticks since it's code!)
- name the final Rmd file "packagethetopiccomesfrom_topic.Rmd". For example, the `filter()` file is named `dplyr_filter.Rmd`
- If you are considering documenting MULTIPLE TOPICS in a single help Rmd, STOP and CONSULT FIRST with Spielman!!
-->

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(dplyr)
# Load any INDIVIDUAL libraries here.
# NEVER load the full tidyverse - In other words `library(tidyverse)` => ABSOLUTELY NOT!!!! Load only what you need, _one at a time_.

# Do NOT read in any datasets (carnivores/msleep)!!! It's taken care of, as long as you develop as follows instead of knitting (really, no knitting!) - 
# 1) devtools::load_all()
# 2) get_help("topic youre currently developing")
```


## Description

The [`bind_cols()`](https://dplyr.tidyverse.org/reference/bind.html) function is part of the [`{dplyr}`](https://dplyr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

This function is used to *bind* the columns of two tibbles (data frames) or arrays into one tibble (data frame). This requires every column to be of equal length.

<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{dplyr}` library, or always use the function with `dplyr::bind_cols()` notation.

```{r, eval=FALSE}
# Load the library
library(dplyr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
dplyr::bind_cols()
```



## Conceptual usage 

```{r, eval=FALSE}
bind_cols(first tibble, second tibble)

bind_cols(first array, second array)

# Remember that both tibbles and both arrays are of equal length in order to bind
```


## Examples

> Need a refresher on the `msleep` dataset? Run `get_help("msleep")` for more information.

<!-- Add a <br> if there is NO DATASET REFRESHER!! Otherwise the refresher takes up enough space. -->

```{r example1}
# This example first shows the creation of two arrays, then the use of `bind_cols()` to bind them.
array1 <- 1:3
array2 <- 4:6
bind_cols(column1 = array1, column2 = array2)

```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. There should be a <br> tab between each example -->


```{r example2}
# This example is similar to the first, but now using tibbles (data frames) derived from the data set `msleep`.

msleep %>% 
  select(name:vore) -> tibble1 # Using a colon between column names includes columns between them.

msleep %>% 
  select(order, sleep_total) -> tibble1 # Using a comma selects individual columns to *select*


# Unfamiliar with pipes (%>%)? - run get_help("%>%")

bind_cols(tibble1, tibble2)
```

<br> <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->

<!-- If needed, add more example chunks below WITH NAMES (example3, example4, etc.), making sure to have a <br> tag in between each example for nicer spacing. If you don't need more examples, don't add any! -->

