---
title: "`dplyr::rename()`"
author: "{introverse} quick docs"
output: 
  html_document:
    highlight: pygments
    theme: flatly
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(dplyr)
```


## Description

The [`rename()`](https://dplyr.tidyverse.org/reference/rename.html) function is part of the [`{dplyr}`](https://dplyr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

We use `rename()` to quickly rename columns in tibbles (data frames). No quotes are needed.


<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{dplyr}` library, or always use the function with `dplyr::rename()` notation.

```{r description, eval=FALSE}
# Load the library
library(dplyr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
dplyr::rename()
```


## Conceptual usage 

```{r usage, eval=FALSE}
tibble %>% 
  rename(new_name_of_column = current_name_of_column)
```

## Examples

> Run `get_help("carnivores")` for a refresher on the `carnivores` dataset.

```{r example1}
# Rename the `name` column in carnivores to `common_name`
carnivores %>% 
  rename(common_name = name)
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->


```{r example2}
# Rename the `name` column in carnivores to `common_name`, 
# and rename the `brainwt` column to `brain_weight_kg`
carnivores %>% 
  rename(common_name = name, 
         brain_weight_kg = brainwt)
```
