---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(dplyr)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;<code>dplyr::if_else()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`if_else()`](https://dplyr.tidyverse.org/reference/if_else.html) function is part of the [`{dplyr}`](https://dplyr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

We use this function to assign a value based on whether a logical statement is `TRUE` or `FALSE`. We often use `case_when()` along with `mutate()` to create new data frame columns. The provided values must be of the same type! 

Note that there is a `{base}` R version of this function `dplyr::ifelse()` (without an underscore!) that can be used in exactly the same way, but the `{base}` version does not enforce that the result is always the same type.



<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{dplyr}` library, or always use the function with `dplyr::if_else()` notation.

```{r description, eval=FALSE}
# Load the library
library(dplyr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
dplyr::if_else()
```


## Conceptual usage 

```{r usage, eval=FALSE}
if_else(logical statement, 
        value if statement is `TRUE`, 
        value if statement is `FALSE`)

tibble %>%
  mutate(new_column = if_else(logical statement, 
                              value for new_column if statement is `TRUE`, 
                              value for new_column if statement is `FALSE`)
  )

```

### Examples

> Run `get_help("carnivores")` for a refresher on the `carnivores` dataset.

```{r example1}
# Return the value 5 if it is TRUE that 10 < 4. Otherwise, return the value 7.
if_else(10 < 4, 5, 7)
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->


```{r example2, error=TRUE}
# Return the value 5 if it is TRUE that 10 < 100. Otherwise, return the value 'string'. 
## But wait! This code will result in an ERROR since 5 is a number but 'string' is character (different types!)",

if_else(10 < 100, 5, 'string')
```

<br> 


```{r example3}
# Use if_else to help create a new column called `how_awake` in `carnivores` that will be:
## "alert" if awake >= 15 and "sleepy" otherwise.
carnivores %>%
  mutate(how_awake = if_else(awake >=15, "alert", "sleepy"))
```