---
title: "`{introverse}` quick docs"
output: 
  html_document:
    theme: readable
    syntax: tango
---

<!-- Welcome to the template! This comment is here to help you get started. The HTML comments will help you fill it out. Please make sure to do the following:
- make the YAML `title` the topic name. For example, the `filter()` docs are titled `dplyr::filter()` (with backticks since it's code!)
- name the final Rmd file "packagethetopiccomesfrom_topic.Rmd". For example, the `filter()` file is named `dplyr_filter.Rmd`
- If you are considering documenting MULTIPLE TOPICS in a single help Rmd, STOP and CONSULT FIRST with Spielman!!
-->

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(dplyr)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>dplyr::bind_rows()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

## Description

The [`bind_rows()`](https://dplyr.tidyverse.org/reference/bind.html) function is part of the [`{dplyr}`](https://dplyr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

This function is used to *bind* the rows of two tibbles (data frames) or arrays into one tibble (data frame). This requires that both tibbles have the same columns.

<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{dplyr}` library, or always use the function with `dplyr::bind_cols()` notation.

```{r, eval=FALSE}
# Load the library
library(dplyr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
dplyr::bind_rows()
```



### Conceptual Usage 

```{r, eval=FALSE}
# Remember that both tibbles must have the same columns to be able to bind them

bind_rows(first tibble, second tibble)
```


## Examples

The examples below use the `carnivores` dataset. Learn more about this dataset with `get_help("carnivores")`.

```{r msleep}
# Show the carnivores dataset
carnivores
```

<br>
<!-- Add a <br> if there is NO DATASET REFRESHER!! Otherwise the refresher takes up enough space. -->

```{r example1}
# Use bind_rows() to combine two smaller tibbles

# First, create two smaller tibbles _from carnivores_ to use for demonstration:

# use dplyr::slice to extract the first row only
carnivores %>% 
  slice(1) -> tibble1 

# use dplyr::slice to extract the second row only
carnivores %>% 
  slice(2) -> tibble2

# Bind rows together:
bind_rows(tibble1, tibble2)
```