---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(dplyr)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;<code>dplyr::count()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>{introverse}</code> quick docs</b>
</div>
<br>

### Description

The [`count()`](https://dplyr.tidyverse.org/reference/count.html) function is part of the [`{dplyr}`](https://dplyr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

We use this function to create a new column (named `n` by default) in a tibble (data frame) that counts the number of rows in specified group. It is a convenient shortcut for: `group_by(column_name) %>% tally() %>% ungroup()`"


<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{dplyr}` library, or always use the function with `dplyr::count()` notation.

```{r description, eval=FALSE}
# Load the library
library(dplyr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
dplyr::count()
```


## Conceptual usage 

```{r usage, eval=FALSE}
tibble %>% 
  count(column to count)

tibble %>% 
  count(first column to count,
        second column to count in all combinations with the first,
        etc. as needed)
```

### Examples

> Run `get_help("msleep")` for a refresher on the `msleep` dataset.

```{r example1}
# Count the number of each vore category in `msleep`
msleep %>%
  count(vore)
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->


```{r example2}
# Count the number of each combination of `vore` and `order` categories in `msleep`
msleep %>%
  count(vore, order)
```
