---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(forcats)
library(dplyr)
library(ggplot2)
library(tidyr)
msleep %>%
  drop_na(vore) %>%                  # Remove any NAs from the vore column with tidyr::drop_na()
  mutate(vore = as.factor(vore))-> msleep_fctvore
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>forcats::fct_relevel()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`fct_relevel()`](https://forcats.tidyverse.org/reference/fct_relevel.html) function is part of the [`{forcats}`](https://forcats.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

We use the `fct_relevel()` function to change the _order_ of categories (levels) in a factor variable based on your custom ordering, using a variety of flexible syntaxes. 

Changing the order of factor levels is commonly performed to change axis order of a factor variable when using plotting with the `{ggplot2}` library.

<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{forcats}` library, or always use the function with `forcats::fct_relevel()` notation.

```{r description, eval=FALSE}
# Load the library
library(forcats)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
forcats::fct_relevel()
```


### Conceptual Usage 

> See examples for what sorts of additional arguments one can provide.

```{r usage, eval=FALSE}
fct_relevel(factor variable to relevel, 
            ..additional arguments for how to relevel..)
```

### Examples

The examples below use a modified version of the `msleep` dataset called `msleep_fctvore`. Learn more about this dataset with `get_help("msleep")`.

In this modified dataset, the `vore` column has been _coerced_ into a factor type (instead of character), and all `NA` values have been removed from that column. (Notice below, the `vore` column is annotated `<fct>` since it's a factor).

```{r msleep, collapse=TRUE}
# Show the modified msleep dataset, msleep_fctvore, with head()
head(msleep_fctvore)

# Show the default order (alphabetical!) of the factor vore column.
# All examples below will use the function fct_relevel() to manipulate this order:
levels(msleep_fctvore$vore)
```
<br>

```{r example1}
# Re-write the `vore` column, releveled to make "omni" come first
msleep_fctvore %>%
  mutate(vore = fct_relevel(vore, "omni")) -> msleep_fctvore_ex1

# Show new levels to confirm "omni" is now first
levels(msleep_fctvore_ex1$vore)
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->


```{r example2}
# Re-write the `vore` column, releveled as: "omni", "herbi", "carni", "insecti"
# You can freely list out the additional arguments inside `fct_relevel()`
msleep_fctvore %>%
  mutate(vore = fct_relevel(vore, 
                            "omni", "herbi", "carni", "insecti")) -> msleep_fctvore_ex2

# Show new levels to confirm new order
levels(msleep_fctvore_ex2$vore)
```


<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->


```{r example3}
# Re-write the `vore` column, releveled to make "insect" second, aka:
# "insecti" should come _after_ the first (1) level. (Note: there is no alternative `before` argument.)
# This fct_relevel approach is helpful when you want to move one level without listing all the levels out
msleep_fctvore %>%
  mutate(vore = fct_relevel(vore, "insecti", after = 1)) -> msleep_fctvore_ex3

# Show new levels to confirm new order
levels(msleep_fctvore_ex3$vore)
```


<br>


```{r example4}
# Re-write the `vore` column, releveled to make "insect" LAST, using `Inf` ("infinity")
# "insecti" should come _after_ the first level
# This fct_relevel approach is helpful when you want to move one level without listing all the levels out

msleep_fctvore %>%
  mutate(vore = fct_relevel(vore, "insecti", after = Inf)) -> msleep_fctvore_ex4

# Show new levels to confirm new order
levels(msleep_fctvore_ex4$vore)
```

<br>

```{r example5}
# Without re-writing the column, reorder the levels for _plotting purposes only_
# Provide fct_relevel(VARIABLE) to ggplot2::aes() to order in your plot
# This affects the x-axis labeling, so it is best practice to clean up with `labs()`
ggplot(msleep_fctvore) +
                            # Order we want `vore` to appear in
  aes(x = fct_relevel(vore, "insecti", "carni", "omni", "herbi"), 
      y = awake) + 
  geom_boxplot() + 
  labs(x = "vore")
```