---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(forcats)
library(dplyr)
library(ggplot2)
library(tidyr)
msleep %>%
  drop_na(vore) %>%                  # Remove any NAs from the vore column with tidyr::drop_na()
  mutate(vore = as.factor(vore))-> msleep_fctvore
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>forcats::fct_reorder()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`fct_reorder()`](https://forcats.tidyverse.org/reference/fct_reorder.html) function is part of the [`{forcats}`](https://forcats.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

We use the `fct_reorder()` function to quickly _reorder_ of categories (levels) in a factor variable _based on the order of values in another variable._ 

Changing the order of factor levels is commonly performed to change axis order of a factor variable when using plotting with the `{ggplot2}` library.

<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{forcats}` library, or always use the function with `forcats::fct_reorder()` notation.

```{r description, eval=FALSE}
# Load the library
library(forcats)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
forcats::fct_reorder()
```


### Conceptual Usage 

```{r usage, eval=FALSE}
fct_reorder(factor variable to change order of, other variable to order by)

fct_reorder(factor variable to change order of, 
            other variable to order by, 
            # Commonly used optional arguments:
            .fun = how to transform the other variable for re-ordering, 
            .desc = TRUE or FALSE) # Ordered in descending order, by default
```


### Examples

The examples below use a modified version of the `msleep` dataset called `msleep_fctvore`. Learn more about this dataset with `get_help("msleep")`.

In this modified dataset, the `vore` column has been _coerced_ into a factor type (instead of character), and all `NA` values have been removed from that column. (Notice below, the `vore` column is annotated `<fct>` since it's a factor).

```{r msleep, collapse=TRUE}
# Show the modified msleep dataset, msleep_fctvore, with head()
head(msleep_fctvore)

# Examples will reorder based on mean(awake), so this shows what to expect:
# Calculate and show the values of mean(awake) per vore. 
msleep_fctvore %>%
  group_by(vore) %>%
  summarize(mean_awake = mean(awake)) %>%
  arrange(mean_awake)
```
<br>


```{r example1}
# reorder the order of `vore` levels according to _mean_ values of _awake_
msleep_fctvore %>%
  mutate(vore = fct_reorder(vore, 
                            awake, 
                            # This argument tells forcats to order vore by MEAN awake
                            .fun = mean)) -> msleep_fctvore_ex1

# Show new levels to confirm they are updated
levels(msleep_fctvore_ex1$vore)
```

<br>

```{r example2}
# reorder the order of `vore` levels according to _mean_ values of _awake_, but in descending order
msleep_fctvore %>%
  mutate(vore = fct_reorder(vore, 
                            awake, 
                            .fun = mean, # order vore by MEAN awake
                            .desc = TRUE)) -> msleep_fctvore_ex2

# Show new levels to confirm they are updated
levels(msleep_fctvore_ex2$vore)
```

<br>

```{r example3}
# reorder the order of `vore` levels according to _mean_ values of _awake_, 
# Here, we first summarize the data so we need only to reorder by `mean_awake`, without needing a `.fun` argument
msleep_fctvore %>%
  group_by(vore) %>%
  summarize(mean_awake = mean(awake)) %>%
  mutate(vore = fct_reorder(vore, mean_awake)) -> msleep_fctvore_ex3

# Show new levels to confirm they are updated
levels(msleep_fctvore_ex3$vore)
```


<br>

```{r example4}
# Without re-writing the column, change the levels for _plotting purposes only_
# Provide fct_reorder(VARIABLE, OTHER VARIABLE, ...) to ggplot2::aes() to order in your plot
# This affects the x-axis labeling, so it is best practice to clean up with `labs()`

# For this plot, we'll reorder `vore` based on MINIMUM AWAKE values
# Recall: The R function to calculate minimums is `min()`
ggplot(msleep_fctvore) +
  aes(x = fct_reorder(vore, awake, .fun = min), 
      y = awake) + 
  geom_boxplot() + 
  labs(x = "vore")
```


<br>

```{r example5}
# Without re-writing the column, change the levels for _plotting purposes only_
# Provide fct_reorder(VARIABLE, OTHER VARIABLE, ...) to ggplot2::aes() to order in your plot
# This affects the x-axis labeling, so it is best practice to clean up with `labs()`

# For this plot, we'll reorder `vore` based on its COUNT, using `dplyr::count()`
msleep_fctvore %>%
  count(vore) %>%
  ggplot() + 
  aes(x = fct_reorder(vore, n), 
      y = n) + 
  geom_col() + 
  labs(x = "vore")
```