---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(forcats)
library(dplyr)
library(ggplot2)
library(tidyr)
msleep %>%
  drop_na(order) %>%                  # Remove any NAs from the vore column with tidyr::drop_na()
  mutate(order = as.factor(order)) -> msleep_fctorder
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>forcats::fct_lump_*()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`fct_lump_*()`](https://forcats.tidyverse.org/reference/fct_lump.html) functions are part of the [`{forcats}`](https://forcats.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

`{forcats}` contains several related functions for _lumping_ together categories (levels) in a factor variable to condense into fewer overall categories. Some of the most useful lumping functions include the following:

| `{forcats}` function | How it lumps |
|:----------|:-------------|
|  `fct_lump_min()` | Lumps together all levels which occur fewer than a specified **minimum number** of times into a single level |
| `fct_lump_prop()` | Lumps together all levels which occur fewer than a specified **proportion** of times into a single level |
| `fct_lump_n()` | Lumps together all levels _except_ for the specified **n** most frequenct levels |

These functions will combine all lumped levels into a single new level, "Other," whose name can be customized. These functions are very useful for working with a factor variable that has a very high number of categories, which may be "overly-busy" to view all at once, especially in a plot. 

Changing the order of factor levels is commonly performed to change axis order of a factor variable when using plotting with the `{ggplot2}` library.

<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{forcats}` library, or always use the function with `forcats::fct_lump_n()` notation.

```{r description, eval=FALSE}
# Load the library
library(forcats)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
forcats::fct_lump_min()
forcats::fct_lump_prop()
forcats::fct_lump_n()
```


### Conceptual Usage 

```{r usage, eval=FALSE}
fct_lump_min(factor_variable, 
             the minimum number of times, 
             other_level = "Name you want to use if not 'Other'")

fct_lump_prop(factor_variable, 
              the minimum proportion of times,
              other_level = "Name you want to use if not 'Other'")


fct_lump_n(factor_variable, 
           the number of high-frequency levels NOT to lump,
           other_level = "Name you want to use if not 'Other'")
```

### Examples

The examples below use a modified version of the `msleep` dataset called `msleep_fctorder`. Learn more about this dataset with `get_help("msleep")`.

In this modified dataset, the `order` column has been _coerced_ into a factor type (instead of character), and all `NA` values have been removed from that column. (Notice below, the `order` column is annotated `<fct>` since it's a factor).

```{r msleep, collapse=TRUE}
# Show the modified msleep dataset, msleep_fctorder, with head()
head(msleep_fctorder)

# To help guide you through examples, use the base R function table() count the number of each order 
# This shows us which categories are more/less frequent
table(msleep_fctorder$order)
```
<br>



```{r example1}
# Lump together all levels occurring fewer than 5 times
msleep_fctorder %>%
  mutate(order = fct_lump_min(order, 5)) -> msleep_fctorder_ex1

# Show new levels to confirm they are updated: Notice the new category "Other" contains all the lumped
levels(msleep_fctorder_ex1$order)
```

<br>

```{r example2}
# Lump together all levels occurring fewer than 5 times, name name new category "Fewer than 5x"
msleep_fctorder %>%
  mutate(order = fct_lump_min(order, 5, other_level = "Fewer than 5x")) -> msleep_fctorder_ex2

# Show new levels to confirm they are updated: Notice the new category "Fewer than 5x" contains all the lumped
levels(msleep_fctorder_ex2$order)
```

<br>

```{r example3}
# Lump together all levels occurring in fewer than 10% of rows
msleep_fctorder %>%
  mutate(order = fct_lump_prop(order, 0.1)) -> msleep_fctorder_ex3

# Show new levels to confirm they are updated
levels(msleep_fctorder_ex3$order)
```


<br>

```{r example4}
# Lump together all levels EXCEPT the 3 most common levels
msleep_fctorder %>%
  mutate(order = fct_lump_n(order, 3)) -> msleep_fctorder_ex4

# Show new levels to confirm they are updated
levels(msleep_fctorder_ex4$order)
```

<br>

```{r example5}
# Without re-writing the column, change the levels for _plotting purposes only_
# Provide, for example, fct_lump_n(VARIABLE, the n value) to ggplot2::aes() to order in your plot
# This affects the x-axis labeling, so it is best practice to clean up with `labs()`
ggplot(msleep_fctorder) +
      # Lump everything EXCEPT the top 4 levels, and use a custom name for the lumped level
  aes(x = fct_lump_n(order, 4, 
                     other_level = "Additional taxonomic orders"), 
      y = awake) + 
  geom_boxplot() + 
  labs(x = "order")
```