---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(forcats)
library(dplyr)
library(ggplot2)
library(tidyr)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;<code>forcats::fct_infreq()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>{introverse}</code> quick docs</b>
</div>
<br>

### Description

The [`fct_infreq()`](https://forcats.tidyverse.org/reference/fct_infreq.html) function is part of the [`{forcats}`](https://forcats.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

We use the `fct_infreq()` function to quickly _reorder_ of categories (levels) in a factor variable based on how many times each level appears. The most commonly appearing level will be first, and the least-commonly appear level will be last. 

Changing the order of factor levels is commonly performed to change axis order of a factor variable when using plotting with the `{ggplot2}` library.

<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{forcats}` library, or always use the function with `forcats::fct_infreq()` notation.

```{r description, eval=FALSE}
# Load the library
library(forcats)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
forcats::fct_infreq()
```


## Conceptual usage 

```{r usage, eval=FALSE}
fct_infreq(factor variable to change order of)
```

### Examples

> Run `get_help("msleep")` for a refresher on the `msleep` dataset.

```{r factorize}
# For examples, we must first create a new version of `msleep` that contains a 
#  factor variable. We will convert `vore` to a factor with base R function`as.factor()`.
msleep %>%
  drop_na(vore) %>%                  # Remove any NAs from the vore column with tidyr::drop_na()
  mutate(vore = as.factor(vore)) %>% # Make the vore column factor type with dplyr::mutate() 
  #Make the `vore` column appear first for easier viewing with dplyr::select() and tidyselect::everything()
  select(vore, everything()) -> msleep_fctvore

# Show the example dataset, which now shows `vore` as <fct> (factor!) type
msleep_fctvore

# Show the default order (alphabetical!) of the factor vore column.
# Examples below will use fct_infreq() to manipulate this order:
levels(msleep_fctvore$vore)

# First, let's see how often each `vore` appears in the dataset using `dplyr::count()`, and
# `dplyr::arrange() to see ordered high-to-low
# Results from examples should "match" this order from high-to-low
msleep_fctvore %>%
  count(vore) %>%
  arrange(desc(n))
```

<br><br>

```{r example1}
# reorder the order of `vore` levels according to frequency
msleep_fctvore %>%
  mutate(vore = fct_infreq(vore)) -> msleep_fctvore_ex1

# Show new levels to confirm they are updated
levels(msleep_fctvore_ex1$vore)
```

<br>

```{r example2}
# Without re-writing the column, change the levels for _plotting purposes only_
# Provide fct_infreq(VARIABLE) to ggplot2::aes() to order in your plot
# This affects the x-axis labeling, so it is best practice to clean up with `labs()`
ggplot(msleep_fctvore) +
  aes(x = fct_infreq(vore), 
      y = awake) + 
  geom_boxplot() + 
  labs(x = "vore")
```