---
title: "`forcats::fct_rev()`"
author: "`{introverse}` quick docs"
output: 
  html_document:
    highlight: pygments
    theme: flatly
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(forcats)
library(dplyr)
library(ggplot2)
library(tidyr)
```


## Description

The [`fct_rev()`](https://forcats.tidyverse.org/reference/fct_rev.html) function is part of the [`{forcats}`](https://forcats.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

We use the `fct_rev()` function to quickly _reverse the order_ of categories (levels) in a factor variable. 

Changing the order of factor levels is commonly performed to change axis order of a factor variable when using plotting with the `{ggplot2}` library.

<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{forcats}` library, or always use the function with `forcats::fct_rev()` notation.

```{r description, eval=FALSE}
# Load the library
library(forcats)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
forcats::fct_rev()
```


## Conceptual usage 

> See examples for what sorts of additional arguments one can provide.

```{r usage, eval=FALSE}
fct_rev(factor variable to reverse order of)
```

## Examples

> Run `get_help("msleep")` for a refresher on the `msleep` dataset.

```{r factorize}
# For examples, we must first create a new version of `msleep` that contains a 
#  factor variable. We will convert `vore` to a factor with base R function`as.factor()`.
msleep %>%
  drop_na(vore) %>%                  # Remove any NAs from the vore column with tidyr::drop_na()
  mutate(vore = as.factor(vore)) %>% # Make the vore column factor type with dplyr::mutate() 
  #Make the `vore` column appear first for easier viewing with dplyr::select() and tidyselect::everything()
  select(vore, everything()) -> msleep_fctvore

# Show the example dataset, which now shows `vore` as <fct> (factor!) type
msleep_fctvore

# Show the default order (alphabetical!) of the factor vore column.
# Examples below will use fct_rev() to manipulate this order:
levels(msleep_fctvore$vore)
```

<br><br>

```{r example1}
# Reverse the order of `vore` levels
msleep_fctvore %>%
  mutate(vore = fct_rev(vore)) -> msleep_fctvore_ex1

# Show new levels to confirm they are reversed
levels(msleep_fctvore_ex1$vore)
```

<br>

```{r example2}
# Without re-writing the column, reverse the levels for _plotting purposes only_
# Provide fct_rev(VARIABLE) to ggplot2::aes() to order in your plot
# This affects the x-axis labeling, so it is best practice to clean up with `labs()`
ggplot(msleep_fctvore) +
  aes(x = fct_rev(vore), 
      y = awake) + 
  geom_boxplot() + 
  labs(x = "vore")
```