---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = TRUE)
library(ggplot2)
msleep %>% tidyr::drop_na(sleep_rem, awake) -> msleep
set.seed(1)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;<code>ggplot2::aes()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`aes()`](https://ggplot2.tidyverse.org/reference/aes.html) function is part of the [`{ggplot2}`](https://ggplot2.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

We use the `aes()` function within `{ggplot2}` plots to **establish aesthetic _mappings_** between visual features in the plot and _columns_ in the tibble (data frame) being plotted. The `aes()` function should be used _only_ to map columns to aesthetics. Other stylings can be directly provided to geoms, without enclosing them in `aes()`.


<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{ggplot2}` library, or always use the function with `ggplot2::aes()` notation.

```{r description, eval=FALSE}
# Load the library
library(ggplot2)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
ggplot2::aes()
```

### Conceptual Usage

```{r usage, eval=FALSE}
# Provide aes() to initial ggplot call so that all geoms inherit
ggplot(tibble to plot, 
       mapping = aes(aspect of plot = column name in tibble, etc.))

# It is optional whether you use the 'mapping' argument keyword with `ggplot()`
# This is equivalent to above:
ggplot(tibble to plot, 
       aes(aspect of plot = column name in tibble, etc.))

# Provide aes on its own so that all geoms inherit
ggplot(tibble to plot) +
  aes(aspect of plot = column name in tibble, etc.)

# Provide aes to a geom directly
ggplot(tibble to plot) +
  geom_whatever(aes(aspect of plot = column name in tibble, etc.))
```

### Examples

> Run `get_help("msleep")` for a refresher on the `msleep` dataset.

```{r example1}
# Make a boxplot with `vore` on the x-axis and `awake` on the y-axis
# This example provides aes() inside ggplot()
# Note, to remove NAs, you must manipulate the data first
ggplot(msleep, 
       aes(x = vore, y = awake)) + 
  geom_boxplot()
```

<br>

```{r example2}
# Make a boxplot with `vore` on the x-axis and `awake` on the y-axis, with aes() on its own
# This examples provides aes() separately, outside of ggplot()
# Note, to remove NAs, you must manipulate the data first
ggplot(msleep) + 
  aes(x = vore, y = awake) + 
  geom_boxplot()
```

<br>

```{r example3}
# Make a boxplot with `vore` on the x-axis and `awake` on the y-axis, with aes() on its own
# Also _fill_ the boxplots based on the values in the `vore` column
# Note, to remove NAs, you must manipulate the data first
ggplot(msleep) + 
  aes(x = vore, y = awake, fill = vore) + 
  geom_boxplot()
```

<br>


```{r example4, error=TRUE}
## CAUTION! This example code results in an error because `vore` is a column but it was not placed into `aes()`
# Note, to remove NAs, you must manipulate the data first
ggplot(msleep) + 
  aes(x = vore, y = awake) + 
  geom_boxplot(fill = vore)
```

<br>

```{r example4.5}
## CAUTION! This example code results in _an inappropriate legend_ and _wrong colors_ because a non-column is provided to `aes()`
# Note, to remove NAs, you must manipulate the data first
ggplot(msleep) + 
  aes(x = vore, y = awake) + 
  geom_boxplot(aes(fill = "blue")) # "blue" is not a column. It's just a fill specification without mapping, and the fill isn't even blue!
```

<br>

```{r example5}
# When using multiple geoms, all will inherit mappings from an initial aes() call
# Note, to remove NAs, you must manipulate the data first
ggplot(msleep) + 
  aes(x = vore, 
      y = awake, 
      fill = vore,
      color = vore) +
  geom_jitter()  +  # also inherits x, y, color, and fill mappings
  geom_boxplot()    # inherits x, y, color, and fill mappings
```

<br>

```{r example6}
# When using multiple geoms, you can decide which geoms get which aesthetic:
# Note, to remove NAs, you must manipulate the data first
ggplot(msleep) + 
  aes(x = vore, y = awake) +       # All geoms inherit the x/y mapping
  geom_jitter(aes(color = vore), pch = 21) + # ONLY the jitter geom gets the color mapping. Using pch=21, we see points are NOT filled, only colored
  geom_boxplot(aes(fill = vore), alpha = 0.5)   # ONLY the boxplot gets the fill mapping. Its color remains default black. Add alpha for transparency
```

```{r example7}
# Another example with multiple geoms: All geoms inherit the aes() mappings. 
# Each vore has its own `geom_smooth()` trendline _because of the color mapping_
# Note, to remove NAs, you must manipulate the data first
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake, color = vore) +       
  geom_point() +  
  geom_smooth(method = "lm", se=FALSE) # Linear trendline without confidence interval, for clarity
```

<br>

```{r example8}
# Another example with multiple geoms: Choose which geoms use which aesthetic
# Note, to remove NAs, you must manipulate the data first
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake) +       
  geom_point(aes(color = vore)) +  
  geom_smooth(method = "lm", se=FALSE) # Does NOT separate trendlines by color: Single trendline for all vores
```

<br>

```{r example9}
# Another example with multiple geoms: Choose which geoms use which aesthetic
# Note, to remove NAs, you must manipulate the data first
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake) +       
  geom_point() +   # Does NOT show points by color
  geom_smooth(aes(color = vore), # Separate trendline for all vores
              method = "lm", 
              se = FALSE) 
```



```{r returnmsleep, echo=F, include=F}
msleep <- introverse::msleep
```