---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = TRUE)
library(ggplot2)
msleep %>% tidyr::drop_na(sleep_rem, awake) -> msleep
set.seed(1)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>ggplot2::xlim()</code> and <code>ggplot2::ylim()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

All functions documented here are part of the [`{ggplot2}`](https://ggplot2.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

The `{ggplot2}` package [provides several functions](https://ggplot2.tidyverse.org/reference/lims.html) for customizing the limits, aka ranges, of the axes in your `{ggplot2}` plots, as described below. `{ggplot2}` will usually pick a reasonable axis range, so using these functions is not necessary unless you need to change the ranges for a specific data visualization goal.

If you also want to customize the specific tickmark (aka "break") locations on axes, either...

+ [See this tutorial on the `{introverse}` website](https://spielmanlab.github.io/introverse/articles/axes.html)
+ Or, run this command in `R` Console: `vignette("axes", package = "introverse")`


```{r description, eval=FALSE}
# Load the library
library(ggplot2)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation, for example with xlim()
ggplot2::xlim() 
```


### Conceptual Usage

```{r conceptual, eval=F}
# Set x-axis limits with `xlim()`
xlim(low x value, high x value)

# Set x-axis limits with `ylim()`
ylim(low y value, high y value)

```

### Examples

The examples below use the `msleep` dataset. Learn more about this dataset with `get_help("msleep")`.

```{r msleep}
# Show the msleep dataset with head()
head(msleep)
```


<br>
```{r example0}
# This example does NOT set limits, but uses ggplot2's default chosen ranges
# Compare this default to the subsequent examples to see how xlim() and ylim() behave
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake) + 
  geom_point() 
```

<br>
```{r example1}
# Set the x-axis limits to range from 0-24, for example
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake) + 
  geom_point() +
  xlim(0, 24)
```

<br>

```{r example2}
# Set the y-axis limits to range from 0-40, for example
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake) + 
  geom_point() +
  ylim(0, 40)
```

<br>


```{r example4}
# Set both the x- and y-axis limits to range from 0-24
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake) + 
  geom_point() +
  xlim(0, 24) +
  ylim(0, 24)
```

<br>

```{r example5, warning=TRUE}
# CAUTION!! If you limit the ranges too much, data is excluded! See the "Warning" from this code.
# Always make sure you are not excluding data that you don't specifically intend to exclude
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake) + 
  geom_point() +
  # Narrow axis ranges will exclude data:
  xlim(1, 3) +
  ylim(2, 7)
```
