---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = TRUE, 
                      fig.width = 8,
                      fig.height = 5)
library(ggplot2)
library(dplyr)
msleep %>% tidyr::drop_na(sleep_rem, awake) -> msleep
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;Faceting in <code>ggplot2</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

All functions documented here are part of the [`{ggplot2}`](https://ggplot2.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

The `{ggplot2}` package provides two useful functions for creating _paneled_, i.e. _faceted_ plots:

+ [`facet_wrap()`](https://ggplot2.tidyverse.org/reference/facet_wrap.html) can be used to facet across one variable. You can specify how many rows and columns should be laid out in the paneling.
+ [`facet_grid()`](https://ggplot2.tidyverse.org/reference/facet_grid.html) is most commonly used to facet across two variables to make a _matrix_ of panels, but can also be used to facet across one variable. 


<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use these functions, you need to either first load the `{ggplot2}` library, or always use the function with `ggplot2::()` notation.

```{r description, eval=FALSE}
# Load the library
library(ggplot2)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation, for example with facet_wrap()
ggplot2::facet_wrap()
```


### Conceptual Usage


```{r conceptual, eval=FALSE}
# Most basic usage:

# We do NOT use aes() to specify variables for faceting. Instead we use vars():
facet_wrap(vars(tibble variable to facet across))

facet_grid(rows = vars(tibble variable to facet across rows),
           cols = vars(tibble variable to facet across columns))

# Or, facet_grid with one variable:
facet_grid(rows = vars(tibble variable to facet across rows))
facet_grid(cols = vars(tibble variable to facet across columns))

```

### Examples

The examples below use a _modified_ version of the `msleep` dataset, where `NA` values have been removed from columns `vore` and `conservation` using `tidyr::drop_na()`. Learn more about the `msleep` with `get_help("msleep")`.

```{r msleep}
# Modify msleep for examples:
msleep %>%
  tidyr::drop_na(vore, conservation) -> msleep_cleaned # Plots are made with `msleep_cleaned`

# Show dataset:
msleep_cleaned
```

<br>

```{r example1}
# Use facet_wrap with default layout to panel scatterplot by vore: Each panel represents a vore.
ggplot(msleep_cleaned) + 
  aes(x = sleep_rem, 
      y = awake, 
      color = vore) + 
  geom_point() + 
  facet_wrap(vars(vore))
```

<br>

```{r example2}
# Use argument `nrow` to specify how many panel rows facet_wrap should plot.
ggplot(msleep_cleaned) + 
  aes(x = sleep_rem, 
      y = awake, 
      color = vore) + 
  geom_point() + 
  facet_wrap(vars(vore), nrow = 1) # All panels will be laid out in ONE ROW
```

<br>


```{r example3}
# Or, use argument `ncol` to specify how many panel columns facet_wrap should plot.
ggplot(msleep_cleaned) + 
  aes(x = sleep_rem, 
      y = awake, 
      color = vore) + 
  geom_point() + 
  facet_wrap(vars(vore), ncol = 1) # All panels will be laid out in ONE COLUMN
```


<br>



```{r example4}
# Use the `scales` argument to allow x-axis limits to vary acrcoss panels depending on panel data ranges
ggplot(msleep_cleaned) + 
  aes(x = sleep_rem, 
      y = awake, 
      color = vore) + 
  geom_point() + 
  facet_wrap(vars(vore), scales = "free_x") # Free the x scales to vary based on data - now panels have different x-axis ranges
```



<br>

```{r example5}
# Use the `scales` argument to allow y-axis limits to vary acrcoss panels depending on panel data ranges
ggplot(msleep_cleaned) + 
  aes(x = sleep_rem, 
      y = awake, 
      color = vore) + 
  geom_point() + 
  facet_wrap(vars(vore), scales = "free_y") # Free the y scales to vary based on data - now panels have different y-axis ranges
```

<br>

```{r example6}
# Use the `scales` argument to allow BOTH x- and y-axis limits to vary acrcoss panels depending on panel data ranges
ggplot(msleep_cleaned) + 
  aes(x = sleep_rem, 
      y = awake, 
      color = vore) + 
  geom_point() + 
  facet_wrap(vars(vore), scales = "free") # Free the BOTH x and y scales to vary based on data - now panels have different axis ranges
```

<br>


```{r example7}
# Use facet_grid to make a paneled plot of conservation and vore
# Note that some panels are empty, meaning there are no data points wiht that combination of those vore/conservation categories in msleep.
ggplot(msleep_cleaned) + 
  aes(x = sleep_rem, 
      y = awake, 
      color = vore) + 
  geom_point() + 
  # Here, vore panels are in rows, and conservation panels are in columns
  facet_grid(rows = vars(vore),
             cols = vars(conservation))
```

<br>

```{r example8}
# Use facet_grid to make a paneled plot of conservation and vore, reversed
# Note that some panels are empty, meaning there are no data points wiht that combination of those vore/conservation categories in msleep.
ggplot(msleep_cleaned) + 
  aes(x = sleep_rem, 
      y = awake, 
      color = vore) + 
  geom_point() + 
  # Here, conservation panels are in rows, and vore panels are in columns
  facet_grid(rows = vars(conservation),
             cols = vars(vore))
```


<br>

```{r example9}
# Free both axis scales with `scales = "free". This can also be done as `scales = "free_x"` or `scales = "free_y"` to free only one axis
ggplot(msleep_cleaned) + 
  aes(x = sleep_rem, 
      y = awake, 
      color = vore) + 
  geom_point() + 
  # Here, conservation panels are in rows, and vore panels are in columns
  facet_grid(rows = vars(conservation),
             cols = vars(vore), 
             scales = "free") # allow both X and Y panel axes ranges to vary based on panel's data
```


<br> 

```{r example10}
# Switch y facet label locations in a faceted plot (works with either facet_wrap or facet_grid)
ggplot(msleep_cleaned) + 
  aes(x = sleep_rem, 
      y = awake, 
      color = vore) + 
  geom_point() + 
  # Here, conservation panels are in rows, and vore panels are in columns
  facet_grid(rows = vars(conservation),
             cols = vars(vore), 
             switch = "y") # This argument switches y facet labels to be on the LEFT
```
<br>


```{r example11}
# Switch x facet label locations in a faceted plot (works with either facet_wrap or facet_grid)
ggplot(msleep_cleaned) + 
  aes(x = sleep_rem, 
      y = awake, 
      color = vore) + 
  geom_point() + 
  # Here, conservation panels are in rows, and vore panels are in columns
  facet_grid(rows = vars(conservation),
             cols = vars(vore), 
             switch = "x") # This argument switches x facet labels to be on the BOTTOM
```

<br>


```{r example12}
# Switch x and y label locations in a faceted plot (works with either facet_wrap or facet_grid)
ggplot(msleep_cleaned) + 
  aes(x = sleep_rem, 
      y = awake, 
      color = vore) + 
  geom_point() + 
  # Here, conservation panels are in rows, and vore panels are in columns
  facet_grid(rows = vars(conservation),
             cols = vars(vore), 
             switch = "both") # This argument switches BOTH x and y facet labels to be on the bottom and left, respectively
```
