---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = TRUE)
library(ggplot2)
msleep %>% tidyr::drop_na(sleep_rem, awake) -> msleep
set.seed(1)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;<code>ggplot2::ggplot()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`ggplot()`](https://ggplot2.tidyverse.org/reference/ggplot.html) function is part of the [`{ggplot2}`](https://ggplot2.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

We use the `ggplot()` function to establish a baseline `{ggplot2}` plot. Note that the function _does not end in the number 2!_ Most commonly, we provide this function with the `data` argument of a tibble (data frame) whose data should be plotted. Optionally, we can also provide _aesthetic mappings_ with `aes()`, which will automatically be applied to all geoms.


<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{ggplot2}` library, or always use the function with `ggplot2::ggplot()` notation.

```{r description, eval=FALSE}
# Load the library
library(ggplot2)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
ggplot2::ggplot()
```

### Conceptual Usage

```{r usage, eval=FALSE}

# You can use the 'data' and 'mapping' argument keywords if you chose, but they are not required
ggplot(data = tibble or data frame to plot, 
       mapping = aes(optional aethestic mappings))

# `ggplot()` assumes the first argument is the data, and the second argument (if present) is aethetics
ggplot(tibble or data frame to plot,
       aes(optional aethestic mappings))

# The aes() argument is not required at all:
ggplot(tibble or data frame to plot)
```

### Examples

> Run `get_help("msleep")` for a refresher on the `msleep` dataset.


```{r example0}
# Calling ggplot by itself establishes a blank canvas using the default theme:
ggplot()
```

<br>

```{r example01}
# Calling ggplot with data and mappings but without a geom sets up axes:
ggplot(msleep, 
       aes(x = vore, y = awake)) 
```

<br>

```{r example1}
# Make a boxplot with `vore` on the x-axis and `awake` on the y-axis
# This example provides aes() inside ggplot()
# Note, to remove NAs, you must manipulate the data first
ggplot(msleep, 
       aes(x = vore, y = awake)) + 
  geom_boxplot()
```

<br>

```{r example2}
# Make a boxplot with `vore` on the x-axis and `awake` on the y-axis, with aes() on its own
# This examples provides aes() separately, outside of ggplot()
# Note, to remove NAs, you must manipulate the data first
ggplot(msleep) + 
  aes(x = vore, y = awake) + 
  geom_boxplot()
```
