---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = TRUE)
library(ggplot2)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;<code>ggplot2::geom_boxplot()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`geom_boxplot()`](https://ggplot2.tidyverse.org/reference/geom_boxplot.html) function is part of the [`{ggplot2}`](https://ggplot2.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

The `geom_boxplot()` function is used within `{ggplot2}` plots to create a boxplot, which visualizes summary statistics of a _numeric continuous_ variable collapsed into its summary statistics: minimum, median, maximum, 1st and 3rd quartiles, with any outliers shown beyond minimum and maximum. Boxplots are often used to _compare_ different distributions across a categorical variable.

#### Required aesthetics:

+ `x`: A categorical variable to visualize boxplots across
+ `y`: A numeric variable to visualize distributions of
  + Note that these can be switched for a horizontal boxplot

#### Commonly used optional aesthetics:

+ `color` (`colour`): The outline color for the boxplot
+ `fill`: The fill for the boxplot
+ `size`: The width of the boxplot outlines
+ `outlier.size`: The size of outlier points
+ `outlier.color` (`outlier.colour`): The color of outlier points
+ `outlier.shape`: The shape of outlier points
+ `outlier.fill`: The fill of outlier points, when the shape is specified to allow a fill


<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{ggplot2}` library, or always use the function with `ggplot2::geom_boxplot()` notation.

```{r description, eval=FALSE}
# Load the library
library(ggplot2)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
ggplot2::geom_boxplot()
```



### Examples

> Run `get_help("msleep")` for a refresher on the `msleep` dataset.

```{r example1}
# Distribution of awake across vore categories
# Note to remove the NAs, you must manipulate the data first
ggplot(msleep) + 
  aes(x = vore, y = awake) + 
  geom_boxplot()
```

<br>

```{r example2}
# Distribution of awake across vore categories with all boxplots filled the same
ggplot(msleep) + 
  aes(x = vore, y = awake) + 
  geom_boxplot(fill = "orange")
```

<br>


```{r example2.5}
# Distribution of awake across vore categories, filled by vore
ggplot(msleep) + 
  aes(x = vore, y = awake, fill = vore) + 
  geom_boxplot()
```

<br>

```{r example3}
# Distribution of awake across vore categories, filled by vore, with customized outliers
ggplot(msleep) + 
  aes(x = vore, y = awake, fill = vore) + 
  geom_boxplot(outlier.size = 0.5, outlier.color = "red")
```


<br>

```{r example4}
# To visualize a single boxplot for all values, use x = "" instead of providing a variable
ggplot(msleep) + 
  aes(x = "", y = awake) + 
  geom_boxplot()
```
