---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = TRUE)
library(ggplot2)
msleep %>% tidyr::drop_na(sleep_rem, awake) -> msleep
set.seed(1)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;Built-in complete themes in <code>{ggplot2}</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

All functions documented here are part of the [`{ggplot2}`](https://ggplot2.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

The `{ggplot2}` package [provides several built-in complete themes](https://ggplot2.tidyverse.org/reference/ggtheme.html) for styling the overall look of your plot. The default theme used is `theme_gray()` (aka `theme_grey()`). You can specify a built-in theme in one of two ways:

+ Use the function `ggplot2::theme_set()` to change _all_ subsequent plots made during the given R session.
+ Add a given complete theme function to your plot directly to change only that plot's complete theme

Available complete themes include the following:

+ `theme_gray()` aka `theme_grey()` _(The default theme.)_
+ `theme_bw()`
+ `theme_linedraw()`
+ `theme_light()`
+ `theme_dark()`
+ `theme_minimal()`
+ `theme_classic()`
+ `theme_void()`


<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use these function, you need to either first load the `{ggplot2}` library, or always use the function with `ggplot2::` notation.

```{r description, eval=FALSE}
# Load the library
library(ggplot2)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation, for example with theme_bw()
ggplot2::theme_bw() 
```


### Examples

> Run `get_help("msleep")` for a refresher on the `msleep` dataset.

```{r example1, eval=F}
# Set the default theme of all subsequent plots to theme_bw(), for example
# Don't forget parentheses!
theme_set(theme_bw())
```

<br>

```{r example2}
# Example of a plot using theme_gray (the default)
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake) + 
  geom_point() +
  # Optional, since this is the default
  theme_gray()
```

<br>

```{r example3}
# Example of a plot using theme_bw()
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake) + 
  geom_point() +
  theme_bw()
```

<br>


```{r example4}
# Example of a plot using theme_linedraw()
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake) + 
  geom_point() +
  theme_linedraw()
```

<br>

```{r example4.5}
# Example of a plot using theme_light()
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake) + 
  geom_point() +
  theme_light()
```

<br>


```{r example5}
# Example of a plot using theme_dark()
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake) + 
  geom_point() +
  theme_dark()
```

<br>

```{r example5.5}
# Example of a plot using theme_minimal()
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake) + 
  geom_point() +
  theme_minimal()
```

<br>

```{r example6}
# Example of a plot using theme_classic()
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake) + 
  geom_point() +
  theme_classic()
```




<br>

```{r example7}
# Example of a plot using theme_void()
# This theme is completely blank!
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake) + 
  geom_point() +
  theme_void()
```



```{r returnmsleep, echo=F, include=F}
msleep <- introverse::msleep
```