---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(ggplot2)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>ggplot2::geom_text() and ggplot2::geom_label()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`geom_text()`](https://ggplot2.tidyverse.org/reference/geom_text.html) and [`geom_label()`](https://ggplot2.tidyverse.org/reference/geom_text.html)  functions are part of the [`{ggplot2}`](https://ggplot2.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

These functions are used within `{ggplot2}` plots to create text in plots. The `geom_text()` function simply writes text, but the `geom_label()` function includes a rectangle behind the text.

#### Required aesthetics:

+ `x`: The x-axis coordinate(s) for the text(s)
+ `y`: The y-axis coordinate(s) for the text(s)


#### Commonly used optional aesthetics:

+ `color` (`colour`): The text color
+ `size`: The text size
+ `fill`: A rectangle fill when using `geom_label()` (not `geom_text()`)



<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use these functions, you need to either first load the `{ggplot2}` library, or always use the function with `ggplot2::geom_text()` (or `ggplot2::geom_label()`) notation.

```{r description, eval=FALSE}
# Load the library
library(ggplot2)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation for example with geom_text()
ggplot2::geom_text()
```



### Examples

The examples below use the `msleep` dataset. Learn more about this dataset with `get_help("msleep")`.

```{r msleep}
# Show the msleep dataset with head()
head(msleep)
```

<br>
```{r example1}
# Plot of awake against sleep_rem ("y against x"), but using `geom_text()` instead of `geom_point()`.
# Using the label argument, all "text points" are the letter "w". This is "just a label"
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake) + 
  geom_text(label = "w")
```

<br>
```{r example1.5}
# Plot of awake against sleep_rem ("y against x"), but using `geom_label()` instead of `geom_point()`.
# Using the label argument, all "text points" are the letter "w". This is "just a label"
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake) + 
  geom_label(label = "w")
```

<br>
```{r example1.75}
# Plot of awake against sleep_rem ("y against x"), but using `geom_label()` instead of `geom_point()`.
# Include "just a fill" for the geom_label rectangles:
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake) + 
  geom_label(label = "w", fill = "yellow")
```

<br>

```{r example2}
# Plot of awake against sleep_rem ("y against x"), but using `geom_text()` instead of `geom_point()`.
# Now we also make all labels blue, and all labels now read 'hello'
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake) + 
  geom_text(label = "hello", color = "blue")
```

<br>

```{r example3, warning=FALSE}
# Plot of awake against sleep_rem ("y against x"), but using `geom_text()` instead of `geom_point()`.
# Now we map the vore column to label
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake, label = vore) + 
  geom_text(color = "blue")
```


<br>

```{r example3.5}
# Plot of awake against sleep_rem ("y against x"), but using `geom_text()` instead of `geom_point()`.
# Now we map the vore column to label AND to color
ggplot(msleep) + 
  aes(x = sleep_rem, 
      y = awake, 
      label = vore, 
      color = vore) + 
  geom_text()
```

<br>

```{r example3.75}
# Plot of awake against sleep_rem ("y against x"), but using `geom_label()` instead of `geom_point()`.
# Now we map FILL and label to vore as an example of using this function (not a good plot though since excessive overlap!)
ggplot(msleep) + 
  aes(x = sleep_rem, 
      y = awake, 
      label = vore, 
      fill = vore) + 
  geom_label()
```

<br>

```{r example4}
# A wrangled-using-dplyr barplot showing the mean awake value for each vore, where bars are labeled with mean awake values.
# Subsequent examples all modify this plot.
# To explore this example, see it in action one line at a time!
# The next example presents a "cleaned" version of this plot
msleep %>%
  # Calculate mean hours awake for each vore
  group_by(vore) %>%
  summarize(mean_hours_awake = mean(awake)) %>%
  ggplot() + 
  aes(x = vore, 
      y = mean_hours_awake,
      # Set the label aesthetic for geom_text
      label = mean_hours_awake) + 
  geom_col() + 
  geom_text()
```

<br>


```{r example5}
# The previous examples labels were at the exact bar height - it would look nicer if they were a little bit higher up on the Y-axis higher
# We can achieve this by specifying aesthetics separately for each geom: 
msleep %>%
  # Calculate mean hours awake for each vore
  group_by(vore) %>%
  summarize(mean_hours_awake = mean(awake)) %>%
  ggplot() + 
  geom_col(aes(x = vore, 
               # column height shoud be the mean hours spent awake
               y = mean_hours_awake)) + 
  geom_text(aes(x = vore, 
                # label height should be 0.5 units or so higher than the bar! You can add any value you want here, in units of the Y-axis
                # trial and error to find how much higher!
                y = mean_hours_awake + 0.5,
                label = mean_hours_awake))
```
<br>


```{r example6}
# This example further modifies the plot by rounding mean values in the label so they are more legible
msleep %>%
  # Calculate mean hours awake for each vore
  group_by(vore) %>%
  summarize(mean_hours_awake = mean(awake)) %>%
  ggplot() + 
  geom_col(aes(x = vore, 
               y = mean_hours_awake)) + 
  geom_text(aes(x = vore, 
                y = mean_hours_awake + 0.5,
                # Round the LABEL (not the bar height!) to two decimal places. See get_help("round")
                label = round(mean_hours_awake, 2)))
```
```{r example7}
# This example even further modifies the bar plot by arranging bars in order of mean hours spent awake with fct_reorder
msleep %>%
  # Calculate mean hours awake for each vore
  group_by(vore) %>%
  summarize(mean_hours_awake = mean(awake)) %>%
  ggplot() + 
  geom_col(aes(x = fct_reorder(vore, mean_hours_awake), # reorder vore
               y = mean_hours_awake)) + 
  geom_text(aes(x = fct_reorder(vore, mean_hours_awake), # reorder vore
                y = mean_hours_awake + 0.5,
                label = round(mean_hours_awake, 2)))
```

```{r example8}
# This final example even further modifies the bar plot by arranging bars in order of mean hours spent awake with fct_reorder
# But this time, the vore variable is reordered before plotting and NAs are removed:
msleep %>%
  # Calculate mean hours awake for each vore
  group_by(vore) %>%
  summarize(mean_hours_awake = mean(awake)) %>%
  # reorder the vore factor here
  mutate(vore = fct_reorder(vore, mean_hours_awake)) %>%
  # drop any NAs with tidyr::drop_na()
  tidyr::drop_na() %>%
  ggplot() + 
  geom_col(aes(x = vore, #
               y = mean_hours_awake)) + 
  geom_text(aes(x = vore,
                y = mean_hours_awake + 0.5,
                label = round(mean_hours_awake, 2)))
```
