---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = TRUE)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>ggplot2::ggsave()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`ggsave()`](https://ggplot2.tidyverse.org/reference/ggsave.html) function is part of the [`{ggplot2}`](https://ggplot2.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

We use this function within `{ggplot2}` plots to export and _save_ plots to files. 

<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{ggplot2}` library, or always use the function with `ggplot2::ggsave()` notation.

```{r description, eval=FALSE}
# Load the library
library(ggplot2)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
ggplot2::ggsave()
```

### Conceptual Usage


```{r conceptual, eval=FALSE}
# Although there are other ways to use the `ggsave()` function, this usage will be the MOST RELIABLE AND REPRODUCIBLE!
ggsave(name of file to save the plot to with extension determining filetype,
       plot variable you want to save,
       # Width and height require some trial/error! Default width/height are not reliably reproducible, so always specify!
       width = width of outputted file in inches,
       height = height of outputted file in inches)
```

### Examples

The examples below use the `msleep` dataset. Learn more about this dataset with `get_help("msleep")`.

```{r msleep}
# Show the msleep dataset with head()
head(msleep)
```

<br>
```{r example1, eval=FALSE}
# Create a scatterplot to save.
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake) + 
  geom_point() -> my_plot

# Save to a PNG image file called "my_saved_plot.png"
ggsave("my_saved_plot.png", # output file name to save a PNG file
       my_plot,             # previously defined variable containing the plot we want to save
       width = 6,           # The image will be 6 inches wide
       height = 6)          #  The image will be 6 inches tall
```

<br>

```{r example2, eval=FALSE}
# Create a scatterplot to save.
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake) + 
  geom_point() -> my_plot

# Save to a PDF image file called "my_saved_plot.pdf"
ggsave("my_saved_plot.pdf", # output file name to save a PDF file
       my_plot,             # previously defined variable containing the plot we want to save
       width = 6,           # The image will be 6 inches wide
       height = 6)          #  The image will be 6 inches tall
```
