---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = TRUE)
library(ggplot2)
library(dplyr)
msleep %>% tidyr::drop_na(sleep_rem, awake) -> msleep
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;<code>ggplot2::scale_*_manual()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

All functions documented here are part of the [`{ggplot2}`](https://ggplot2.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

The `{ggplot2}` package [provides several useful functions](https://ggplot2.tidyverse.org/reference/scale_manual.html) for create your own aesthetic mapping scales. With these functions, you can manually customize how colors, fills, shapes, sizes, alpha (transparency), and and linetypes are mapped to your data. _You should only use these functions when you have MAPPED an aesthetic._ If you are trying to specify a color, fill, shape, etc. directly without any mapping, simply provide the argument to the geom of interest.

All functions are named `scale_AESTHETIC-TYPE_manual()`, where `AESTHETIC-TYPE` is a type of aesthetic mapping you have specified. These functions include:

| Function | What we use it for
|:-----------|:-------------
|  `scale_color_manual()` (aka `scale_colour_manual()`)   | Specify custom colors to use for mapping
|  `scale_fill_manual()`  | Specify custom fills to use for mapping
|  `scale_shape_manual()`  | Specify custom fills to use for mapping
|  `scale_alpha_manual()`  | Specify custom alpha (transparency) levels to use for mapping
|  `scale_size_manual()`  | Specify custom sizes to use for mapping. This is generally _not recommended_ unless you really know what you're doing, since it is easy to distort sizes and make interpretation difficult
|  `scale_linetype_manual()`  | Specify custom linetypes to use for mapping


<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use these function, you need to either first load the `{ggplot2}` library, or always use the function with `ggplot2::()` notation.

```{r description, eval=FALSE}
# Load the library
library(ggplot2)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation, for example with scale_color_manual()
ggplot2::scale_color_manual()
```


### Conceptual Usage


```{r conceptual, eval=FALSE}
# Most basic usage:
scale_AESTHETIC-TYPE_manual(values = c(array, of, values, to, use, for, mapping))

# Note that the legend name can also be specified with `ggplot2::labs()`
scale_AESTHETIC-TYPE_manual(name = Name of legend associated with this mapping,
                            values = c(array, of, values, to, use, for, mapping))

# Most basic usage:
scale_AESTHETIC-TYPE_manual(values = c(array, of, values, to, use, for, mapping),
                            na.value = specify value to use for data mapping to NA)

```

### Examples

The examples below use the `msleep` dataset. Learn more about this dataset with `get_help("msleep")`.

```{r msleep}
# Show the msleep dataset with head()
head(msleep)
```

<br>
```{r example1}
# Specify custom colors for the `vore` color mapping, ensuring the number of values MATCHES the number of vores (except NAs)
ggplot(msleep) + 
  aes(x = sleep_rem, 
      y = awake, 
      color = vore) + 
  geom_point() +
  scale_color_manual(
    # There are 4 vores, so we need 4 colors
    values = c("blue", "red", "orange", "pink")
  )
```

<br>


```{r example1.5}
# Specify custom colors for the `vore` color mapping, ensuring the number of values MATCHES the number of vores (except NAs)
# This time, use the `name` argument for a customized legend title
ggplot(msleep) + 
  aes(x = sleep_rem, 
      y = awake, 
      color = vore) + 
  geom_point() +
  scale_color_manual(
    name = "Customized new legend title",
    # There are 4 vores, so we need 4 colors
    values = c("blue", "red", "orange", "pink")
  )
```

<br>
```{r example2}
# Instead of the previous example, you can also `labs()` to specify the legend title 
ggplot(msleep) + 
  aes(x = sleep_rem, 
      y = awake, 
      color = vore) + 
  geom_point() +
  scale_color_manual( values = c("blue", "red", "orange", "pink")) + 
  labs(color = "Customized new legend title")
```

<br>


```{r example3}
# To customize the color used for NA values, use the argument `na.value`
ggplot(msleep) + 
  aes(x = sleep_rem, 
      y = awake, 
      color = vore) + 
  geom_point() +
  scale_color_manual(
    values = c("blue", "red", "orange", "pink"),
    na.value = "yellow" # Now all vore that are NA will be yellow instead of the default dark gray
  )
```


<br>

```{r example4}
# CAUTION!! If you don't include an aesthetic in the first place, none of your manual values appear
ggplot(msleep) + 
  aes(x = sleep_rem, 
      y = awake) + # forgetting the color = vore means no colors and no legend show up 
  geom_point() +
  scale_color_manual(values = c("blue", "red", "orange", "pink"))
```


<br>

```{r example5}
# Specify custom shapes for the `vore` shape mapping
# Use `get_help("geom_point")` to see different types of shapes in R plotting
ggplot(msleep) + 
  aes(x = sleep_rem, 
      y = awake, 
      shape = vore) + 
  geom_point() +
  scale_shape_manual(
    values = c(1, 2, 3, 4), # There are 4 vores, so we need 4 shapes
    na.value = 19 # optionally, specify a shape for the NAs
  )
```


<br>

```{r example6}
# Specify custom fills for the `vore` fill mapping
ggplot(msleep) + 
  aes(x = vore, 
      y = awake, 
      fill = vore) + 
  geom_boxplot() +
  scale_fill_manual(
    values = c("blue", "red", "orange", "pink"),
    na.value = "yellow" 
  )
```


<br>

```{r example7}
# CAUTION!! If you specify a fill aesthetic but you don't use scale_FILL_manual, it won't work
# Instead, resulting fills will be _the default fills_
ggplot(msleep) + 
  aes(x = vore, 
      y = awake, 
      fill = vore) + 
  geom_boxplot() +
  # Accidentally applying a manual color scale doesn't change the plot, since a FILL aesthetic was used
  scale_color_manual(
    values = c("blue", "red", "orange", "pink"),
    na.value = "yellow" 
  )
```


<br>

```{r example8}
# Specify custom alphas for the vore alpha mapping
# We also include a fill aesthetic to vore so that we have fills to make transparent
ggplot(msleep) + 
  aes(x = vore, 
      y = awake, 
      alpha = vore, 
      fill = vore) + 
  geom_boxplot() +
  scale_alpha_manual(
    # alpha can range from 0-1. Notice how fills become increasingly opaque
    values = c(0.2, 0.4, 0.6, 0.8),
    na.value = 1
  )
```

<br>


```{r example9}
# Create a dataset over time to demonstrate working with `linetype`
# Use `get_help("tribble")` to learn about making your own tibbles directly in R.
tibble::tribble(
  ~year, ~value, ~data_group,
  2000, 5, "A",
  2000, 7, "B",
  2000, 8, "C",
  2001, 7, "A",
  2001, 12, "B",
  2001, 11, "C",
  2002, 9, "A",
  2002, 14, "B",
  2002, 16, "C"
) -> time_demo_tibble

# Show tibble:
time_demo_tibble

# Plot value over time, with one line per group, using `scale_linetype_manual()` to customize lines
# See `get_help("geom_line")` for more information about types of line types
ggplot(time_demo_tibble) + 
  aes(x = year, 
      y = value,
      group = data_group, # needed for grouped line plots
      linetype = data_group) + 
  geom_line() +
  scale_linetype_manual(
    values = c(3,1,2)
  )
```

```{r, returnmsleep, echo=F, include=F}
msleep <- introverse::msleep
```