---
title: "`ggplot2::geom_col()`"
author: "`{introverse}` quick docs"
output: 
  html_document:
    highlight: pygments
    theme: flatly
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = TRUE)
library(ggplot2)
library(forcats)
```


## Description

The [`geom_col()`](https://ggplot2.tidyverse.org/reference/geom_col.html) function is part of the [`{ggplot2}`](https://ggplot2.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

The `geom_col()` function is used within `{ggplot2}` plots to create barplots where the bar heights' are directly taken from a variable in the dataset. When you call `geom_col()`, `{ggplot2}` will automatically count the number of observations in each category of your primary variable and then plot that calculated value.

To instead plot bars whose heights are equal to _the counted number of categories_, it is recommended to use `ggplot2::geom_bar()`.


#### Required aesthetics:

+ `x`: A categorical variable to visualize bars across
+ `y`: A variable that determines the height of the bars
  + Note that these can be switched for a horizontal bar plot

#### Commonly used optional aesthetics:

+ `color` (`colour`): The outline color for the bars
+ `fill`: The fill for the bars
+ `size`: The width of the bar outlines


<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{ggplot2}` library, or always use the function with `ggplot2::geom_col()` notation.

```{r description, eval=FALSE}
# Load the library
library(ggplot2)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
ggplot2::geom_col()
```



## Examples

> Run `get_help("msleep")` for a refresher on the `msleep` dataset, and run `get_help("carnivores")` for a refresher on the `carnivores` dataset.


```{r example1}
# Plot the literal values of `awake` for each mammal in carnivores
ggplot(carnivores) + 
  aes(x = name, y = awake) + 
  geom_col()
```

<br>


```{r example2}
# Plot the literal values of `awake` for each mammal in carnivores, with additional stylings
ggplot(carnivores) + 
  aes(x = name, y = awake) + 
  geom_col(color = "goldenrod",
           size = 2,
           fill = "blue")
```

<br>
```{r example3}
# Horizontally plot the literal values of `awake` for each mammal in carnivores
ggplot(carnivores) + 
  aes(x = awake, y = name) + # note that these were switched to make it horizontal
  geom_col()
```

<br>

```{r example4}
# Use forcats::fct_reorder() to order the names by their awake values for a cleaner look
# It best practice to also update the x-axis title 
ggplot(carnivores) + 
  aes(x = fct_reorder(name, awake), 
      y = awake) + 
  geom_col() + 
  labs(x = "name")
```

