---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = TRUE)
library(ggplot2)
library(dplyr)
msleep %>% tidyr::drop_na(sleep_rem, awake) -> msleep
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;<code>ggplot2::labs()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`labs()`](https://ggplot2.tidyverse.org/reference/labs.html) function is part of the [`{ggplot2}`](https://ggplot2.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

We use this function within `{ggplot2}` plots to add _labels_, including the following types of labels:


| Label type | Usage as argument to `labs()` |
|:----------|:-------------|
| X-axis label | `labs(x = "x-axis label")` |
| Y-axis label | `labs(y = "y-axis label")` |
| Plot title | `labs(title = "plot title")` |
| Plot subtitle | `labs(subtitle = "plot subtitle")` |
| Plot caption | `labs(caption = "plot caption")` |
| Legend title | `labs(type of aesthetic in legend = "name of legend")` <br> A type of aesthetic may be `fill`, `color`, etc. |



<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{ggplot2}` library, or always use the function with `ggplot2::labs()` notation.

```{r description, eval=FALSE}
# Load the library
library(ggplot2)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
ggplot2::labs()
```



### Examples

> Run `get_help("msleep")` for a refresher on the `msleep` dataset.

```{r example1}
# A fully labeled scatterplot of awake against sleep_rem
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake) + 
  geom_point() +
  labs(x = "Time in REM sleep (hours)",
       y = "Time spent awake (hours)",
       title = "Plot of REM sleep against awake time",
       subtitle = "What a great scatterplot subtitle this is!",
       caption = "Captions are a good place to include the Plot Author")
```

<br>

```{r example2}
# Labeling the legend created by the `color` aesthetic
# Note to remove NAs, you must manipulate the data first
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake, color = vore) + 
  geom_point() +
  labs(x = "Time in REM sleep (hours)",
       y = "Time spent awake (hours)",
       color = "Custom legend title for vore color mapping")
```

<br>


```{r example3}
# Labeling the legend created by the `fill` aesthetic
# Note to remove NAs, you must manipulate the data first
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake, fill = vore) + 
  geom_point(shape = 21) +
  labs(x = "Time in REM sleep (hours)",
       y = "Time spent awake (hours)",
       fill = "Custom legend title for vore fill mapping")
```


<br>

```{r example4}
## CAUTION!! This legend title is buggy and does not appear,
# since the plot doesn't contain a fill aesthetic - it contains a color
# The default legend title is used.
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake, color = vore) + 
  geom_point() +
  labs(x = "Time in REM sleep (hours)",
       y = "Time spent awake (hours)",
       fill = "This title isn't going to appear since there is no fill aes")
```


```{r, returnmsleep, echo=F, include=F}
msleep <- introverse::msleep
```