---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = TRUE)
library(ggplot2)
library(forcats)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>ggplot2::geom_segment()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`geom_segment()`](https://ggplot2.tidyverse.org/reference/geom_segment.html) function is part of the [`{ggplot2}`](https://ggplot2.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

The `geom_segment()` function is used within `{ggplot2}` plots to create line segments. Line segments have x and y start and end coordinates.


#### Required aesthetics:

+ `x`: The x-coordinate for the line segment to start at
+ `xend`: The x-coordinate for the line segment to end at
+ `y`: The y-coordinate for the line segment to start at
+ `yend`: The y-coordinate for the line segment to end at

#### Commonly used optional aesthetics:

+ `color` (`colour`): The color of the line segment
+ `size`: The width of the line segment
+ `linetype`: The style of the line segment, where options include `"solid"`, `"dashed"`, `"dotted"`, `"dotdash"`, `"longdash"`, `"twodash"`. The default is `"solid"`.


<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{ggplot2}` library, or always use the function with `ggplot2::geom_segment()` notation.

```{r description, eval=FALSE}
# Load the library
library(ggplot2)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
ggplot2::geom_segment()
```



### Examples

Some of the examples below use the `msleep` and `carnivores` datasets. Learn more about this dataset with `get_help("msleep")` and `get_help("carnivores")`.

```{r msleep-carnivores}
# Show the msleep dataset with head()
head(msleep)

# Show the carnivores dataset
carnivores
```

<br>

```{r example1}
# Create a tibble with tibble::tribble() with values for the line segment we want to draw, and draw it!
example_tibble <- tibble::tribble(
  ~x_var, ~xend_var, ~y_var, ~yend_var,
  0,         3.5,        2,       4
)

# Look at tibble:
example_tibble

ggplot(example_tibble) + 
  aes(x = x_var, 
      xend = xend_var,
      y = y_var,
      yend = yend_var) + 
  geom_segment()

```

<br>

```{r example2}
# Scatterplot of awake against sleep_rem, where we add a line segment in at specified coordinates
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake) + 
  geom_point() + 
  # Add a segment:
  geom_segment(x = 0, xend = 2, y = 7, yend = 9, color = "red")
```

<br>

```{r example3}
# Scatterplot of awake against sleep_rem, where we add a line segment in at specified coordinates
# This example uses a different linetype
ggplot(msleep) + 
  aes(x = sleep_rem, y = awake) + 
  geom_point() + 
  # Add a segment:
  geom_segment(x = 0, xend = 2, y = 7, yend = 9, color = "red", linetype = "dashed")
```

<br>

```{r example4}
# Plot the literal values of `awake` for each name in the carnivores dataset, as a lollipop plot
ggplot(carnivores) + 
  aes(x = name, y = awake) +
  geom_point() + 
  # To understand this, think carefully about the grammar and how data maps to these aesthetics
  geom_segment(aes(x = name,   # For a vertical line, x starts and stops at same coordinate. 
                   xend = name,
                   y = 0, # y starts from 0
                   yend = awake)) # and goes to the awake value
```
