---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = TRUE)
library(ggplot2)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>ggplot2::geom_histogram()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`geom_histogram()`](https://ggplot2.tidyverse.org/reference/geom_histogram.html) function is part of the [`{ggplot2}`](https://ggplot2.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

The `geom_histogram()` function is used within `{ggplot2}` plots to create a histogram, which shows the distribution of a _numeric continuous_ variable. The x-axis shows values of the variable, and the y-axis shows binned counts of how often values in a given range occur. By default, `geom_histogram()` uses 30 bins.

#### Required aesthetics:

+ `x`: The numeric variable whose distribution the histogram visualizes. 
  + Note that it is also allowed to make a horizontal histogram by providing a `y` aesthetic instead of an `x` (but never both)

#### Commonly used optional aesthetics:

+ `color` (`colour`): The outline color for the histogram bars
+ `fill`: The fill for the histogram bars
+ `size`: The width of the histogram bar outlines
+ `bins`: How many bins should be used (will be overridden by `binwidth` if both are provided)
+ `binwidth`: The width of each bin (will override `bins` if both are provided)

<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{ggplot2}` library, or always use the function with `ggplot2::geom_histogram()` notation.

```{r description, eval=FALSE}
# Load the library
library(ggplot2)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
ggplot2::geom_histogram()
```



### Examples

The examples below use the `msleep` dataset. Learn more about this dataset with `get_help("msleep")`.

```{r msleep}
# Show the msleep dataset with head()
head(msleep)
```

<br>
```{r example1}
# By default, histograms have no color outline, are filled with dark gray, and use 30 bins
# If you do not specify binning, you will get a _normal_ message telling you 30 were chosen
ggplot(msleep) + 
  aes(x = awake) + 
  geom_histogram()
```

<br>

```{r example2}
# Change the binning with `bins`
ggplot(msleep) + 
  aes(x = awake) + 
  geom_histogram(bins = 10)
```

<br>

```{r example3}
# Change the binning with `binwidth`
ggplot(msleep) + 
  aes(x = awake) + 
  geom_histogram(binwidth = 0.5)
```

<br>

```{r example4}
# Include custom fill and color
ggplot(msleep) + 
  aes(x = awake) + 
  geom_histogram(color = "black", fill = "purple")
```


<br>

```{r example5}
# Include custom fill and color, with a larger outline size
# When you have many arguments, placing on separate lines makes code easier to read and work with
ggplot(msleep) + 
  aes(x = awake) + 
  geom_histogram(color = "black", 
                 fill = "purple", 
                 size = 2)
```



<br>

```{r example6}
# A horizontal histogram using y instead of x aesthetic
ggplot(msleep) + 
  aes(y = awake) + 
  geom_histogram()
```


