---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = TRUE)
library(ggplot2)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;<code>ggplot2::geom_density()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`geom_density()`](https://ggplot2.tidyverse.org/reference/geom_density.html) function is part of the [`{ggplot2}`](https://ggplot2.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

The `geom_density()` function is used within `{ggplot2}` plots to create a density plot, which shows the distribution of a _numeric continuous_ variable with the appearance of a "smoothened histogram." The x-axis shows values of the variable, and the y-axis shows the statistical density calculation (this is not a matched value in the dataset!).

#### Required aesthetics:

+ `x`: The numeric variable whose distribution to visualize
  + Note that it is also allowed to make a horizontal density plot by providing a `y` aesthetic instead of an `x` (but never both)

#### Commonly used optional aesthetics:

+ `color` (`colour`): The outline color for the density plot
+ `fill`: The fill for the density plot
+ `alpha`: Transparency for the density plot fill, ranging from 0-1 where 0 is fully transparent and 1 is fully opaque.
+ `size`: The width of the density plot outline


<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{ggplot2}` library, or always use the function with `ggplot2::geom_density()` notation.

```{r description, eval=FALSE}
# Load the library
library(ggplot2)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
ggplot2::geom_density()
```



### Examples

> Run `get_help("msleep")` for a refresher on the `msleep` dataset.

```{r example1}
ggplot(msleep) + 
  aes(x = awake) + 
  geom_density()
```

<br>

```{r example2}
# Include custom fill and color
ggplot(msleep) + 
  aes(x = awake) + 
  geom_density(fill = "purple", color = "orange")
```

<br>


```{r example2.5}
# Include custom fill, color, and outline size
# When you have many arguments, placing on separate lines makes code easier to read and work with
ggplot(msleep) + 
  aes(x = awake) + 
  geom_density(fill = "purple", 
               color = "orange",
               size = 2)
```

<br>

```{r example3}
# Show overlapping density plots in a single panel
# Important to use `alpha` (find a good one with trial/error) so data is not hidden
# Note: to remove NAs, you must manipulate the data first
ggplot(msleep) + 
  aes(x = awake, fill = vore) + 
  geom_density(alpha = 0.6)
```


<br>

```{r example4}
# A horizontal density plot using y instead of x aesthetic
ggplot(msleep) + 
  aes(y = awake) + 
  geom_density(fill = "purple")
```
