---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(glue)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>glue::glue()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`glue()`](https://glue.tidyverse.org/reference/glue.html) function is part of the identically-named [`{glue}`](https://glue.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).


We use `glue()` conveniently combine multiple strings into one. `glue()` is similar to the base R function `paste()`, except it makes it much easier to _include variables_ in your string. By default, `glue()` will not insert any space or chatacter between strings being combined. You can change this behavior with the argument `sep`.

<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{glue}` library, or always use the function with `glue::glue()` notation.

```{r description, eval=FALSE}
# Load the library
library(glue)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
glue::glue()
```


### Conceptual Usage 

```{r usage, eval=FALSE}
glue("combine this string {and this variable} into a full single string")
```

### Examples

Some examples below use the `carnivores` dataset. Learn more about this dataset with `get_help("carnivores")`.

```{r carnivores}
# Show the carnivores dataset
carnivores
```

<br>

```{r example1}
# Quickly create strings that contain variables using {} around variable names, without ending quotes!
best_animals <- "bears"
worst_animals <- "mosquitos"
glue("We love {best_animals}, but we hate {worst_animals}.")
```

```{r example2}
# Another example of embedding variables into a string. Again, quotes enclose the entire argument to `glue()`
favorite_package <- "introverse"
glue("My favorite R package is the {favorite_package}.")
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->

```{r example3}
# Create a new column called `full_info` in carnivores that is comprised of: "name (genus)"
# Again, directly refer to column names but surround them in curly braces {}
# You'll notice the column comes out a <chr> type. That's ok! It behaves like <chr>
carnivores %>% 
  dplyr::mutate(full_info = glue("{name} ({genus})"))
```

<br>

```{r example4}
# You can also use as.character() to ensure a <chr> column type that was created with glue()
# Use `get_help("as.character") to learn more
carnivores %>% 
  # Create column and coerce it into a character type
  dplyr::mutate(full_info = glue("{name} ({genus})"),
                full_info = as.character(full_info)) 
```

<br>
```{r example5}
# Create a new column called `description` in carnivores that is comprised of the sentence:
#    "The species {name} is awake, on average, {awake} hours per day."
carnivores %>% 
  dplyr::mutate(description = glue("The species {name} is awake {awake} hours per day."))
```