---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      collapse = TRUE)
library(dplyr)
`%<>%` <- magrittr::`%<>%`
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>magrittr::%<>%</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The combined symbol [`%<>%`](https://magrittr.tidyverse.org/reference/compound.html) is an operator in the [`{magrittr}`](https://magrittr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/). This operator simultaneously pipes _and_ redefines the first input as the result. It is useful for simultaneously wrangling and modifying tibbles (data frames) in place.

To learn more about `magrittr` piping, run `get_help("%>%")`.

To use the assignment pipe, you need to either first load the `{tidyverse}` or the `{magrittr}` library.

```{r description, eval=FALSE}
# Load the tidyverse (most common)
library(tidyverse)

# Or load magrittr
library(magrittr)
```

### Conceptual Usage 

```{r usage, eval=FALSE}
tibble %<>%
  mutate(new_column = contents of new column)

# The above is equivalent to:
tibble %>%
  mutate(new_column = contents of new column) -> tibble
```

### Examples

The example below uses the `carnivores` dataset. Learn more about this dataset with `get_help("carnivores")`.

```{r carnivores}
# Show the carnivores dataset
carnivores
```

<br>

```{r example1}
# Simultaneously rename the column `name` to `species` and redefine carnivores
carnivores %<>%
  rename(species = name)

carnivores
```


```{r, include=FALSE, echo=FALSE}
# reassign in case
carnivores %<>%
  rename(name = species)
```