---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      collapse = TRUE)
library(dplyr)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;<code>magrittr::%>%</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>{introverse}</code> quick docs</b>
</div>
<br>

### Description

The combined symbol [`%>%`](https://magrittr.tidyverse.org/reference/pipe.html) is an operator in the [`{magrittr}`](https://magrittr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/). This operator is referred to as "pipe," but it should not be confused with the symbol `|` (on backslash key), which is also known as pipe.

The `%>%` operator "sends in" input to a given function. It provides a convenient approach to constructing _pipelines_, where the output from one command can be seamlessly provided into the next command. 

To use the pipe, you need to either first load the `{tidyverse}`, or one of the core `{tidyverse}` libraries.

```{r description, eval=FALSE}
# Load the tidyverse (most common)
library(tidyverse)

# Or load a core tidyverse library such as dplyr
library(dplyr)
```


## Conceptual usage 

```{r usage, eval=FALSE}
# Non-pipe version of code:
some_function(argument)

# Pipe version of code:
argument %>%
  some_function()
    
#############################

# Non-pipe version of code:
some_function(argument1, argument2)

# Pipe version of code:
argument1 %>%
  some_function(argument2)
    
############################

# Non-pipe version of code:
some_other_other_function(some_other_function(some_function(argument1, argument2)))

# Or, a different way to write that non-pipe code:
output1 <- some_function(argument1, argument2)
output2 <- some_other_function(output1)
output3 <- some_other_other_function(output2)
  
# Pipe version of code:
argument1 %>%
  some_function(argument2) %>%
  some_other_function() %>%
  some_other_other_function()
```

### Examples

> Run `get_help("carnivores")` for a refresher on the `carnivores` dataset.

```{r example1}
# Calculate the natural log of 5

# Non-pipe version of code:
log(5)

# Pipe version of code:
5 %>%
  log()
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->


```{r example2}
# Calculate the log of 5 in base 7

# Non-pipe version of code:
log(5, 7)

# Pipe version of code:
5 %>%
  log(7)
```
<br> 



```{r example3}
# Calculate the log of 5 in base 7, and THEN calculate the sqrt of the result.

# Non-pipe version of code:
sqrt( log(5, 7) )

# Or, a different non-pipe version:
step1 <- log(5,7)
sqrt(step1)

# Pipe version of code:
5 %>%
  log(7) %>%
  sqrt()
```

<br> 

```{r example4, collapse=TRUE}
# Use dplyr::mutate to add a new column to carnivores called `new_col`

# Non-pipe version of code. _This is NOT recommended._
mutate(carnivores, new_col = 7)

# Pipe version of code. _This IS recommended._
carnivores %>%
  mutate(new_col = 7)
```

<br> 

```{r example5, collapse=TRUE}
# Use dplyr::mutate to add a new column to carnivores called `new_col`
# and then use dplyr::select to keep only `name` and `new_col` columns.

# Non-pipe version of code. _This is NOT recommended._
select(mutate(carnivores, new_col = 7), name, new_col)

# Or, a different non-pipe version (also NOT recommended):
carnivores2 <- mutate(carnivores, new_col = 7)
select(carnivores2, name, new_col)

# Pipe version of code. _This IS recommended._
carnivores %>%
  mutate(new_col = 7) %>%
  select(name, new_col)
```
