---
title: "The {magrittr} assignment pipe `%>%`"
author: "`{introverse}` quick docs"
output: 
  html_document:
    highlight: pygments
    theme: flatly
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      collapse = TRUE)
library(dplyr)
`%<>%` <- magrittr::`%<>%`
```


## Description

The combined symbol [`%<>%`](https://magrittr.tidyverse.org/reference/compound.html) is an operator in the [`{magrittr}`](https://magrittr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/). This operator simultaneously pipes _and_ redefines the first input as the result. It is useful for simultaneously wrangling and modifying tibbles (data frames) in place.

To learn more about `magrittr` piping, run `get_help("%>%")`.

To use the assignment pipe, you need to either first load the `{tidyverse}` or the `{magrittr}` library.

```{r description, eval=FALSE}
# Load the tidyverse (most common)
library(tidyverse)

# Or load magrittr
library(magrittr)
```

## Conceptual usage 

```{r usage, eval=FALSE}
tibble %<>%
  mutate(new_column = contents of new column)

# The above is equivalent to:
tibble %>%
  mutate(new_column = contents of new column) -> tibble
```

## Examples

> Run `get_help("carnivores")` for a refresher on the `carnivores` dataset.

```{r example1}
# Simultaneously rename the column `name` to `species` and redefine carnivores
carnivores %<>%
  rename(species = name)

carnivores
```


```{r, include=FALSE, echo=FALSE}
# reassign in case
carnivores %<>%
  rename(name = species)
```