---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;Mathematical operators
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

Mathematical operators (symbols to do math with) in R include the following:
  

| Operator | What it does 
|:-----------|:-------------
|  `+`      | Addition 
|  `-`      | Subtraction 
|  `*`      | Multiplication 
|  `/`      | Division 
|  `^`      | Exponentiation  
|  `**`     | Exponentiation (no space in between the `**`) 
|  `%%`     | Modulo: Find the remainder from "long division"

All operators can be used on individual numbers or with numeric arrays.

<br>

### Examples


```{r example1}
# Add two numbers
5+8
```

<br>

```{r example2}
# Subtract two numbers
5-8
```

<br>

```{r example3}
# Divide two numbers
5/8
```

<br>

```{r example4}
# Multiply two numbers
5*8
```

<br>

```{r example5}
# Raise 5 to the 8th power
5^8
```


<br>

```{r example6}
# Raise 5 to the 8th power
5**8
```

<br>

```{r example7}
# Find the remainder of 8 divided by 5
8 %% 5
```

<br>

```{r example8}
# Add the number 5 to every value in a numeric array
5 + c(45, 65, 85, 105)
```

<br>

```{r example9}
# Cube every number in a numeric array
c(45, 65, 85, 105)^3
```