---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;Logical operators
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

Logical operators are symbols you can make comparisons with **that tell you `TRUE` or `FALSE`**. A _logical statement_ is any piece of code that evaluates to `TRUE` or `FALSE`, i.e. code that uses a logical operator. Note that in other computing languages, these operators and statements are often called "boolean" (boo-lee-an),  not logical.

Logical operators in `R` include the following, using placeholders `X` and `Y` to represent values/variables being compared:


| Operator | What it asks |
|:----------|:-------------|
|  `X == Y` | Is `X` equal to `Y`? |
|  `X < Y` | Is `X` less than `Y`? |
| `X > Y` | Is `X` greater than `Y`? |
| `X <= Y` | Is `X` less than or equal to `Y`? |
| `X >= Y` | Is `X` greater than or equal to `Y`? |
| `X != Y` | Is `X` not equal to `Y`? |
| `!(X < Y)` | Is `X` not less than `Y`? |
| `!(X > Y)` | Is `X` not greater than `Y`? |
| `X %in% Y_ARRAY` | Is the value `X` present in the array `Y_ARRAY`? |
| `&` | Used to combine multiple logical statements with "and," which means _all_ statements must be `TRUE` to achieve `TRUE`. |
| `|` |  Used to combine multiple logical statements with "or," which means _at least one_ statement must be `TRUE` to achieve `TRUE`. This symbol is known as "pipe" and is located on the backslash key above the return/enter key on your keyboard. |


<br>

### Examples

Some examples below use the `carnivores` dataset. Learn more about this dataset with `get_help("carnivores")`.

```{r carnivores}
# Show the carnivores dataset
carnivores
```

<br>

```{r example1}
# Is 10 equal to 12?
10 == 12
```

<br>

```{r example2}
# Is 14 greater than 13.5?
14 > 13.5
```

<br>

```{r example3}
# Is the length of this array equal to 6? 
length( c(1, 1, 2, 3, 5, 8, 13)  ) == 6
```

<br>

```{r example4}
# Does the following string have less than or equal to 7 characters?
nchar("im a string being evaluated") <= 7
```

<br>

```{r example5}
# Is 10 greater than 5 _and_ 11 less than 10?
10 > 5 & 11 < 10
```

<br>

```{r example6}
# Is 10 greater than 5 _or_ 11 less than 10?
10 > 5 | 11 < 10
```

<br>

```{r example7}
# Is 55 not equal to 27?
55 != 27
```

<br>

```{r example8}
# Is the word "elephant" in the following array?
"elephant" %in% c("hippopotamus", "giraffe", "zebra", "elephant")
```

<br>

```{r example9}
# Is 'Panthera' in the genus column of carnivores?
"Panthera" %in% carnivores$genus
```