---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(readr)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;<code>readr::read_*()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>{introverse}</code> quick docs</b>
</div>
<br>

### Description

All functions documented here are part of the [`{readr}`](https://tibble.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

The `{readr}` package [provides several useful functions](https://readr.tidyverse.org/reference/read_delim.html) for _reading_ delimited data files into R. Delimited data files have rows and columns, where columns are _separated aka delimited_ by a certain character, such as a comma `,` or tab symbol. 

When using these functions, your data automatically gets read in as a tibble. _This is usually preferrable to using base R functions like `read.csv()` or `read.table()`, which create data frames._

These functions include:

| Function | What kinds of files it reads
|:-----------|:-------------
|  `read_csv()`   | CSV (_comma_ separated values) files  
|  `read_tsv()`   | TSV (_tab_ separated values) files  
|  `read_csv2()` | Files with a semi-colon `;` delimiter. (This format is commonly used in certain European countries which use commas instead of periods in their decimals)
|  `read_delim()` | Files with any delimiter symbol, which you specify with the argument `delim`

<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use these function, you need to either first load the `{readr}` library, or always use the function with `readr::` notation.

```{r description, eval=FALSE}
# Load the library
library(readr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
readr::read_csv()
readr::read_tsv()
readr::read_csv2()
readr::read_delim()
```


## Conceptual usage 

```{r usage, eval=FALSE}
read_csv("path/to/file/to/read/in/file.csv")
read_tsv("path/to/file/to/read/in/file.tsv")
read_csv2("path/to/file/to/read/in/file.csv")
read_delim("path/to/file/to/read/in/file.csv", delim = ",") # for example, a comma delimiter 
```

### Examples

<br>
```{r example1, eval=FALSE}
# Read in (hypothetical) data file called "data.csv" which lives in the _same directory_ as your code,
# and save to a variable `my_data`
my_data <- read_csv("data.csv")
```

<br>
```{r example2, eval=FALSE}
# Read in (hypothetical) data file called "data.tsv" which lives in the relative directory "datasets/"
# and save to a variable `my_data`
my_data <- read_tsv("dataset/data.tsv")
```

<br>
```{r example3, eval=FALSE}
# Read in (hypothetical) data file called "data_weird.txt" which lives in the relative directory "datasets/"
# and uses the weird approach of exclamation marks `!` to delimit (separate) columns
my_data <- read_delim("datasets/data_weird.txt", delim = "!")
```

<br>
```{r example4, error=TRUE}
# CAUTION! If you have a typo or an incorrectly specified path to the data, you will get an error saying:
# Your file 'does not exist in current working directory'
my_data <- read_csv("this_file_doesnt_exist.csv")
```

