---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(readr)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;<code>readr::write_*()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

All functions documented here are part of the [`{readr}`](https://readr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

The `{readr}` package [provides several useful functions](https://readr.tidyverse.org/reference/writing_delim.html) for _writing_, i.e. saving, tibbles (data frames) to files. 

These functions include:

| Function | What kinds of files it creates
|:-----------|:-------------
|  `write_csv()`   | CSV (_comma_ separated values) files  
|  `write_tsv()`   | TSV (_tab_ separated values) files  
|  `write_csv2()` | Files with a semi-colon `;` delimiter. (This format is commonly used in certain European countries which use commas instead of periods in their decimals)
|  `write_delim()` | Files with any delimiter symbol, which you specify with the argument `delim`

<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use these function, you need to either first load the `{readr}` library, or always use the function with `readr::` notation.

```{r description, eval=FALSE}
# Load the library
library(readr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
readr::write_csv()
readr::write_tsv()
readr::write_csv2()
readr::write_delim()
```


## Conceptual usage 

```{r usage, eval=FALSE}
write_csv(tibble or data frame variable you want to save,
          "path/to/file/to/write/file.csv")

write_tsv(tibble or data frame variable you want to save,
          "path/to/file/to/write/file.tsv")

write_csv2(tibble or data frame variable you want to save,
          "path/to/file/to/write/file.csv2")

write_delim(tibble or data frame variable you want to save,
          "path/to/file/to/write/file.txt",
          delim = ",") # For example, comma delimiter
```

### Examples

<br>
```{r example1, eval=FALSE}
# Save the `msleep` tibble as a CSV to a file called "msleep.csv"
write_csv(msleep, "msleep.csv")
```

<br>
```{r example2, eval=FALSE}
# Save the `msleep` tibble as a TSV to a file called "msleep.csv" which lives in the relative directory "datasets/"
write_tsv(msleep, "datasets/msleep.tsv")
```

<br>
```{r example3, eval=FALSE}
# Save the `msleep` tibble as a delimited file "msleep_weird.txt" which lives in the relative directory "datasets/"
# and uses the weird approach of exclamation marks `!` to delimit (separate) columns
write_delim(msleep, "datasets/msleep_weird.txt", delim = "!")
```
