---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(stringr)
set.seed(1)
single_sentence <-  sample(stringr::sentences,1)
set.seed(18)
fruits <- sample(stringr::fruit, 6)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>stringr::str_count()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`str_count()`](https://stringr.tidyverse.org/reference/str_count.html) function is part of the [`{stringr}`](https://stringr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).


We use the `str_count()` function to count the number of times a given smaller string or regular expression (a special type of pattern-matching string) appears in a larger string.

All `{stringr}` functions (and `R` itself) are _case sensitive_, which means uppercase and lowercase letters are viewed as different characters. In other words, searching for "A" will not match "a", and searching for "a" will not match "A."

<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{stringr}` library, or always use the function with `stringr::str_count()` notation.

```{r description, eval=FALSE}
# Load the library
library(stringr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
stringr::str_count()
```


### Conceptual Usage 

```{r usage, eval=FALSE}
stringr::str_count("string to count observations in", "substring to look for in bigger string")

stringr::str_count(c("array", "of", "strings", "to", "count", "observations", "in"), 
                   "substring to look for in each string in the array")
```

### Examples

The examples use the variables shown below, as well as the `carnivores` dataset. Learn more about the `carnivores` with `get_help("carnivores")`.

```{r exdata, collapse=TRUE}
# Show all variables and datasets used in examples:

# A single string
single_sentence

# An array of strings
fruits

# A tibble
carnivores
```

<br>


```{r example1}
# Count the number of times the substring "ou" appears in `single_sentence`
str_count(single_sentence, "ou")
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->

```{r example2}
# Count the number of times the capital letter "A" appears in `single_sentence`
str_count(single_sentence, "A")
```
<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->

```{r example3}
# Count the number of times the letter "o" appears in _each string_ in the array `fruits`
# The result is an array that counts each string's number of "o"s
str_count(fruits, "o")
```


```{r example4}
# Count the number of times the letter "a" appears in each genus in carnivores
str_count(carnivores$genus, "a")
```


<br>

```{r example5}
# Use dplyr::mutate() to create a new column in carnivores containing counts of the letter "a" in `genus`
carnivores %>%
  dplyr::mutate(number_a_in_genus = str_count(genus, "a"))
```
