---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(stringr)
set.seed(1)
single_sentence <-  sample(stringr::sentences,1)
set.seed(18)
fruits <- sample(stringr::fruit, 6)
fruits
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;Converting between uppercase and lowercase in <code>{stringr}</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`{stringr}`](https://stringr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/), includes several useful functions for converting strings or arrays of strings between uppercase and lowercase. These functions include:


| Function | How it affects strings | 
|:---------|:-----------------------|
| `str_to_upper()` | Makes all letters in a string _uppercase_ |
| `str_to_lower()` | Makes all letters in a string _lowercase_ |
| `str_to_title()` | Converts a string to _title case_, which means the first letter of each word will uppercase and the rest will be lowercase. |
| `str_to_sentence()` | Converts a string to _sentence case_, where the first letter of the first word is uppercase, and all remaining letters are lowercase (like a sentence). |

<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{stringr}` library, or always use the function with `stringr::()` notation.

```{r description, eval=FALSE}
# Load the library
library(stringr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation with str_to_upper(), for example
stringr::str_to_upper()
```


### Conceptual Usage 

```{r usage, eval=FALSE}
stringr::str_to_upper("string or array of strings to convert to uppercase") 
stringr::str_to_lower("string or array of strings to convert to lowercase") 
stringr::str_to_title("string or array of strings to convert to title case") 
stringr::str_to_sentence("string or array of strings to convert to sentence case") 

```

### Examples

The examples use the variables shown below, as well as the `carnivores` dataset. Learn more about the `carnivores` with `get_help("carnivores")`.

```{r exdata, collapse=TRUE}
# Show all variables and datasets used in examples:

# A single string
single_sentence

# An array of strings
fruits

# A tibble
carnivores
```

```{r example1}
# Convert `single_sentence` to all uppercase
str_to_upper(single_sentence)
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->

```{r example2}
# Convert `single_sentence` to title case
str_to_title(single_sentence)
```
<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->

```{r example3}
# Convert all strings in `fruits` to sentence case
str_to_sentence(fruits)
```


<br>

```{r example4}
# Convert all strings in `fruits` to title case
str_to_title(fruits)
```

<br>

```{r example5}
# Convert the column `name` in carnivores to all uppercase using dplyr::mutate
carnivores %>%
  dplyr::mutate(name = str_to_upper(name))
```


<br>

```{r example6}
# Convert the column `genus` in carnivores to all lowercase using dplyr::mutate
carnivores %>%
  dplyr::mutate(genus = str_to_lower(genus))
```

