---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(stringr)
set.seed(1)
single_sentence <-  sample(stringr::sentences,1)
set.seed(18)
fruits <- sample(stringr::fruit, 6)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:24px;"> 
  &nbsp;<code>stringr::str_replace()</code> and <code>stringr::str_replace_all()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`str_replace()` and `str_replace_all()`](https://stringr.tidyverse.org/reference/str_replace.html) functions are part of the [`{stringr}`](https://stringr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

We use these function find and replace patterns, including strings or or regular expression (a special type of pattern-matching string), in strings. 

The `str_replace()` function will replace _only the first occurrence_ of the given pattern with the given replacement, but `str_replace_all()` will replace _all_ occurrences of the given pattern with the given replacement.

All `{stringr}` functions (and `R` itself) are _case sensitive_, which means uppercase and lowercase letters are viewed as different characters. In other words, searching for "A" will not match "a", and searching for "a" will not match "A."

<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{stringr}` library, or always use the function with `stringr::()` notation.

```{r description, eval=FALSE}
# Load the library
library(stringr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation with str_replace(), for example
stringr::str_replace()
```


### Conceptual Usage 

```{r usage, eval=FALSE}
stringr::str_replace("string to search in", 
                     "pattern to search for",
                     "replacement to replace only the first occurrance of pattern")

stringr::str_replace_all("string to search in", 
                         "pattern to search for",
                         "replacement to replace all occurrances")
```

### Examples

The examples use the variables shown below, as well as the `carnivores` dataset. Learn more about the `carnivores` with `get_help("carnivores")`.

```{r exdata, collapse=TRUE}
# Show all variables and datasets used in examples:

# A single string
single_sentence

# An array of strings
fruits

# A tibble
carnivores
```

<br>


```{r example1}
# Replace the first occurrances of "the" with "THE" in single_sentence
str_replace(single_sentence, "the", "THE")
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->

```{r example2}
# Replace all  occurranceso f "the" with "THE" in single_sentence
str_replace_all(single_sentence, "the", "THE")
```
<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->

```{r example3}
# Replace the first occurrence of letter "i" with the word "EYE" in all strings in fruits
# This will replace the first occurrence separately in each string in fruits
str_replace(fruits, "i", "EYE")
```


<br>

```{r example4}
# Replace all occurrence of letter "i" with the word "EYE" in all strings in fruits
str_replace_all(fruits, "i", "EYE")
```

<br>

```{r example5}
# Replace all occurrences of the substring "berry" with "bear-y" in all fruit strings
# When you know you want all occurrences replaced, it's generally safer to use `str_replace_all()`
str_replace_all(fruits, "berry", "bear-y")
```


<br>

```{r example6}
# Use dplyr::mutate() to create a new column `name_improved` in carnivores that 
#   replaces all occurrences of "seal" in the column `name` with "furbabies"
carnivores %>%
  dplyr::mutate(name_improved = str_replace_all(name, "seal", "furbabies"))
```

