---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(stringr)
set.seed(1)
single_sentence <-  sample(stringr::sentences,1)
set.seed(18)
fruits <- sample(stringr::fruit, 6)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>stringr::str_starts()</code> and <code>stringr::str_ends()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`str_starts()` and `str_ends()`](https://stringr.tidyverse.org/reference/str_starts.html) functions are part of the [`{stringr}`](https://stringr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

We use these functions to ask _if it is `TRUE` or `FALSE`_ that a certain string either _starts with_ (`str_starts()`) or _ends with_ (`str_end()`) a given pattern. A common usage of this function is with `dplyr::filter()` to subset rows from tibbles (data frames) that start and/or end with a certain set of characters.

All `{stringr}` functions (and `R` itself) are _case sensitive_, which means uppercase and lowercase letters are viewed as different characters. In other words, searching for "A" will not match "a", and searching for "a" will not match "A."

<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{stringr}` library, or always use the function with `stringr::` notation.

```{r description, eval=FALSE}
# Load the library
library(stringr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation, for example with str_starts() 
stringr::str_starts()
```


### Conceptual Usage 

```{r usage, eval=FALSE}
stringr::str_starts("string to search in", 
                    "substring you are asking if the first argument starts with")

stringr::str_starts(c("array", "of", "strings", "to", "search", "in"), 
                   "substring you are asking if the first argument starts with")

stringr::str_ends("string to search in", 
                    "substring you are asking if the first argument ends with")

stringr::str_ends(c("array", "of", "strings", "to", "search", "in"), 
                   "substring you are asking if the first argument ends with")
```

### Examples

The examples use the variables shown below, as well as the `carnivores` dataset. Learn more about the `carnivores` with `get_help("carnivores")`.

```{r exdata, collapse=TRUE}
# Show all variables and datasets used in examples:

# A single string
single_sentence

# An array of strings
fruits

# A tibble
carnivores
```

<br>

```{r example1}
# Does `single_sentence` start with "Hello"?
str_starts(single_sentence, "Hello")
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->

```{r example2}
# Does `single_sentence` end with a period?
# To search for a literal period, you need to write: "\\." (include two backslashes)
str_ends(single_sentence, "\\.")
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->

```{r example3}
# Do strings in `fruits` start with "p"?
str_starts(fruits, "p")
```


<br>


```{r example4}
# Do strings in `fruits` end with "y"?
str_ends(fruits, "y")
```

<br>
```{r example5}
# Keep only rows where it is `TRUE` that the `genus` ends in "s" dplyr::filter
carnivores %>%
  dplyr::filter(str_ends(genus, "s"))
```

<br>


```{r example6}
# Keep only rows where it is `FALSE` that genus starts with "C"
# For this, we negate the result of `str_starts()` using `!` to get `FALSE` instead of `TRUE`
carnivores %>%
  dplyr::filter(!str_starts(genus, "C"))
```