---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(tibble)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;<code>tibble::as_tibble()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`as_tibble()`](https://tibble.tidyverse.org/reference/as_tibble.html) function is part of the [`{tibble}`](https://tibble.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

We use the `as_tibble()` function to _coerce_ a variable into being a tibble. Usually, we use this to coerce a data frame into a tibble to achieve improved printing and other tibble behaviors.

<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{tibble}` library, or always use the function with `tibble::as_tibble()` notation.

```{r description, eval=FALSE}
# Load the library
library(tibble)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
tibble::as_tibble()
```


## Conceptual usage 

```{r usage, eval=FALSE}
as_tibble(variable to coerce into a tibble)
```

### Examples

<br>
```{r example1}
# Coerce the built-in data frame iris into a tibble
iris_tibble <- as_tibble(iris)

iris_tibble
```
