---
title: "`tibble::as_tibble()`"
author: "`{introverse}` quick docs"
output: 
  html_document:
    highlight: pygments
    theme: flatly
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(tibble)
```


## Description

The [`as_tibble()`](https://tibble.tidyverse.org/reference/as_tibble.html) function is part of the [`{tibble}`](https://tibble.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

We use the `as_tibble()` function to _coerce_ a variable into being a tibble. Usually, we use this to coerce a data frame into a tibble to achieve improved printing and other tibble behaviors.

<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{tibble}` library, or always use the function with `tibble::as_tibble()` notation.

```{r description, eval=FALSE}
# Load the library
library(tibble)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
tibble::as_tibble()
```


## Conceptual usage 

```{r usage, eval=FALSE}
as_tibble(variable to coerce into a tibble)
```

## Examples

<br>
```{r example1}
# Coerce the built-in data frame iris into a tibble
iris_tibble <- as_tibble(iris)

iris_tibble
```
