---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(tibble)
example_tibble <- tribble(
  ~food,         ~color,     ~food_type,
  "cucumber",    "green",    "fruit",
  "pomegranate", "red",      "fruit",
  "carrot",      "orange",    "root")
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>tibble::tibble()</code> and <code>tibble::tribble()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`tibble()`](https://tibble.tidyverse.org/reference/tibble.html) and [`tribble()`](https://tibble.tidyverse.org/reference/tribble.html) functions are part of the [`{tibble}`](https://tibble.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

We use the functions `tibble()` and `tribble()` functions to quickly create tibbles from scratch within R. You can use either function depending on your personal preference.


<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use these functions, you need to either first load the `{tibble}` library, or always use the function with `tibble::` notation.

```{r description, eval=FALSE}
# Load the library
library(tibble)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
tibble::tibble()
tibble::tribble()
```


### Conceptual Usage 

```{r usage, eval=FALSE}
# We give arrays to the tibble() function. 
# Arrays should be of the same length and will become the columns
tibble(column_name1 = values for column 1,
       column_name2 = values for column 2,
       etc.)


# We use special syntax for tribble that writes the data "appearing like a table"
# The tildes `~` are used to indicate column names
tribble(
  ~column_name1,         ~column_name2,        ~etc for more columns,
  first row of column1,  first row of column2,  etc., 
  second row of column1, second row of column2, etc.
)
```

### Examples


<br>
```{r showexample}
# The examples below use tibble::tibble() and tibble::tribble to recreate this tibble:
example_tibble
```


<br>

```{r example1}
# Re-create example_tibble with tibble::tibble()
tibble(
  food = c("cucumber", "pomegranate", "carrot"),
  color = c("green", "red", "orange"),
  food_type = c("fruit", "fruit", "root")
)
```

<br>

```{r example2}
# Re-create example_tibble with tibble::tribble()
# Use *spacing* to help guide yourself

tribble(
  ~food,         ~color,     ~food_type,
  "cucumber",    "green",    "fruit",
  "pomegranate", "red",      "fruit",
  "carrot",      "orange",    "root")
```

