---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(tidyr)
tibble::tribble(
  ~first_name, ~last_name, 
  "Rosalind", "Franklin", 
  "Lynn",  "Margulis",
  "Barbara", "McClintock"
) -> biologists

```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-20px;">
 <span style = "font-weight:bold;font-size:40px;"> 
  &nbsp;<code>tidyr::unite()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`unite()`](https://tidyr.tidyverse.org/reference/unite.html) function is part of the [`{tidyr}`](https://tidyr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).


We use `unite()` to combine two columns in a tibble (data frame) into a single columns. By default, column values are united with an underscore (`_`), but this can be changed with the `sep` argument.

<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{tidyr}` library, or always use the function with `tidyr::unite()` notation.

```{r description, eval=FALSE}
# Load the library
library(tidyr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
tidyr::unite()
```


### Conceptual Usage 

```{r usage, eval=FALSE}
tibble %>% 
  unite(name of new column, 
        first column to unite, 
        second column to unite)

tibble %>% 
  unite(name of new column, 
        first column to unite, 
        second column to unite, 
        sep = "optional character saying how to unite the values")
```

### Examples

The first two examples use this dataset, which contains some names of notable biologists:

```{r exdata}
biologists
```

```{r example1}
# Combine `first_name` and `last_name` into `full_name`
biologists %>% 
  unite(full_name, first_name, last_name)
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->

```{r example2}
# Combine `first_name` and `last_name` into `full_name`, and KEEP the two originals
biologists %>% 
  unite(full_name, first_name, last_name,
        remove = FALSE)
```

```{r example3}
# Combine `first_name` and `last_name` into `full_name`, 
# but separate with a space instead of underscore
biologists %>% 
  unite(full_name, first_name, last_name,
        sep = " ")
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->

```{r example4}
# Combine `first_name` and `last_name` into `full_name_reverse`, a different format
# Note the order of arguments and the separator used
biologists %>% 
  unite(full_name_reverse, last_name, first_name,
        sep = ", ")
```