---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(tidyr)
tb_cases_wide <- tidyr::table4a
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>tidyr::pivot_longer()</code>
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The [`pivot_longer()`](https://tidyr.tidyverse.org/reference/pivot_longer.html) function is part of the [`{tidyr}`](https://tidyr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

We use this function to convert a _wide_ tibble (data frame) into a _long_ tibble (data frame). 

When using `pivot_longer()`, you must specify the columns to be pivoted, which may require the use of some `{tidyselect}` helper functions. To get help with `{tidyselect}` from the `{introverse}`, use `show_topics("tidyselect")` to see available docs.

To use this function, you need to either first load the `{tidyr}` library, or always use the function with `tidyr::pivot_longer()` notation.

```{r description, eval=FALSE}
# Load the library
library(tidyr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
tidyr::pivot_longer()
```


### Conceptual Usage 

```{r usage, eval=FALSE}
tibble %>% 
  pivot_longer(columns to pivot from wide to long,
               names_to = "name of new column that will contain the old column names",
               values_to = "name of new column that will contain the values from the old columns")

```

### Examples

This example use the wide dataset `billboard`, which comes from the `{tidyr}` package and contains song rankings for the Billboard top 100 songs in the year 2000.


```{r billboard}
# Show billboard dataset
billboard
```

```{r example0}
# Use pivot_longer() and the tidyselect helper starts_with to pivot all columns that start with "wk"
billboard %>%
  pivot_longer(starts_with("wk"),
               names_to = "week", 
               values_to = "song_ranking")
```

<br>

This example example uses the dataset `tb_cases_wide`, which is a wide tibble showing how many cases of Tuberculosis were recorded in each country in the given year.

```{r exdata}
tb_cases_wide
```

```{r example1}
# Pivot tb_cases longer.
# Use backticks `` or quotes "" to refer to columns whose names start with a number
tb_cases_wide %>% 
  pivot_longer(`1999`:`2000`, # we want to pivot columns 1999 and 2000
               names_to = "year",  # this column will contain 1999 and 2000
               values_to = "number_of_tb_cases") # will contain information previously in columns 1999 and 2000
```
