---
title: "`tidyr::drop_na()`"
author: "`{introverse}` quick docs"
output: 
  html_document:
    highlight: pygments
    theme: flatly
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(tidyr)
```


## Description

The [`drop_na()`](https://tidyr.tidyverse.org/reference/drop_na.html) function is part of the [`{tidyr}`](https://tidyr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

It is used to remove rows containing `NA`s from a tibble (data frame), either from the full dataset or from certain specified column(s).

<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{tidyr}` library, or always use the function with `tidyr::drop_na()` notation.

```{r description, eval=FALSE}
# Load the library
library(tidyr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
tidyr::drop_na()
```


## Conceptual usage 

```{r usage, eval=FALSE}
# Remove all NAs from the tibble
tibble %>% 
  drop_na()

tibble %>% 
  drop_na(column to drop NAs from)

tibble %>% 
  drop_na(columns, to, drop, NAs, from)
```

## Examples

> Run `get_help("msleep")` for a refresher on the `msleep` dataset.

```{r example1}
# Remove all NAs in the dataset
msleep %>% 
  drop_na()
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->


```{r example2}
# Remove all rows with NA in the `vore` column
msleep %>% 
  drop_na(vore)
```

<br> 


```{r example3}
# Remove all rows with NA in the `vore` or `conservation` columns
msleep %>% 
  drop_na(vore, conservation)
```
