---
title: "`tidyr::pivot_wider()`"
author: "`{introverse}` quick docs"
output: 
  html_document:
    highlight: pygments
    theme: flatly
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(tidyr)
library(dplyr)
tb_cases_long <- tidyr::table1 %>% select(-population) 
```


## Description

The [`pivot_wider()`](https://tidyr.tidyverse.org/reference/pivot_wider.html) function is part of the [`{tidyr}`](https://tidyr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

We use this function to convert a _long_ tibble (data frame) into a _wide_ tibble (data frame). 

To use this function, you need to either first load the `{tidyr}` library, or always use the function with `tidyr::pivot_wider()` notation.

```{r description, eval=FALSE}
# Load the library
library(tidyr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
tidyr::pivot_wider()
```


## Conceptual usage 

```{r usage, eval=FALSE}
tibble %>% 
  pivot_wider(names_from = "column whose values will become new column names",
              values_from = "column whose values will be in those new columns")

```

## Examples


This example uses the dataset `tb_cases_long`, a long tibble which shows how many cases of Tuberculosis were recorded in each country in the given year.

```{r exdata}
tb_cases_long
```

```{r example1}
# Pivot tb_cases wider
tb_cases_long %>% 
  pivot_wider(names_from = "year", # new columns will be created from values in `year`
              values_from = "cases") # those new columns contain values from `cases`
```
