---
title: "{introverse} quick docs"
output: 
  html_document:
    highlight: tango
    theme: readable
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(dplyr)
library(tidyr)
```

```{css custom, echo=FALSE}
.title{
    display: none;
}
blockquote {
  display: block;
  font-size: 0.95em;
  border-left: 3px solid #7FB036;
  padding:10px;
  margin-left:10px;
}
```

<div>
<br>
 <img src = "logo.png" style = "float:left;height:130px;margin-top:-30px;">
 <span style = "font-weight:bold;font-size:30px;"> 
  &nbsp;<code>{tidyselect}()</code> helper functions
 </span>
 <br>
 <b>&nbsp;&nbsp;&nbsp;<code>get_help()</code> docs</b>
</div>
<br>

### Description

The  [`{tidyselect}`](https://tidyselect.r-lib.org/) package is part of the [`{tidyverse}`](https://www.tidyverse.org/), and it provides a useful set of _helper functions_ for conveniently (de-)selecting columns from tibbles (data frames). In fact, it is functions in `{tidyselect}` which enable much [`{dplyr}`](https://dplyr.tidyverse.org/) and [`{tidyr}`](https://tidyr.tidyverse.org/) functionality, including selecting columns with `dplyr::select()` or specifying columns to pivot with `tidyr::pivot_longer()`, and as a consequence all `{tidyselect}` functionality is by-definition included when you load core `{tidyverse}` packages like `{dplyr}` and `{tidyr}`. The `{introverse}` lead author, notably, prefers to call this functionality "select magic."

We generally use these helper functions as part of our _arguments_ to functions that deal with column selection, such as `dplyr::select()`, `dplyr::pull()`, or `tidyr::pivot_longer()`. Helper functions include the following:

> By adding a minus sign to all function _arguments_, you can make them _deselect_ instead of selecting the given columns

| Function | What it selects | 
|:---------|:---------------|:-----------|
| `everything()` | All columns, except those already explicitly stated |
| `contains()`     |  All columns that contain a given string argument |
| `starts_with()`     | All columns that start with a given string, i.e. prefix |
| `ends_with()`     | All columns that end with a given string, i.e. suffix |
| `last_col()`     | The last column in the tibble (data frame)
| `matches()`     |  All columns that match a regular expression (a special type of pattern-matching string) |

To use these function, you need to either first load the _`{dplyr}`_ (the `{dplyr}` library conveniently includes the `{tidyselect}` functionality!) library, or always use the function with `dplyr::()` notation.

```{r description, eval=FALSE}
# Load a core library like dplyr ot tidyr
library(dplyr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation, for example with the everything() function
dplyr::everything()
```


### Examples

The first examples below use the `msleep` dataset. Learn more about this dataset with `get_help("msleep")`.

```{r msleep}
# Show the msleep dataset with head()
head(msleep)
```

<br>

```{r example1}
# Use everything() to move the column brainwt to the front
# This code says: first select brainwt, and then _everything else_
msleep %>% 
  select(brainwt, everything())
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->


```{r example2}
# Use contains() to keep only columns that contain the letter "w"
msleep %>%
  select(contains("w"))
```

<br> 

```{r example2.5}
# Use contains() to REMOVE columns that contain the letter "w", with a minus sign
msleep %>%
  select(-contains("w"))
```

<br> 


```{r example3}
# Use starts_with() to keep only columns that start with "sleep"
msleep %>%
  select(starts_with("sleep"))
```

<br> 

```{r example3.5}
# Use starts_with() to REMOVE columns that start with "sleep", with a minus sign
msleep %>%
  select(-starts_with("sleep"))
```

<br> 

```{r example4}
# Use ends_with() to keep only columns that end with "wt"
msleep %>%
  select(ends_with("wt"))
```

<br> 

```{r example4.5}
# Use ends_with() to REMOVE columns that end with "wt", with a minus sign
msleep %>%
  select(-ends_with("sleep"))
```

<br> 

```{r example5}
# Tidyselect allows us to use colons `:` to select a range of columns
# Select vore through awake
msleep %>%
  select(vore:awake)
```

<br> 

```{r example5.25}
# DE-select vore through awake, with a minus sign
msleep %>%
  select(-(vore:awake))
```


<br>
```{r example5.5}
# Use last_col() to keep only the last-appearing column in msleep
msleep %>%
  select(last_col())
```

<br>
```{r example5.75}
# Use last_col() to keep the columns `sleep_cycle` through the last-appearing column in msleep
msleep %>%
  select(sleep_cycle:last_col())
```


<br>
```{r example6}
# ADVANCED EXAMPLE:
# Use matches() to keep only columns that match the regex `\\w+_\\w+`
msleep %>%
  select(matches("\\w+_\\w+"))
```

<br>


The remaining examples use the wide dataset `billboard`, which comes from the `{tidyr}` package and contains song rankings for the Billboard top 100 songs in the year 2000.


```{r billboard}
# Show billboard dataset
billboard
```

<br>


```{r example7}
# Use tidyr::pivot_longer and starts_with to pivot all columns that start with "wk"
billboard %>%
  pivot_longer(starts_with("wk"),
               names_to = "week", 
               values_to = "song_ranking")
```

<br>

```{r example8}
# Use ends_with to select artist, track, and columns ending in the number 5
billboard %>%
  select(artist, track, ends_with("5"))
```