---
title: "`dplyr::filter()`"
author: "`{introverse}` quick docs"
output: 
  html_document:
    theme: flatly
    highlight: pygments
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
# If your topic is from a library, load the INDIVIDUAL library here.
# DO NOT, I REPEAT __DO NOT__, LOAD THE FULL TIDYVERSE. ONLY LOAD INDIVIDUAL LIBRARIES THAT ARE NEEDED.
# IN OTHER WORDS, `library(tidyverse)` <-- ABSOLUTELY NOT!!!!!!!!!
library(dplyr)
pengs <- readr::read_csv(
  system.file("extdata", "pengs.csv", package = "introverse") 
)
```

<br>

## Description

The [`filter()`](https://dplyr.tidyverse.org/reference/filter.html) function is part of the [`{dplyr}`](https://dplyr.tidyverse.org/) package, which is part of the [`{tidyverse}`](https://www.tidyverse.org/).

It is used to **subset rows** from tibbles (data frames). Rows are kept if the given logical statement (code that gives `TRUE` or `FALSE`) is `TRUE`. Use this function if you only want to keep certain rows, aka certain observations."


<!-- Add the following if the topic is a function that does NOT come with base R, and uncomment/fill in the subsequent multiline HTML comment: -->
To use this function, you need to either first load the `{dplyr}` library, or always use the function with `dplyr::filter()` notation.

```{r, eval=FALSE}
# Load the library
library(dplyr)
# Or, load the full tidyverse:
library(tidyverse)

# Or, use :: notation
dplyr::filter()
```



## Conceptual usage 

```{r, eval=FALSE}
tibble %>% 
  filter(logical statement)

tibble %>% 
  filter(logical statement,
         another logical statement,
         and more logical statements that you want to all be TRUE)

```

## Examples

> Need a reminder of the dataset `pengs`? You can see the full dataset with `View(pengs)` and learn more about the dataset with `get_help("pengs")`.

```{r}
# Subset pengs to keep only rows (observations) that are 'Gentoo' species (rows where it is TRUE that species == 'Gentoo')" 
pengs %>% 
  filter(species == "Gentoo")
```

<br>  <!-- Do not delete the <br>, and make sure to keep the previous and following lines blank. -->


```{r}
# Subset pengs to keep only rows with flipper lengths greater than 200
pengs %>%
  filter(flipper_length_mm > 200)
```

<br> 


```{r}
# Subset pengs to keep only rows where the species is 'Gentoo' and the island is 'Biscoe'
pengs %>%
  filter(species == "Gentoo", island == "Biscoe")
```

<br> 

```{r}
# Subset pengs to keep only rows where the species is _either_ 'Gentoo' or 'Adelie'
# In other words, it should be TRUE that the species is %in% the array c("Gentoo", "Adelie"). 
# Remember: We do _not_ use `==` when asking if a value is in an array.
pengs %>%
  filter(species %in% c("Gentoo", "Adelie"))
```
