---
title: "Stringr module"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(stringr)
knitr::opts_chunk$set(echo = FALSE)
```


## Using stringr

### Exercise 


Write the R code required to count the length of the string "abcde".

```{r count-length, exercise=TRUE}

```

```{r count-length-solution}
str_length("abcde")
```

Write the R code required to count the length of the string "abcde fg" What do you notice about including a space in the string?

```{r count-length-2, exercise=TRUE}

```

```{r count-length-2-solution}
str_length("abcde fg")
```

Write the R code required to count the length of the string "how long is this string?"

```{r count-length-3, exercise=TRUE}

```

```{r count-length-3-solution}
str_length("how long is this string")
```

Write the R code required to count the length of the string "12.98". What do you notice about the use of punctuation in the string?

```{r count-length-4, exercise=TRUE}

```

```{r count-length-4-solution}
str_length("12.98")
```

Write the R code required to count the length of the string "24 hours in a day".

```{r count-length-5, exercise=TRUE}

```

```{r count-length-5-solution}
str_length("24 hours in a day")
```


Write the R code required to turn the sentence "i love coding" to all uppercase.

> Hint: save your string to a variable first.

```{r turn-all-uppercase, exercise=TRUE}

```

```{r turn-all-uppercase-solution}
"I love coding" -> x
str_to_upper(x)
```

Write the R code required to turn the sentence "CODING IS AMAZING!!" to all lowercase.

```{r turn-all-lowercase, exercise=TRUE}

```


```{r turn-all-lowercase-solution}
"CODING IS AMAZING!!" -> x
str_to_lower(x)
```

Write the R code required to write the sentence "I LOVE CODING!" in all lowercase, using the Turkish locale-sensitivity. 

```{r turn-all-lowercase-tr, exercise=TRUE}

```

```{r turn-all-lowercase-tr-solution}
"I LOVE CODING!" -> x
str_to_lower(x, "tr")
```

Write the R code required to write the phrase "uppercase fun" in all uppercase.

```{r uppercase-fun, exercise=TRUE}

```

```{r uppercase-fun-solution}
"uppercase fun" -> x
str_to_upper(x)
```

Write the R code required to write the phrase "LOWERCASE FUN" in all lowercase.

```{r lowercase-fun, exercise=TRUE}

```


```{r lowercase-fun-solution}
"LOWERCASE FUN" -> x
str_to_lower(x)
```

Write the R code required to order "w, p, r" by the correct number.

> Hint: save your string to a variable

```{r order-by-number, exercise=TRUE}

```

```{r order-by-number-solution}
c("w", "p", "r") -> x
str_order(x)
```

Write the R code required to order "w, p, r" in the correct order.

```{r order-by-number-2, exercise=TRUE}

```

```{r order-by-number-3-solution}
c("w", "p", "r") -> x
str_sort(x)
```

Write the R code required to order "3, 7, 2" in the correct order.

```{r order-by-number-4, exercise=TRUE}

```

```{r order-by-number-4-solution}
c("3", "7", "2") -> x
str_sort(x)
```

Add extra whitespace to both sides of the words "coding" "fun"

```{r whitespace-1, exercise=TRUE}

```


```{r whitespace-1-solution}
c("coding", "fun") -> x
str_pad(x, 10, "both")
```


Wrap the poem "Shall I Compare Thee to a Summer's Day" by William Shakespeare so that the length of the lines are as similar as possible.

"Shall I compare thee to a summer’s day?
Thou art more lovely and more temperate.
Rough winds do shake the darling buds of May,
And summer’s lease hath all too short a date.


```{r wrap, exercise=TRUE}

```

```{r wrap-solution}
poem <-str_c(" 'Shall I compare thee to a summer's day? Thou art more lovely and more temperate. Rough winds do shake their darling buds of May , and summer's lease hath all too short a date.'")

cat(str_wrap(poem, width = 40))
```

Write the R code required to split "1-2-3-4-5" into a fixed number of pieces.

> Hint: remember to include the symbols your numbers are divided by in your code.

```{r split, exercise=TRUE}

```

```{r split-solution}
str_split("1-2-3-4-5", "-")
```

Write the R code required to split "1/2/3/4/5" into a fixed number of pieces.

```{r split-2, exercise=TRUE}

```

```{r split-2-solution}
str_split("1/2/3/4/5", "/")
```

Write the R code required to split "1-2-3-4-5" into 2 fixed pieces.

```{r split-3, exercise=TRUE}

```

```{r split-3-solution}
str_split_fixed("1-2-3-4-5", "-", n=2)
```

Write the R code required to split "c-o-d-i-n-g" into 3 fixed pieces.

```{r split-4, exercise=TRUE}

```


```{r split-4-solution}
str_split_fixed("c-o-d-i-n-g", "-", n=3)
```

Use the function `fixed()` to detect if the sequences "hello" and "hello" match.

```{r fixed, exercise=TRUE}

```

```{r fixed-solution}
x1 <- "hello"
x2 <- "hello"
c(x1, x2)
x1==x2
```

Use the function `fixed()` to detect if the sequences "hello:)" and "hello" match.
```{r fixed-2, exercise=TRUE}

```

```{r fixed-2-solution}
x1 <- "hello:)"
x2 <- "hello"
c(x1, x2)
x1==x2
```

Use the function `fixed()` to detect if the sequences "HellO" and "hello" match.

```{r fixed-3, exercise=TRUE}

```

```{r fixed-3-solution}
x1 <- "HellO"
x2 <- "hello"
c(x1, x2)
x1==x2
```

Write the R code to define "I love coding" to a variable, then split each character individually.

```{r split-sentence-by-characters, exercise=TRUE}

```

```{r split-sentence-by-characters-solution}
"I love coding" -> x
str_split(x, "")
```

Write the R code required to count how many characters are present in the phrase "I love coding" after f-rst diving it into individual characters.

```{r split-sentence-by-characters-count, exercise=TRUE}

```

```{r split-sentence-by-characters-count-solution}
"I love coding" -> x
str_split(x, "")
str_count(x, "")
```

Use `boundary()` to split the sentence "Coding is so much fun." into separate words. 

```{r split-sentence-by-words, exercise=TRUE}

```

```{r split-sentence-by-words-solution}
"Coding is so much fun" -> x
str_split(x, boundary("word"))
``` 

Use `boundary()` and another stringr function learned earler to split the sentence "We are learning stringr." into separate words. Next, count how many words are present.

```{r split-sentence-by-words-count, exercise=TRUE}

```

```{r split-sentence-by-words-count-solution}
"We are learning stringr" -> x
str_split(x, boundary("word"))
str_count(x, boundary("word"))
``` 




