---
title: "Question Bank Set 1"
author: "Shemaiah Sotrah"
date: "6/24/2020"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(learnr)
```


### Set 1: Using R as a calculator

1. What is 5 times 6?

```{r calc_q1, exercise =TRUE}

```
  
<div id="calc_q1-hint">
**Hint:**  Use the asterisk `*` to multiply numbers
</div>

2. What is 600 divided by 2.53?
 
``` {r calc_q2, exercise = TRUE}

```

<div id="calc_q2-hint">
**Hint:**  Use the forward slash `/` to divide numbers
</div>

3. What is 138 squared?
  
``` {r calc_q3, exercise = TRUE}

```

<div id="calc_q3-hint">
**Hint:**  Use the carot symbol `^` *or* two asterisks (`**`) to raise numbers to an exponent.
</div>

4. What is the square root of 14?
  
``` {r calc_q4, exercise = TRUE}

```  

<div id="calc_q4-hint">
**Hint:**   Square roots are like raising to the power of 0.5
</div>

### Set 1: Answer Key
``` {r, calc_q1-answer}
#Solution to Question 1
5*6
```

``` {r, calc_q2-answer}
#Solution to Question 2
600/2.53
```

``` {r, calc_q3-answer}
#Solution to Question 3
138**2 
#Equally valid approach
138^2 
```

``` {r, calc_q4-answer}
14^0.5
```


### Set 2: Working with logic: TRUE and FALSE

1. Use a *logical operator* to ask if 56 is greater than 7.
  
``` {r logical_q1, exercise = TRUE}

```  

<div id="logical_q1-hint">
**Hint:**  Logical operators include: `>`, `<`, `>=`, `<=`, `==`
</div>

2. Use a *logical operator* to ask if 110 is equal to 12.
  
``` {r logical_q2, exercise = TRUE}

``` 
  
<div id="logical_q2-hint">
**Hint:**  When comparing *equality* of numbers, you MUST use two equals!! `==`
</div>

3. Use a *logical operator* to ask if 32 is less than or equal to 88.
  
``` {r logical_q3, exercise = TRUE}

``` 

<div id="logical_q3-hint">
**Hint:**  `<` is less than, but `<=` is less than *or equal to*`
</div>

4. Use a *logical operator* to ask if 100 is equal to -10 squared.
  
``` {r logical_q4, exercise = TRUE}

``` 

<div id="logical_q4-hint">
**Hint:** Don't forget to use parentheses!!
</div>

5. Use a *logical operator* to ask if 100 is *not equal* to -55. 
  
``` {r logical_q5, exercise = TRUE}

``` 

<div id="logical_q5-hint">
**Hint:** Check if values are *not equal* with the logical operator `!=`. Generally speaking, exclamation marks negate logical statements.
</div>

6. Use a *logical operator* to ask if 100 is *not less than* 88. 
 
``` {r logical_q6, exercise = TRUE}

``` 

<div id="logical_q6-hint">
**Hint:** *Negate* any other logical operator (besides `==`, which becomes `!=`), by *enclosing the statement inside parentheses prefixed by an exclamation mark, e.g. `!(...statement...)`
</div>

7. In all questions above, you wrote *single logical statements* - you used a logical operator to evaluate if a single thing was TRUE or FALSE. This time, write a *compound logical statement* to ask if: 100 is equal to 100 *and* 66 is equal to 66.

``` {r logical_q7, exercise = TRUE}

``` 

```{r logical_q7-hint-1}
#Combine statements with *and* using the ampersand `&`
```

```{r logical_q7-hint-2}
#Statements with "and" will only evaluate as TRUE if *both* conditions are TRUE. `
```

8. Write a *compound logical statement* to ask if 100 is greater than to 87 *and* 66 is less than 12.
  
``` {r logical_q8, exercise = TRUE}

``` 

<div id="logical_q8-hint">
**Hint:**  Combine statements with *and* using the ampersand `&`
</div>

9. Similar to the previous question, write a *compound logical statement* to ask if 100 is greater than to 87 *OR* 66 is less than 12.
  
``` {r logical_q9, exercise = TRUE}

``` 

```{r logical_q9-hint-1}
#Combine statements with *or* using the "pipe operator" `|`. This is not the letter L or I - this symbol is located on the backslash `\` key. Press Shift + `\` to get it.
```

```{r logical_q9-hint-2}
#Statements with "or" will evaluate as TRUE if *at least one* condition is TRUE. `
```

### Set 2: Answer Key
``` {r, logical_q1-answer}
#Solution to Question 1
56 > 7
```

``` {r, logical_q2-answer}
#Solution to Question 2
110 == 12
```

``` {r, logical_3-answer}
#Solution to Question 3
32 <= 88
```

``` {r, logical_q4-answer}
#Solution to Question 4
100 == (-10)^2
```

``` {r, logical_q5-answer}
#Solution to Question 5
100 != -55
```

``` {r, logical_q6-answer}
#Solution to Question 6
!(100 < 88)
```

``` {r, logical_q7-answer}
#Solution to Question 7
100 == 100 & 66 == 66
```

``` {r, logical_q8-answer}
#Solution to Question 8
100 > 87  & 66 < 12
```

``` {r, logical_q9-answer}
#Solution to Question 9
100 > 87 | 66 < 12
```

### Set 3: Defining and understanding variables in R

1. The code example below defines a variable. Answer the multiple choice questions about this code:
```{r}
y <- 8.5
```

```{r quiz1, echo=FALSE}
quiz(caption = "Multiple Choice",
  question("What is the *name* of this variable?",
     answer("`<-`"),
     answer("y", correct = TRUE),
     answer("8.5"),
     answer("We can't tell.")
     ),
  question("What is the *value* of this variable?",
     answer("`<-`"),
     answer("y"),
     answer("8.5", correct = TRUE),
     answer("We can't tell.")
     ),
  question("What is the *type* of this variable?",
     answer("Character"),
     answer("Numeric", correct = TRUE),
     answer("Logical"),
     answer("We can't tell."
     )
  )
)
```

<!-- Shemaiah: for these questions, include the option:   exercise.lines=8     -->

2. Using the code presented in question 1 to help guide you, define a variable named `banana_price` whose value is 0.55 (aka 55 cents). 

``` {r variables_q2, exercise = TRUE, exercise.lines=8}

``` 

<div id="variables_q2-hint">
**Hint:**  Use the assignment operator `<-` to assign the value 0.55 to the variable name `banana_price`
</div>

3. Again, define the `banana_price` variable (since these questions don't communicate with each other!!). Use R as a calculator to determine the cost of 17 bananas.

``` {r variables_q3, exercise = TRUE, exercise.lines=8}

``` 
  
<div id="variables_q3-hint">
**Hint:**  Perform this operation using the `banana_price` variable itself.
</div>

4. Again, define the `banana_price` variable. Define a second variable called `num_bananas` with the value for how many bananas you're buying. Using both your variables, use R as a calculator to calculate the total cost.

``` {r variables_q4, exercise = TRUE, exercise.lines=8}

``` 
  
<div id="variables_q4-hint">
**Hint:**  We are learning to use variables whenever possible instead of the direct numbers. Using values in calculations rather than variables representing those values is called **hard-coding** and can lead to bugs!
</div>

5. Again, define your two variables, `banana_price` and `num_bananas`. Again calculate the total cost using these variables, *but this time*, save the output to a new variable called `total_cost`. Output the final value of `total_cost` by simply typing the variable name on its own line.
  
``` {r variables_q5, exercise = TRUE, exercise.lines=8}

``` 

<!-- Shemaiah: make sure to KEEP THE \ in front of $. -->
6. Let's work with *logical operations* and variables. Imagine you have \$5 to spend at the grocery store. Can you afford to buy 30 bananas (you are apparently making *a lot* of quarantine banana bread)? To answer this question, first *define* four variables (`banana_price`, `num_bananas`[ but what's the value this time?!], `total_cost`, and `total_money` which represents your budget). Then, use logical operations to compare `total_cost` and `total_money`. You should arrive at an answer of TRUE *if it is affordable*, and FALSE if you can't buy this many bananas.
   
``` {r variables_q6, exercise = TRUE, exercise.lines=8}

``` 

<div id="variables_q6-hint">
**Hint:**  A TRUE solution is one where `total_money` is *greater than or equal to* the `total_cost`. Or, where `total_cost` is *less than or equal to* `total_money`.
</div>

7. The grocery store visit is getting more interesting. Sure, banana bread is all well and good, but it's certainly not as good as chocolate-chip banana bread. You've recently come into some money, so you now have \$20 to spend!! You want to buy 4 bags of chocolate chips that cost \$1.50 each as well as 16 bananas which, due to some evil price-gouging, now cost 60 cents (\$0.60) each. Can you afford this purchase? 
   
``` {r variables_q7, exercise = TRUE, exercise.lines=8}

``` 

```{r variables_q7-hint-1}
#Make sure to define all variables with *meaningful variable names* first, *then* perform calculations to determine the `total_cost`, and then compare `total_cost` to `total_money`. In the end, you will have created six variables.
```

```{r variables_q7-hint-2}
#Don't put spaces into your variable names *ever*. Instead, use underscores to represent spaces.
```

### Set 3: Answer Key

``` {r, variables_q2-answer}
#Solution to Question 2
 banana_price <- 0.55
```

``` {r, variables_q3-answer}
#Solution to Question 3
banana_price <- 0.55
banana_price * 17
```

``` {r, variables_q4-answer}
#Solution to Question 4
banana_price <- 0.55
num_bananas <- 17
banana_price * num_bananas
```

``` {r, variables_q5-answer}
#Solution to Question 5
banana_price <- 0.55
num_bananas <- 17
total_cost <- banana_price * num_bananas
total_cost
```

``` {r, variables_q6-answer}
#Solution to Question 6
banana_price <- 0.55
num_bananas <- 30
total_money <- 5
total_cost <- banana_price * num_bananas
total_cost <= total_money
```

``` {r, variables_q7-answer}
#Solution to Question 7
banana_price <- 0.6
num_bananas <- 16
choc_price <- 1.5
num_choc  <- 4
total_money <- 20
total_cost <- (banana_price * num_bananas) + (choc_price * num_choc)
total_cost <= total_money
```


