---
title: "Question Bank Set 2"
author: "Shemaiah Sotrah"
date: "6/24/2020"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(learnr)
```


### Set 1: Working with functions in R

1. Use the function `sqrt()` to calculate the square root of 206.
  
``` {r func_q1, exercise = TRUE}

```
  
<div id= "func_q1-hint">
**Hint:** Provide 206 as an *argument* to the function `sqrt()`
</div>

2. Use the function `log()` to calculate the *natural log* (this is actually the default behavior of this function) of 206.
  
``` {r func_q2, exercise = TRUE}

```
  
<!-- This question's chunk should have the ?log in it --> 
3. Looking up documentation for how to use functions in R (and other languages!) is *very important* - don't try to be a psychic!! To see the documentation for a function, type `?nameoffunction`. To see this in action, run the code below. Have a particular look at the sections "Usage" and "Arguments." Then, use the function `log()` to calculate the log in base 10 of 206.
 
``` {r func_q3, exercise = TRUE}
?log()

```  

<div id= "func_q3-hint">
**Hint:** The function `log()` can optionally take a *second* argument specifying the base. For example, `log(500, 6)` takes the log of 500 in base 6. Make sure you see where this information is described in the `?log` output!
</div>

4. The function `class()` will tell you what *type* of data a given variable or value is. Use this function to determine the *type* of value 7.
  
``` {r func_q4, exercise = TRUE}

```  

<div id= "func_q4-hint">
**Hint:** Provide 7 as the argument to `class()`
</div>

5. Now, use `class()` to determine the *type* of value "7".
  
``` {r func_q5, exercise = TRUE}

```    
  
<div id= "func_q5-hint">
**Hint:** Don't forget the quotes! What did you learn from the last two questions about how quotes work?
</div>

6. Let's combine our skills! Define a variable called `x` whose value is 7 (no quotes!). Then, determine its type with `class()`, but **save this output** to a new variable called `type_x`. Finally, use logical operators with this variable to ask: Is the type of the variable equal to (hint!) numeric?
  
``` {r func_q6, exercise = TRUE}

``` 

```{r func_q6-hint-1}
#Don't forget to use `==` when comparing equality of values. 
```

```{r func_q6-hint-2}
#If you receive an error, it may be because you forgot to put quotes around *"numeric"*. If you don't use quotes, R thinks you're referring to a variable called `numeric`, but no such variable exists! R doesn't want to use variables that don't exist, because....they don't exist. 
```

### Set 1: Answer Key

``` {r, func_q1-answer}
#Solution to Question 1
sqrt(206)
```

``` {r, func_q2-answer}
#Solution to Question 2
log(206)
```

``` {r, func_q3-answer}
#Solution to Question 3
log(206, 10)
```

``` {r, func_q4-answer}
#Solution to Question 4
 class(7)
```

``` {r, func_q5-answer}
#Solution to Question 5
class("7")
```

``` {r, func_q6-answer}
#Solution to Question 6
x <- 7
type_x <- class(x)
type_x == "numeric"
```

### Set 2: Working with arrays in R

Please name these questions `array_qX` where X is the number (1, 2, 3, ...).

> Note: The terms "array" and "vector" both refer to the same thing and are used interchangeably, but a "list" is something else entirely! These exercises don't consider lists at all.


1. Define an array variable called `array_numbers` that contains the values 1, 2, 3, 4, and 5. Then, type the variable name itself to make sure its contents look right.
 
``` {r array_q1, exercise = TRUE}

```          

<div id= "array_q1-hint">
**Hint:** Use the function `c()` ("combine") to define your array
</div>

2. Define an array variable called `array_numbers2` that again contains the values 1, 2, 3, 4, and 5, **but!** this time use a colon `:` instead of `c()`. Again, type the variable name itself to make sure its contents look right.
             
``` {r array_q2, exercise = TRUE}

```  

<div id= "array_q2-hint">
**Hint:**  `5:7` defines an array with values 5, 6, 7. Can you adapt this approach to the given question
</div>

3. Define an array variable called `even_numbers` that contains the values 2, 4, 6, 8, and 10 using the `c()` function.
  
``` {r array_q3, exercise = TRUE}

```  

4. Define an array variable called `even_numbers` that contains the values 2, 4, 6, 8, and 10 using a new function (!) `seq()`. The chunk below shows an example of using `seq()`. Run this code to see the function in action, and then *adapt* the code to make the `even_numbers` array.

```{r array_q4, exercise = TRUE}
seq(1, 34, by = 3)

```    

```{r array_q4-hint-1}
#You may want to run `?seq` to see more about how the function works!
```

```{r array_q4-hint-2}
#Always print out your output to be sure the variable you've created is correct. Keep trying until you get there - coding is all about trial and error! 
```

5. The function `length()` tells you how many items are in a given array. Define an array using `seq()` as the multiples of 3 from 0-999 (i.e. 0, 3, 6, 9, ..., 993, 996, 999) and then use `length` to determine how many numbers it contains.
  
```{r array_q5, exercise = TRUE}

```    

<div id= "array_q5-hint">
**Hint:** First make sure (print print print!) that you are defining this array correctly. *Only once* you can see that the array is correct should you calculate its length.
</div>

<!-- a lot of questions coming up should be pre-populated with baseline chunk code--> 
6. One place where R really shines is *vectorization* - it can perform an operation on *all items in a vector* at the same time! The code below multiplies every value in the array `example_array` by 5. Run this code and examine its output to be sure you understand it. Then, modify the code to instead divide every item in the array by 10.
  
```{r array_q6, exercise = TRUE}
example_array <- c(5, 78, -3, 11, 6, 0)
example_array * 5
```      

<div id= "array_q6-hint">
**Hint:** When modifying the code be careful not to delete the definition of `example_array`.
</div>

7. We can also perform logical operations in a vectorized context. The code below asks if all values in `example_array` are greater than 0. Run this code and examine its output to be sure you understand it. Then, modify the code to instead ask if every item is *less than or equal to 0**.
  
```{r array_q7, exercise = TRUE}
example_array <- c(5, 78, -3, 1, 6, 0)
example_array > 0
```  

```{r array_q7-hint-1}
#When modifying the code be careful not to delete the definition of `example_array`.
```

```{r array_q7-hint-2}
#This type of output would be referred to as a *logical array*.
```

8. In fact, logical values TRUE and FALSE evaluate to numbers 1 and 0, respectively. Use the function `sum()` (adds up values in an array) to calculate the sum of the output from `example_array > 0`. 
              
```{r array_q8, exercise = TRUE}

```                 

```{r array_q8-hint-1}
#Don't forget to define `example_array` again here.
```

```{r array_q8-hint-2}
#You can provide the full expression `example_array > 0` to the function `sum()`, or save that logical expression's output to a variable, and use the variable as the argument to `sum()`.
```

### Set 2: Answer Key

``` {r, array_q1-answer}
#Solution to Question 1
array_numbers <- c(1,2,3,4,5)
array_numbers
```

``` {r, array_q2-answer}
#Solution to Question 2
array_numbers2 <- 1:5
array_numbers2
```

``` {r, array_q3-answer}
#Solution to Question 3
even_numbers <- c(2,4,6,8,10)
even_numbers
```

``` {r, array_q4-answer}
#Solution to Question 4
even_numbers <- seq(2, 10, 2)
even_numbers
```

``` {r, array_q5-answer}
#Solution to Question 5
mul3 <- seq(0, 999, 3)
length(mul3)
```

``` {r, array_q6-answer}
#Solution to Question 6
example_array <- c(5, 78, -3, 11, 6, 0)
example_array / 10
```

``` {r, array_q7-answer}
#Solution to Question 7
example_array <- c(5, 78, -3, 1, 6, 0)
example_array <= 0
```

``` {r, array_q8-answer}
#Solution to Question 8
example_array <- c(5, 78, -3, 1, 6, 0)
sum(example_array > 0)
```

### Set 3: Working with data frames

This set of exercises will use a data frame (aka "tibble" - we will learn the distinction between these terms!) called `msleep`, which is built-in to R. This dataset contains information about some species of mammals as follows:

+ `name`: The common name
+ `genus`: The taxonomic genus
+ `vore`: Dietary strategy (carnivore, omnivore or herbivore)
+ `order`: The taxonomic order
+ `conservation`: The species' conservation status
+ `sleep_total`: Total amount of sleep, in hours, the species tends to experience
+ `sleep_rem`: Total amount of REM sleep, in hours, the species tends to experience
+ `sleep_cycle`: Length of sleep cycle, in hours, the species tends to experience
+ `awake`: Amount of time spent awake, in hours, the species tends to experience
+ `brainwt`: Average brain weight, in kilograms, for this species
+ `bodywt`: Average body weight, in kilograms, for this species


First, in the code chunk below, examine the data frame using functions we have seen `head()`, `nrow()`, and `ncol()` to get an overall sense of the data frame itself. Then,  answer the following multiple choice questions about this data frame.
```{r examine-msleep, exercise=T, exercise.lines = 5}



```
<!-- the above has no solution - just a chunk for them to practice and answer subsequent multiple choice -->

1. 

```{r msleep_q1, echo=FALSE}
question("How many rows are in this data frame?",
         answer("6"),
         answer("83", correct = TRUE),
         answer("11"),
         answer("We can't tell")
         )
```

2. 

```{r msleep_q2, echo=FALSE}
question("How many columns are in this data frame?",
         answer("6"),
         answer("83"),
         answer("11", correct = TRUE),
         answer("We can't tell")
         )
```

3. 

```{r msleep_q3, echo=FALSE}
question("How many columns in the data frame contain *numeric* values?",
         answer("6", correct = TRUE),
         answer("5"),
         answer("All columns are numeric"),
         answer("No columns are numeric")
         )
```

4. 

```{r msleep_q4, echo=FALSE}
question("How many columns in the data frame contain *character* (string) values?",
         answer("6"),
         answer("5", correct = TRUE),
         answer("All columns are numeric"),
         answer("No columns are numeric")
         )
```

5. 

```{r msleep_q5, echo=FALSE}
question("How many columns in the data frame contain *logical* (TRUE or FALSE) values?",
         answer("6"),
         answer("5"),
         answer("All columns are logical"),
         answer("No columns are logical", correct = TRUE)
         )
```

<!-- Now back to code questions -->

6. Technically, columns in data frames are actually *named arrays*. We can access them to work with directly using a dollar sign `$`. The code below extracts the `name` variable from the data frame. Run and understand this code, and then modify it to extract the `awake` column.

```{r msleep_q6, exercise = TRUE}
msleep$name
``` 

<div id= "msleep_q6-hint">
**Hint:**  Make sure to *not* use any quotes.
</div>

7. Finally, let's get to everyone's favorite part - statistics! Below shows you how to calculate the *average* (mean!) hours that these mammals are awake for. Try out these other functions to calculate other summary statistics: `median()`, `range()`, `min()`, `max()`, `sd()` (standard d
```{r msleep_q7, exercise = TRUE}
mean(msleep$awake)
``` 

8. Using your new skills, calculate the _median_ hours of REM sleep (columns `sleep_rem`) these mammals get. The output might not look as expected!! To start to understand why, view the full contents of this column (i.e. `msleep$sleep_rem`). You will see some values "NA", meaning there is *missing data*. Missing data is very common - real life science (especially biology!) is tricky, and sometimes we can't gather as much data as we want. 

```{r msleep_q8, exercise = TRUE}

``` 

9. So, how do we calculate the median from the question above? We need to include the *second argument* `na.rm=TRUE` which tells median to ignore ("rm" for remove) NA's when performing calculations. See this code in action below, and then modify it to calculate the _mean_ of this column instead. 

```{r msleep_q9, exercise = TRUE}
 median(msleep$sleep_rem, na.rm = TRUE)
``` 

10. The column `vore` is a *character* column - it contains strings rather than numbers. Let's examine this column using the function `unique()`, which returns all unique values in a given array - What are the unique vores in this dataset?

```{r msleep_q10, exercise = TRUE}

``` 

<div id= "msleep_q10-hint">
**Hint:** Remember, you can always look up how to use a function! Try running `?unique` if you're stuck.
</div>

### Set 3: Answer Key

``` {r, msleep_q6-answer}
#Solution to Question 6
msleep$awake
```

``` {r, msleep_q7-answer}
#Solution to Question 7
median(msleep$awake)
range(msleep$awake)
min(msleep$awake)
max(msleep$awake)
sd(msleep$awake)
```

``` {r, msleep_q8-answer}
#Solution to Question 8
median(msleep$sleep_rem)
```

``` {r, msleep_q9-answer}
#Solution to Question 9
mean(msleep$sleep_rem, na.rm = TRUE)
```

``` {r, msleep_q10-answer}
#Solution to Question 10
unique(msleep$vore)
```
