---
title: "Introduction to R"
author: "Stephanie J. Spielman"
output: 
  learnr::tutorial:
    theme: united
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(learnr)
library(ggplot2) ## for msleep dataset
```

# CB2R Data Science Workshop, Summer 2020
<br><br><br>

## Set 1: Using R as a calculator
<br><br>

<br>

#### 1. What is 5 times 6?

```{r calc_q1, exercise =TRUE}

```
  
<div id="calc_q1-hint">
**Hint:**  Use the asterisk `*` to multiply numbers
</div>

<br>

#### 2. What is 600 divided by 2.53?
 
```{r calc_q2, exercise = TRUE}

```

<div id="calc_q2-hint">
**Hint:**  Use the forward slash `/` to divide numbers
</div>

<br>

#### 3. What is 138 squared?
  
```{r calc_q3, exercise = TRUE}

```

<div id="calc_q3-hint">
**Hint:**  Use the carot symbol `^` *or* two asterisks (`**`) to raise numbers to an exponent.
</div>

<br>

#### 4. What is the square root of 14?
  
```{r calc_q4, exercise = TRUE}

``` 

<div id="calc_q4-hint">
**Hint:**  Square roots are like raising to the power of 0.5
</div>


## Set 2: Working with logical values
<br><br>


Logical operators:

| Symbol | Asks if... | 
|:--------|:------------|
|`==`    | Two values are equal |
|`!=`    | Two values are *not* equal |
| `>`    | LHS (left-hand side) is greater than RHS (right-hand side) |
| `>=`   | LHS is greater than or equal to RHS | 
| `<`    | LHS is less than RHS  |
| `<=`   | LHS is less than or equal to RHS | 
| `&`    | Used to combine logical statements and ask if **BOTH** are `TRUE`
| `|`    | Used to combine logical statements and ask if **AT LEAST ONE** is `TRUE`. Type this symbol with `Shift`+`\`.

<br>

#### 1. Use a *logical operator* to ask if 56 is greater than 7.
  
```{r logical_q1, exercise = TRUE}

``` 

<br>

#### 2. Use a *logical operator* to ask if 110 is equal to 12.
  
```{r logical_q2, exercise = TRUE}

```
  
<div id="logical_q2-hint">
**Hint:**  When comparing *equality* of numbers, you MUST use two equals!! `==`
</div>

<br>

#### 3. Use a *logical operator* to ask if 32 is less than or equal to 88.
  
```{r logical_q3, exercise = TRUE}

```

<br>

#### 4. Use a *logical operator* to ask if 100 is equal to -10 squared.
  
```{r logical_q4, exercise = TRUE}

```

<div id="logical_q4-hint">
**Hint:** Don't forget to use parentheses!! Otherwise, your answer will not be right.
</div>

<br>

#### 5. Use a *logical operator* to ask if 100 is *not equal* to -55. 
  
```{r logical_q5, exercise = TRUE}

```

<div id="logical_q5-hint">
**Hint:** Check if values are *not equal* with the logical operator `!=`. Generally speaking, exclamation marks negate logical statements.
</div>

<br>

#### 6. Use a *logical operator* to ask if 100 is *not less than* 88. *Negating* the `<` operator, and all other operators besides `==`, you need to *enclose* the whole statement inside parentheses prefixed by an exclamation mark, like: `!(...statement...)` (where `...statement...` is whatever you are testing).
 
```{r logical_q6, exercise = TRUE}

```


<br>

#### 7. In all questions above, you wrote *single logical statements* - you used a logical operator to evaluate if a single thing was TRUE or FALSE. This time, write a *compound logical statement* to ask if: 100 is equal to 100 *and* 66 is equal to 66.

```{r logical_q7, exercise = TRUE}

```


```{r logical_q7-hint-1}
# Combine statements with *and* using the ampersand `&`
```

```{r logical_q7-hint-2}
# Statements with "and" will only evaluate as TRUE if *both* conditions are TRUE. `
```

<br>

#### 8. Write a *compound logical statement* to ask if 100 is greater than 87 *and* 66 is less than 12.
  
```{r logical_q8, exercise = TRUE}

```

<div id="logical_q8-hint">
**Hint:**  Combine statements with *and* using the ampersand `&`
</div>

<br>

#### 9. Similar to the previous question, write a *compound logical statement* to ask if 100 is greater than to 87 *OR* 66 is less than 12.
  
```{r logical_q9, exercise = TRUE}

```

```{r logical_q9-hint-1}
# Combine statements with *or* using the "pipe operator" `|`. This is not the letter L or I - this symbol is located on the backslash `\` key. Press Shift + `\` to get it.
```

```{r logical_q9-hint-2}
# Statements with "or" will evaluate as TRUE if *at least one* condition is TRUE. `
```

## Set 3: Defining variables in R
<br><br>


<br>

#### 1. The code example below defines a variable. Answer the multiple choice questions about this code:
```{r}
y <- 8.5
```

```{r quiz1, echo=FALSE}
quiz(caption = "Multiple Choice",
  question("What is the *name* of this variable?",
     answer("`<-`", message = "This is the assignment operator."),
     answer("y", correct = TRUE),
     answer("8.5"),
     answer("We can't tell."),
     allow_retry = TRUE
     ),
  question("What is the *value* of this variable?",
     answer("`<-`",  message = "This is the assignment operator."),
     answer("y"),
     answer("8.5", correct = TRUE),
     answer("We can't tell."),
     allow_retry = TRUE
     ),
  question("What is the *type* of this variable?",
     answer("Character"),
     answer("Numeric", correct = TRUE),
     answer("Logical"),
     answer("We can't tell."),
     allow_retry = TRUE
  )
)
```

<br>

> **Before you begin, an important message!** After you define a variable, **always** print out the variable to make sure it was defined correctly! In these exercises, you can reveal the value of a variable simply by typing and running the variable's name. Try it out below! First, run the code "as is" - this will define y as equal to 8.5 internally, but it won't do anything else! Then, type y directly to print its output:

```{r demo_print, exercise=TRUE}
## Gives no output on its own, but does something *internally*. 
y <- 8.5 

## Next, type y below and re-run this code chunk. The value 8.5 will be printed out
```

<br>

#### 2. Using the code presented in question 1 to help guide you, define a variable named `banana_price` whose value is 0.55 (aka 55 cents). 

```{r variables_q2, exercise = TRUE, exercise.lines=8}

```

<div id="variables_q2-hint">
**Hint:**  Use the assignment operator `<-` to assign the value 0.55 to the variable name `banana_price`
</div>

<br>

#### 3. Again, define the `banana_price` variable (since these questions don't communicate with each other!!). Use R as a calculator to determine the cost of 17 bananas.

```{r variables_q3, exercise = TRUE, exercise.lines=8}

```
  
<div id="variables_q3-hint">
**Hint:**  Perform this operation using the `banana_price` variable itself.
</div>

<br>

#### 4. Again, define the `banana_price` variable. Define a second variable called `num_bananas` with the value for how many bananas you're buying. Using both your variables, use R as a calculator to calculate the total cost.

```{r variables_q4, exercise = TRUE, exercise.lines=8}

```
  
<div id="variables_q4-hint">
**Hint:**  We are learning to use variables whenever possible instead of the direct numbers. Using values in calculations rather than variables representing those values is called **hard-coding** and can lead to bugs!
</div>

<br>

#### 5. Again, define your two variables, `banana_price` and `num_bananas`. Again calculate the total cost using these variables, *but this time*, save the output to a new variable called `total_cost`. 
  
```{r variables_q5, exercise = TRUE, exercise.lines=8}

```

<br>

#### 6. Let's work with *logical operations* and variables. Imagine you have \$5 to spend at the grocery store. Can you afford to buy 30 bananas (you are apparently making *a lot* of quarantine banana bread)? 
To answer this question, first *define* four variables (`banana_price`, `num_bananas`[ but what's the value this time?!], `total_cost`, and `total_money` which represents your budget). Then, use logical operations to compare `total_cost` and `total_money`. You should arrive at an answer of TRUE *if it is affordable*, and FALSE if you can't buy this many bananas.
   
```{r variables_q6, exercise = TRUE, exercise.lines=8}

```

<div id="variables_q6-hint">
**Hint:**  A `TRUE` solution is one where `total_money` is *greater than or equal to* the `total_cost`. Or, where `total_cost` is *less than or equal to* `total_money`.
</div>

<br>

#### 7. The grocery store visit is getting more interesting. Sure, banana bread is all well and good, but it's certainly not as good as chocolate-chip banana bread. You've recently come into some money, so you now have \$20 to spend!! You want to buy 4 bags of chocolate chips that cost \$1.50 each as well as 16 bananas which, due to some evil price-gouging, now cost 60 cents (\$0.60) each. Can you afford this purchase? 
   
```{r variables_q7, exercise = TRUE, exercise.lines=8}

```

<div id= "variables_q7-hint">
**Hint:** Make sure to define all variables with *meaningful variable names* first, *then* perform calculations to determine the `total_cost`, and then compare `total_cost` to `total_money`. In the end, you will have created six variables. **BUT!!** Make sure to *never ever EVER* put spaces into your variable names. Instead, use underscores to represent spaces.
</div>

## Set 4: Working with functions
<br><br>


<br>

#### 1. Use the function `sqrt()` to calculate the square root of 206.
  
```{r func_q1, exercise = TRUE}

```
  
<div id= "func_q1-hint">
**Hint:** Provide 206 as an *argument* to the function `sqrt()`
</div>

<br>

#### 2. Use the function `log()` to calculate the *natural log* (this is actually the default behavior of this function) of 206.
  
```{r func_q2, exercise = TRUE}

```
  
<br>

#### 3. Looking up documentation for how to use functions in R (and other languages!) is *very important* - don't try to be a psychic!! To see the documentation for a function, type `?nameoffunction`. To see this in action, run the code below. Have a particular look at the sections "Usage" and "Arguments." Then, use the function `log()` to calculate the log in base 10 of 206.
 
```{r func_q3, exercise = TRUE}
?log()

``` 

<div id= "func_q3-hint">
**Hint:** The function `log()` can optionally take a *second* argument specifying the base. For example, `log(500, 6)` takes the log of 500 in base 6. Make sure you see where this information is described in the `?log` output!
</div>

<br>

#### 4. The function `class()` will tell you what *type* of data a given variable or value is. Use this function to determine the *type* of value 7.
  
```{r func_q4, exercise = TRUE}

``` 

<div id= "func_q4-hint">
**Hint:** Provide 7 as the argument to `class()`
</div>

<br>

#### 5. Now, use `class()` to determine the *type* of value "7".
  
```{r func_q5, exercise = TRUE}

```   
  
<div id= "func_q5-hint">
**Hint:** Don't forget the quotes! What did you learn from the last two questions about how quotes work?
</div>

<br>

#### 6. Let's combine our skills! Define a variable called `x` whose value is 7 (no quotes!). Then, determine its type with `class()`, but **save this output** to a new variable called `type_x`. Finally, use logical operators with this variable to ask: Is the type of the variable equal to numeric?
  
```{r func_q6, exercise = TRUE}

```

```{r func_q6-hint-1}
# Don't forget to use `==` when comparing equality of values. 
```

```{r func_q6-hint-2}
# If you receive an error, it may be because you forgot to put quotes around "numeric". If you don't use quotes, R thinks you're referring to a variable called `numeric`, but no such variable exists! R doesn't want to use variables that don't exist, because....they don't exist. 
```

## Set 5: Working with arrays
<br><br>
> Note: The terms "array" and "vector" both refer to the same thing and are used interchangeably, but a "list" is something else entirely! These exercises don't consider lists at all.


<br>

#### 1. Define an array variable called `array_numbers` that contains the values 1, 2, 3, 4, and 5. Then, type and run the variable name itself to make sure its contents look right.
 
```{r array_q1, exercise = TRUE}

```         

<div id= "array_q1-hint">
**Hint:** Use the function `c()` ("combine") to define your array
</div>

<br>

#### 2. Define an array variable called `array_numbers2` that again contains the values 1, 2, 3, 4, and 5, **but!** this time use a colon `:` instead of `c()`. Again, type the variable name itself to make sure its contents look right.
             
```{r array_q2, exercise = TRUE}

``` 

<div id= "array_q2-hint">
**Hint:**  `5:7` defines an array with values 5, 6, 7. Can you adapt this approach to the given question
</div>

<br>

#### 3. Define an array variable called `even_numbers` that contains the values 2, 4, 6, 8, and 10 using the `c()` function.
  
```{r array_q3, exercise = TRUE}

``` 

<br>

#### 4. Define an array variable called `even_numbers` that contains the values 2, 4, 6, 8, and 10 using a new function (!) `seq()`. The chunk below shows an example of using `seq()`. Run this code to see the function in action, and then *adapt* the code to make the `even_numbers` array.

```{r array_q4, exercise = TRUE}
seq(1, 25, by = 3)
```   

```{r array_q4-hint-1}
# You may want to run `?seq` to see more about how the function works!
```

```{r array_q4-hint-2}
# Always print out your output to be sure the variable you've created is correct. Keep trying until you get there - coding is all about trial and error! 
```

<br>

#### 5. The function `length()` tells you how many items are in a given array. Define an array using `seq()` as the multiples of 3 from 0-999 (i.e. 0, 3, 6, 9, ..., 993, 996, 999) and then use tge function `length()` to determine how many numbers it contains.
  
```{r array_q5, exercise = TRUE}

```   

<div id= "array_q5-hint">
**Hint:** First make sure (print print print!) that you are defining this array correctly. *Only once* you can see that the array is correct should you calculate its length.
</div>

<br>

#### 6. One place where R really shines is *vectorization* - it can perform an operation on *all items in a vector* at the same time! The code below multiplies every value in the array `example_array` by 5. Run this code and examine its output to be sure you understand it. Then, modify the code to instead divide every item in the array by 10.
  
```{r array_q6, exercise = TRUE}
example_array <- c(5, 78, -3, 11, 6, 0)
example_array * 5
```     

<div id= "array_q6-hint">
**Hint:** When modifying the code be careful not to delete the definition of `example_array`.
</div>

<br>

#### 7. We can also perform logical operations in a vectorized context. The code below asks if all values in `example_array` are greater than 0. Run this code and examine its output to be sure you understand it. Then, modify the code to instead ask if every item is *less than or equal to 0** (but be careful not to delete the `example_array` definition when modifying!).
  
```{r array_q7, exercise = TRUE}
example_array <- c(5, 78, -3, 1, 6, 0)
example_array > 0
``` 

<div id= "array_q7-hint">
 This type of output would be referred to as a *logical array*.
</div>

<br>

#### 8. In fact, logical values `TRUE` and `FALSE` evaluate to numbers 1 and 0, respectively. Use the function `sum()` (adds up values in an array) to calculate the sum of the output from `example_array > 0`. (Don't forget to define `example_array` again here!)
              
```{r array_q8, exercise = TRUE}

```                

<div id= "array_q8-hint">
 You can provide the full expression `example_array > 0` to the function `sum()`, or save that logical expression's output to a variable, and use the variable as the argument to `sum()`.
</div>

## Set 6: Data frames
<br><br>

This set of exercises will use a *data frame* (aka *tibble*- we will learn the distinction between these terms!) called `msleep`, which has been pre-loaded for you. This dataset contains information about some species of mammals as follows:

+ `name`: The common name
+ `genus`: The taxonomic genus
+ `vore`: Dietary strategy (carnivore, omnivore or herbivore)
+ `order`: The taxonomic order
+ `conservation`: The species' conservation status
  + `cd` --> conservation-dependent
  + `en` --> endangered
  + `lc` --> least concern
  + `nt` --> near-threatened
  + `vu` --> vulnerable    
+ `sleep_total`: Total amount of sleep, in hours, the species tends to experience
+ `sleep_rem`: Total amount of REM sleep, in hours, the species tends to experience
+ `sleep_cycle`: Length of sleep cycle, in hours, the species tends to experience
+ `awake`: Amount of time spent awake, in hours, the species tends to experience
+ `brainwt`: Average brain weight, in kilograms, for this species
+ `bodywt`: Average body weight, in kilograms, for this species


First, in the code chunk below, examine the data frame using functions we have seen `head()`, `nrow()`, and `ncol()` to get an overall sense of the data frame itself. Then,  answer the following multiple choice questions about this data frame.
```{r examine-msleep, exercise=T, exercise.lines = 8}
msleep ## Reveals the whole data frame

## Examine below with head(), nrow(), ncol()


## Bonus! Try using the function str() to examine msleep


```

<br>

#### 1. 

```{r msleep_q1, echo=FALSE}
question("How many rows are in this data frame?",
         answer("6"),
         answer("83", correct = TRUE),
         answer("11"),
         answer("We can't tell"),
         allow_retry = TRUE
         )
```

<br>

#### 2. 

```{r msleep_q2, echo=FALSE}
question("How many columns are in this data frame?",
         answer("6"),
         answer("83"),
         answer("11", correct = TRUE),
         answer("We can't tell"),
         allow_retry = TRUE
         )
```

<br>

#### 3. 

```{r msleep_q3, echo=FALSE}
question("How many columns in the data frame contain *numeric* values?",
         answer("6", correct = TRUE),
         answer("5"),
         answer("All columns are numeric"),
         answer("No columns are numeric"),
         allow_retry = TRUE
         )
```

<br>

#### 4. 

```{r msleep_q4, echo=FALSE}
question("How many columns in the data frame contain *character* (string) values?",
         answer("6"),
         answer("5", correct = TRUE),
         answer("All columns are numeric"),
         answer("No columns are numeric")
         )
```

<br>

#### 5. 

```{r msleep_q5, echo=FALSE}
question("How many columns in the data frame contain *logical* (TRUE or FALSE) values?",
         answer("6"),
         answer("5"),
         answer("All columns are logical"),
         answer("No columns are logical", correct = TRUE)
         )
```


<br>

#### 6. Technically, columns in data frames are actually *named arrays*. We can access them to work with directly using a dollar sign `$`. The code below extracts the `name` variable from the data frame. Run and understand this code, and then modify it to extract the `awake` column.

```{r msleep_q6, exercise = TRUE}
msleep$name
``` 

<div id= "msleep_q6-hint">
**Hint:**  Make sure to *not* use any quotes, i.e. do **not** run `msleep$"name"`
</div>

<br>

#### 7. Finally, let's get to everyone's favorite part - statistics! Below shows you how to calculate the *average* (mean!) hours that these mammals are awake for. Try out these other functions to calculate other summary statistics: `median()`, `range()`, `min()`, `max()`, `sd()` (standard deviation) on the `awake` column. 
```{r msleep_q7, exercise = TRUE}
mean(msleep$awake)
``` 

<br>

#### 8. Using your new skills, calculate the _median_ hours of REM sleep (columns `sleep_rem`) these mammals get. The output might not look as expected!! To start to understand why, view the full contents of this column (i.e. `msleep$sleep_rem`). You will see some values "NA", meaning there is *missing data*. Missing data is very common - real life science (especially biology!) is tricky, and sometimes we can't gather as much data as we want. 

```{r msleep_q8, exercise = TRUE}

``` 

<br>

#### 9. So, how do we calculate the median from the question above? We need to include the *second argument* `na.rm=TRUE` which tells median to ignore ("rm" for remove) NA's when performing calculations. See this code in action below, and then modify it to calculate the _mean_ of this column instead. 

```{r msleep_q9, exercise = TRUE}
 median(msleep$sleep_rem, na.rm = TRUE)
``` 

<br>

#### 10. The column `vore` is a *character* column - it contains strings rather than numbers. Let's examine this column using the function `unique()`, which returns all unique values in a given array - What are the unique vores in this dataset?

```{r msleep_q10, exercise = TRUE}

``` 

<div id= "msleep_q10-hint">
**Hint:** Remember, you can always look up how to use a function! Try running `?unique` if you're stuck.
</div>

<br>

#### 11. We can use the function `summary()` to summarize arrays, and even entire data frames. Run this function twice: first on the `awake` column and then on `msleep` itself (without indexing any columns), and behold the information!

```{r msleep_q11, exercise = TRUE}

``` 

<br>

#### An important learning moment!
You'll notice in the full `summary()` output for `msleep`, there isn't much useful information for the character columns. You may have expected all the *categories* in these columns to be revealed. In fact, R does not treat character variables this way. To force R to consider a vector specifically as containing *categories*, you need to convert it into a new type of data called **factor**, which is a special type of variable in R for, you guessed it, working with categories. The code below shows how the column `vore` would be summarized when converted to a factor with the function `as.factor()`. Run and understand this code.

```{r msleep_factors, exercise=TRUE}
## Treat it regularly
summary(msleep$vore)

## Treat it as a factor. The column is NOT redefined permanently!!
## Spaces are used for clarity, but they are not needed
summary(  as.factor(msleep$vore) )

## Redefine the column permanently like this:
msleep$vore <- as.factor(msleep$vore)

## We can confirm it is now a factor like this. It outputs "factor" not "character"!
class(msleep$vore)
class(msleep$vore) == "factor" ## Another way to confirm - this will be TRUE if we converted to factor properly

## Now, run summary on just vore column. It's treated forever as factor since we REDFINED it
summary(msleep$vore)

## The function levels() can be used on a factor to see its categories
levels(msleep$vore)
```

<div id= "msleep_factors-hint">
**Hint:** The `levels()` function will *only* work on a factor. If you run `levels()` on a variable that is NOT type factor, the function will return `NULL`. 
</div>

<br>

#### 12. *Redefine* the column `conservation` to be a factor, and reveal its categories using `levels()`. Use code from the demo above to help you! 

```{r msleep_q12, exercise=TRUE}

```

## Answer Keys

### Set 1
<br>

<br>

#### 1. What is 5 times 6?
```{r, calc_q1-answer}
5*6
```

<br>

#### 2. What is 600 divided by 2.53?
```{r, calc_q2-answer}
600/2.53
```

<br>

#### 3. What is 138 squared?
```{r, calc_q3-answer, collapse=TRUE}
138**2 

#Equally valid approach
138^2 
```

<br>

#### 4. What is the square root of 14?
```{r, calc_q4-answer}
14^0.5
```


### Set 2
<br>

<br>

#### 1. Use a *logical operator* to ask if 56 is greater than 7.

```{r, logical_q1-answer}
56 > 7
```

<br>

#### 2. Use a logical operator to ask if 110 is equal to 12.

```{r, logical_q2-answer}
110 == 12
```

<br>

#### 3. Use a logical operator to ask if 32 is less than or equal to 88.

```{r, logical_3-answer}
32 <= 88
```

<br>

#### 4. Use a logical operator to ask if 100 is equal to -10 squared.

```{r, logical_q4-answer}
100 == (-10)^2
```

<br>

#### 5. Use a logical operator to ask if 100 is *not* equal to -55. 

```{r, logical_q5-answer}
100 != -55
```

<br>

#### 6. Use a logical operator to ask if 100 is *not* less than 88. 

```{r, logical_q6-answer}
!(100 < 88)
```

<br>

#### 7. In all questions above, you wrote *single logical statements* - you used a logical operator to evaluate if a single statement was `TRUE` or `FALSE`. This time, write a *compound logical statement* to ask if: 100 is equal to 100 *and* 66 is equal to 66.

```{r, logical_q7-answer}
100 == 100 & 66 == 66
```

<br>

#### 8. Write a compound logical statement to ask if 100 is greater than 87 *and* 66 is less than 12.

```{r, logical_q8-answer}
100 > 87  & 66 < 12
```

<br>

#### 9. Similar to the previous question, write a compound logical statement to ask if 100 is greater than 87 *OR* 66 is less than 12.

```{r, logical_q9-answer}
100 > 87 | 66 < 12
```

### Set 3
<br>

<br>

#### 2. Using the code presented in question 1 to help guide you, define a variable named `banana_price` whose value is 0.55 (aka 55 cents). 

```{r, variables_q2-answer}
 banana_price <- 0.55
```

<br>

#### 3. Again, define the `banana_price` variable (since these questions don't communicate with each other!!). Use R as a calculator to determine the cost of 17 bananas.

```{r, variables_q3-answer}
banana_price <- 0.55
banana_price * 17
```

<br>

#### 4. Again, define the `banana_price` variable. Define a second variable called `num_bananas` with the value for how many bananas you're buying. Using both your variables, use R as a calculator to calculate the total cost.

```{r, variables_q4-answer}
banana_price <- 0.55
num_bananas <- 17
banana_price * num_bananas
```

<br>

#### 5. Again, define your two variables, `banana_price` and `num_bananas`. Again calculate the total cost using these variables, *but this time*, save the output to a new variable called `total_cost`. Output the final value of `total_cost` by simply typing the variable name on its own line.

```{r, variables_q5-answer}
banana_price <- 0.55
num_bananas <- 17
total_cost <- banana_price * num_bananas
total_cost
```

<br>

#### 6. Let's work with *logical operations* and variables. Imagine you have \$5 to spend at the grocery store. Can you afford to buy 30 bananas (you are apparently making *a lot* of quarantine banana bread)? To answer this question, first *define* four variables (`banana_price`, `num_bananas`[ but what's the value this time?!], `total_cost`, and `total_money` which represents your budget). Then, use logical operations to compare `total_cost` and `total_money`. You should arrive at an answer of TRUE *if it is affordable*, and FALSE if you can't buy this many bananas.

```{r, variables_q6-answer}
banana_price <- 0.55
num_bananas <- 30
total_money <- 5
total_cost <- banana_price * num_bananas
total_cost <= total_money
```

<br>

#### 7. The grocery store visit is getting more interesting. Sure, banana bread is all well and good, but it's certainly not as good as chocolate-chip banana bread. You've recently come into some money, so you now have \$20 to spend!! You want to buy 4 bags of chocolate chips that cost \$1.50 each as well as 16 bananas which, due to some evil price-gouging, now cost 60 cents (\$0.60) each. Can you afford this purchase? 

```{r, variables_q7-answer}
banana_price <- 0.6
num_bananas <- 16
choc_price <- 1.5
num_choc  <- 4
total_money <- 20
total_cost <- (banana_price * num_bananas) + (choc_price * num_choc)
total_cost <= total_money
```


### Set 4

<br>

#### 1. Use the function `sqrt()` to calculate the square root of 206.

```{r, func_q1-answer}
sqrt(206)
```

<br>

#### 2. Use the function `log()` to calculate the *natural log* (this is actually the default behavior of this function) of 206.

```{r, func_q2-answer}
log(206)
```

<br>

#### 3. Use the function `log()` to calculate the log in base 10 of 206.

```{r, func_q3-answer}
log(206, 10)
```

<br>

#### 4. Use the function `class()` to determine the *type* of value 7.

```{r, func_q4-answer}
class(7)
```

<br>

#### 5. Now, use `class()` to determine the *type* of value "7".

```{r, func_q5-answer}
class("7")
```

<br>

#### 6. Let's combine our skills! Define a variable called `x` whose value is 7 (no quotes!). Then, determine its type with `class()`, but **save this output** to a new variable called `type_x`. Finally, use logical operators with this variable to ask: Is the type of the variable equal to (hint!) numeric?

```{r, func_q6-answer}
x <- 7
type_x <- class(x)
type_x == "numeric"
```


### Set 5

<br>

#### 1. Define an array variable called `array_numbers` that contains the values 1, 2, 3, 4, and 5. Then, type the variable name itself to make sure its contents look right.

```{r, array_q1-answer}
array_numbers <- c(1,2,3,4,5)
array_numbers
```

<br>

#### 2. Define an array variable called `array_numbers2` that again contains the values 1, 2, 3, 4, and 5, **but!** this time use a colon `:` instead of `c()`. Again, type the variable name itself to make sure its contents look right.

```{r, array_q2-answer}
array_numbers2 <- 1:5
array_numbers2
```

<br>

#### 3. Define an array variable called `even_numbers` that contains the values 2, 4, 6, 8, and 10 using the `c()` function.

```{r, array_q3-answer}
even_numbers <- c(2,4,6,8,10)
even_numbers
```

<br>

#### 4. Define an array variable called `even_numbers` that contains the values 2, 4, 6, 8, and 10 using a new function `seq()`. 

```{r, array_q4-answer}
even_numbers <- seq(2, 10, 2)
even_numbers
```

<br>

#### 5. The function `length()` tells you how many items are in a given array. Define an array using `seq()` as the multiples of 3 from 0-999 (i.e. 0, 3, 6, 9, ..., 993, 996, 999) and then use `length()` to determine how many numbers it contains.

```{r, array_q5-answer}
mul3 <- seq(0, 999, 3)
length(mul3)
```

<br>

#### 6. Modify the code to instead divide every item in the array by 10.
  
```{r, array_q6-answer}
example_array <- c(5, 78, -3, 11, 6, 0)
example_array / 10
```

<br>

#### 7. Modify the code to instead ask if every item is *less than or equal to 0**.

```{r, array_q7-answer}
example_array <- c(5, 78, -3, 1, 6, 0)
example_array <= 0
```

<br>

#### 8. Use the function `sum()` (adds up values in an array) to calculate the sum of the output from `example_array > 0`. 


```{r, array_q8-answer}
example_array <- c(5, 78, -3, 1, 6, 0)
sum(example_array > 0)
```


### Set 6
<br>

#### 6. Modify this code to extract the `awake` column.


``` {r, msleep_q6-answer}
msleep$awake
```

<br>

#### 7. Try out these other functions to calculate other summary statistics: `median()`, `range()`, `min()`, `max()`, `sd()` (standard deviation) on the `awake` column.

``` {r, msleep_q7-answer}
median(msleep$awake)
range(msleep$awake)
min(msleep$awake)
max(msleep$awake)
sd(msleep$awake)
```

<br>

#### 8. Using your new skills, calculate the _median_ hours of REM sleep (columns `sleep_rem`) these mammals get. The output might not look as expected!! To start to understand why, view the full contents of this column (i.e. `msleep$sleep_rem`). You will see some values "NA", meaning there is *missing data*. Missing data is very common - real life science (especially biology!) is tricky, and sometimes we can't gather as much data as we want. 

``` {r, msleep_q8-answer}
median(msleep$sleep_rem)
```

<br>

#### 9. So, how do we calculate the median from the question above? We need to include the *second argument* `na.rm=TRUE` which tells median to ignore ("rm" for remove) NA's when performing calculations. See this code in action below, and then modify it to calculate the _mean_ of this column instead. 

``` {r, msleep_q9-answer}
mean(msleep$sleep_rem, na.rm = TRUE)
```

<br>

#### 10. The column `vore` is a *character* column - it contains strings rather than numbers. Let's examine this column using the function `unique()`, which returns all unique values in a given array - What are the unique vores in this dataset?

``` {r, msleep_q10-answer}
unique(msleep$vore)
```

<br>

#### 11. We can use the function `summary()` to summarize arrays, and even entire data frames. Run this function twice: first on the `awake` column and then on `msleep` itself (without indexing any columns), and behold the information!

``` {r, msleep_q11-answer}
summary(msleep$awake)
summary(msleep)
```

<br>

#### 12. *Redefine* the column `conservation` to be a factor, and reveal its categories using `levels()`. Use code from the demo above to help you! 

``` {r, msleep_q12-answer}
msleep$conservation <- as.factor(msleep$conservation)
levels(msleep$conservation)
```

