---
title: "Introduction to R"
author: "Stephanie J. Spielman"
output: 
  learnr::tutorial:
    theme: united
    css: css/custom-css.css
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(learnr)
library(ggplot2) ## for msleep dataset
tutorial_options(exercise.lines = 10)
```

# Data Science for Biologists, Fall 2020
<br><br><br>

## Set 1: Using R as a calculator

<br>

1. What is 5 times 6?

> Hint:  Use the asterisk `*` to multiply numbers

```{r calc_q1, exercise =TRUE}

```
  
```{r calc_q1-solution}
5*6
```

<br>

2. What is 600 divided by 2.53?

> Hint:  Use the forward slash `/` to divide numbers
 
```{r calc_q2, exercise = TRUE}

```

```{r calc_q2-solution}
600/2.53
```
<br>

3. What is 138 squared?

> Hint:  Use the carot symbol `^` *or* two asterisks (`**`) to raise numbers to an exponent.

```{r calc_q3, exercise = TRUE}

```

```{r calc_q3-solution, collapse=TRUE}
138**2 

#Equally valid approach
138^2 
```

<br>

4. What is the square root of 14?
  
> Hint:  Square roots are like raising to the power of 0.5
 
```{r calc_q4, exercise = TRUE}

``` 

```{r calc_q4-solution}
14^0.5
```

5. What is the answer to: 11 plus 10 all divided by 3?

> Hint:  Remember your PEMDAS!

  
```{r calc_q5, exercise = TRUE}

``` 

```{r calc_q5-solution}
(11+10)/3
```



## Set 2: Working with logical values

<br><br>


Logical operators:

| Symbol | Asks if... | 
|:--------|:------------|
|`==`    | Two values are equal |
|`!=`    | Two values are *not* equal |
| `>`    | LHS (left-hand side) is greater than RHS (right-hand side) |
| `>=`   | LHS is greater than or equal to RHS | 
| `<`    | LHS is less than RHS  |
| `<=`   | LHS is less than or equal to RHS | 
| `&`    | Used to combine logical statements and ask if **BOTH** are `TRUE`
| `|`    | Used to combine logical statements and ask if **AT LEAST ONE** is `TRUE`. Type this symbol with `Shift`+`\`.

<br>

1. Use a *logical operator* to ask if 56 is greater than 7.
  
```{r logical_q1, exercise = TRUE}

``` 

```{r logical_q1-solution}
56 > 7
```
<br>

2. Use a *logical operator* to ask if 110 is equal to 12.

> Hint:  When comparing *equality* of numbers, you MUST use two equals!! `==`


```{r logical_q2, exercise = TRUE}

```

```{r logical_q2-solution}
110 == 12
```

<br>

3. Use a *logical operator* to ask if 32 is less than or equal to 88.
  
```{r logical_q3, exercise = TRUE}

```

```{r logical_q3-solution}
32 <= 88
```

<br>

4. Use a *logical operator* to ask if 100 is equal to -10 squared.

> Hint: Remember your PEMDAS and don't forget to use parentheses!! Otherwise, your answer will not be right.
  
```{r logical_q4, exercise = TRUE}

```
```{r logical_q4-solution}
100 == (-10)^2
```

<br>

5. Use a *logical operator* to ask if 100 is *not equal* to -55. 
  
> Hint: Check if values are *not equal* with the logical operator `!=`. Generally speaking, exclamation marks negate logical statements.

```{r logical_q5, exercise = TRUE}

```

```{r logical_q5-solution}
100 != -55
```

<br>

6. Use a *logical operator* to ask if 100 is *not less than* 88. *Negating* the `<` operator, and all other operators besides `==`, you need to *enclose* the whole statement inside parentheses prefixed by an exclamation mark, like: `!(...statement...)` (where `...statement...` is whatever you are testing). 
 
```{r logical_q6, exercise = TRUE}

```
```{r logical_q6-solution}
!(100 < 88)
```


<br>

7. In all questions above, you wrote *single logical statements* - you used a logical operator to evaluate if a single thing was TRUE or FALSE. This time, write a *compound logical statement* to ask if: 100 is equal to 100 *and* 66 is equal to 66. Statements with "and" will only evaluate as `TRUE` if BOTH conditions are `TRUE`. Simply use an ampersand `&` to combine logical statements. To get you started, there is an example below you can adapt!

```{r logical_q7, exercise = TRUE}
# This example code asks: is 1 equal to 1 and 2 equal to 2?
1 == 1 & 2 == 2
# This is equivalent, but the parentheses are your choice!
(1 == 1) & (2 == 2)

```
```{r logical_q7-solution}
100 == 100 & 66 == 66
```

<br>

8. Write a *compound logical statement* to ask if 100 is greater than 87 *and* 66 is less than 12.
  
```{r logical_q8, exercise = TRUE}

```
```{r logical_q8-solution}
100 > 87  & 66 < 12
```
<br>

9. Similar to the previous question, write a *compound logical statement* to ask if 100 is greater than to 87 *OR* 66 is less than 12. Statements with "or" will evaluate as `TRUE` if *at least one* condition is `TRUE`. Rather than `&`, we need to use this symbol for "or": `|`. This is not the letter L or I - this symbol is located on the backslash `\` key, above the return/enter key. Press Shift + `\` to get it.
  
```{r logical_q9, exercise = TRUE}

```
```{r logical_q9-solution}
100 > 87 | 66 < 12
```



## Set 3: Working with variables in R

<br><br>


<br>

1. The code example below defines a variable. Answer the multiple choice questions about this code:
```{r}
y <- 8.5
```

```{r quiz1, echo=FALSE}
quiz(caption = "Multiple Choice",
  question("What is the *name* of this variable?",
     answer("`<-`", message = "This is the assignment operator."),
     answer("y", correct = TRUE),
     answer("8.5"),
     answer("We can't tell."),
     allow_retry = TRUE
     ),
  question("What is the *value* of this variable?",
     answer("`<-`",  message = "This is the assignment operator."),
     answer("y"),
     answer("8.5", correct = TRUE),
     answer("We can't tell."),
     allow_retry = TRUE
     ),
  question("What is the *type* of this variable?",
     answer("Character"),
     answer("Numeric", correct = TRUE),
     answer("Logical"),
     answer("We can't tell."),
     allow_retry = TRUE
  )
)
```

<br>

> **Before you begin, an important message!** After you define a variable, **always** print out the variable to make sure it was defined correctly! In these exercises, you can reveal the value of a variable simply by typing and running the variable's name. Try it out below! First, run the code "as is" - this will define y as equal to 8.5 internally, but it won't do anything else! Then, type y directly to print its output:

```{r demo_print, exercise=TRUE}
## Gives no output on its own, but does something *internally*. 
y <- 8.5 

## Next, type y below and re-run this code chunk. The value 8.5 will be printed out
```

<br>

2. Using the code presented in question 1 to help guide you, define a variable named `banana_price` whose value is 0.55 (aka 55 cents). 

> Hint:  Use the assignment operator `<-` to assign the value 0.55 to the variable name `banana_price`.

```{r variables_q2, exercise = TRUE}

```
```{r variables_q2-solution}
 banana_price <- 0.55
```


<br>

3. Again, define the `banana_price` variable (since these questions don't communicate with each other!!). Use R as a calculator to determine the cost of 17 bananas. Perform this operation using the `banana_price` variable itself.


```{r variables_q3, exercise = TRUE}

```
```{r variables_q3-solution}
banana_price <- 0.55
banana_price * 17
```
  
<br>

4. Again, define the `banana_price` variable. Define a second variable called `num_bananas` with the value for how many bananas you're buying. Using both your variables, use R as a calculator to calculate the total cost.

> We are learning to use variables whenever possible instead of the direct numbers. Using values in calculations rather than variables representing those values is called **hard-coding** and can lead to bugs!

```{r variables_q4, exercise = TRUE}

```
```{r variables_q4-solution}
banana_price <- 0.55
num_bananas <- 17
banana_price * num_bananas
```
<br>

5. Again, define your two variables, `banana_price` and `num_bananas`. Again calculate the total cost using these variables, *but this time*, save the output to a new variable called `total_cost`. 
  
```{r variables_q5, exercise = TRUE}

```

```{r variables_q5-solution}
banana_price <- 0.55
num_bananas <- 17
total_cost <- banana_price * num_bananas
total_cost
```
<br>

### WARNING!
#### You'll really want your pen-and-paper now to help organize your thoughts!! Thinking through word problems while coding is HARD! First, plan your general steps using pen-and-paper, then figure out how to turn those steps into code.

<br>

6. Let's work with *logical operations* and variables. Imagine you have \$5 to spend at the grocery store. Can you afford to buy 30 bananas (you are apparently making *a lot* of quarantine banana bread)? 
To answer this question, first *define* four variables (`banana_price`, `num_bananas`[ but what's the value this time?!], `total_cost`, and `total_money` which represents your budget). Then, use logical operations to compare `total_cost` and `total_money`. You should arrive at an answer of TRUE *if it is affordable*, and FALSE if you can't buy this many bananas.

> Hint:  A `TRUE` solution is one where `total_money` is *greater than or equal to* the `total_cost`. Alternatively, where `total_cost` is *less than or equal to* `total_money`.
   
```{r variables_q6, exercise = TRUE}

```
```{r variables_q6-solution}
banana_price <- 0.55
num_bananas <- 30
total_money <- 5
total_cost <- banana_price * num_bananas
total_cost <= total_money
```

<br>

7. The grocery store visit is getting more interesting. Sure, banana bread is all well and good, but it's certainly not as good as chocolate-chip banana bread. You've recently come into some money, so you now have \$20 to spend!! You want to buy 4 bags of chocolate chips that cost \$1.50 each as well as 16 bananas which, due to some evil price-gouging, now cost 60 cents (\$0.60) each. Can you afford this purchase? 

> Hint: Make sure to define all variables with *meaningful variable names* first, *then* perform calculations to determine the `total_cost`, and then compare `total_cost` to `total_money`. In the end, you will have created six variables. **BUT!!** Make sure to *never ever EVER* put spaces into your variable names. Instead, use underscores to represent spaces.
   
```{r variables_q7, exercise = TRUE}

```


```{r variables_q7-solution}
banana_price <- 0.6
num_bananas <- 16
choc_price <- 1.5
num_choc  <- 4
total_money <- 20
total_cost <- (banana_price * num_bananas) + (choc_price * num_choc)
total_cost <= total_money
```

## Set 4: Working with functions

<br>

1. Use the function `sqrt()` to calculate the square root of 206.
  
> Hint: Provide 206 as an *argument* to the function `sqrt()`

```{r func_q1, exercise = TRUE}

```
```{r func_q1-solution}
sqrt(206)
```

<br>

2. Use the function `log()` to calculate the *natural log* (this is actually the default behavior of this function) of 206.
  
```{r func_q2, exercise = TRUE}

```
```{r func_q2-solution}
log(206)
```

<br>

3. Looking up documentation for how to use functions in R (and other languages!) is *very important* - don't try to be a psychic!! To see the documentation for a function, type `?nameoffunction`. To see this in action, run the code below. Have a particular look at the sections "Usage" and "Arguments." Then, use the function `log()` to calculate the log in base 10 of 206.

> Hint: The function `log()` can optionally take a *second* argument specifying the base. For example, `log(500, 6)` takes the log of 500 in base 6. Make sure you see where this information is described in the `?log` output!

> Note: Sometimes these official help pages are written really poorly. Over time, you'll become more comfortable with them! For now, focus on sections "Usage" and "Arguments."
 
```{r func_q3, exercise = TRUE}
?log()

``` 

```{r func_q3-solution}
log(206, 10)
```
<br>

4. The function `class()` will tell you what *type* of data a given variable or value is. Use this function to determine the *type* of value 7.

> Hint: Provide 7 as the argument to `class()`

  
```{r func_q4, exercise = TRUE}

``` 
```{r func_q4-solution}
class(7)
```

<br>

5. Now, use `class()` to determine the *type* of value `"7"`.
  
> Hint: Don't forget the quotes! Think: what did you learn from the last two questions about how quotes work?

```{r func_q5, exercise = TRUE}

```   
```{r func_q5-solution}
class("7")
```

<br>

6. Let's combine our skills! Define a variable called `x` whose value is 7 (no quotes!). Then, determine its type with `class()`, but **save this output** to a new variable called `type_x`. Finally, use logical operators with this variable to ask: Is the type of the variable equal to numeric?

> Hint: Don't forget to use `==` when comparing equality of values. 

> Hint: If you receive an error, it may be because you forgot to put quotes around "numeric". If you don't use quotes, R thinks you're referring to a variable called `numeric`, but no such variable exists! R doesn't want to use variables that don't exist, because....they don't exist. 
  
```{r func_q6, exercise = TRUE}

```
```{r func_q6-solution}
x <- 7
type_x <- class(x)
type_x == "numeric"
```



## Set 5: Working with arrays


> Note: The terms "array" and "vector" both refer to the same thing and are used interchangeably, but a "list" is something else entirely! These exercises don't consider lists at all.


<br>

1. Define an array variable called `array_numbers` that contains the values 1, 2, 3, 4, and 5. Then, type and run the variable name itself to make sure its contents look right.

> Hint: Use the function `c()` ("combine") to define your array. The code `c(8,9,10)` returns an array with the values 8, 9, and 10. Can you adapt this code to the question?

 
```{r array_q1, exercise = TRUE}

```         
```{r array_q1-solution}
array_numbers <- c(1,2,3,4,5)
array_numbers
```
<br>

2. Define an array variable called `array_numbers2` that again contains the values 1, 2, 3, 4, and 5, **but!** this time use a colon `:` instead of `c()`. Again, type the variable name itself to make sure its contents look right.
     
> Hint:  `5:7` returns an array with values 5, 6, 7. Can you adapt this code to the question?

        
```{r array_q2, exercise = TRUE}

``` 
```{r array_q2-solution}
array_numbers2 <- 1:5
array_numbers2
```

<br>

3. Define an array variable called `even_numbers` that contains the values 2, 4, 6, 8, and 10 using the `c()` function.
  
```{r array_q3, exercise = TRUE}

``` 
```{r array_q3-solution}
even_numbers <- c(2,4,6,8,10)
even_numbers
```


<br>

4. The function `length()` tells you how many items are in a given array. Define an array with values of your choosing (be brave!!) and use the function `length()` to determine how many values it contains.

> Hint: First make sure (print print print!) that you are defining your array correctly. *Only once* you can see that the array is correct should you calculate its length. **Write your code one line at a time for success!!**
  
```{r array_q4, exercise = TRUE}

```   
```{r array_q4-solution}
# A random array to demonstrate:
my_favorite_things <- c("raindrops on roses", "whiskers on kittens", "bright copper kettles", "warm_woolen_mittens")
length(my_favorite_things)
```


<br>

5. One place where R really shines is *vectorization* - it can perform an operation on *all items in a vector* at the same time! The code below multiplies every value in the array `example_array` by 5. Run this code and examine its output to be sure you understand it. Then, modify the code to instead divide every item in the array by 10.

> Hint: When modifying the code be careful not to delete the definition of `example_array`.

  
```{r array_q5, exercise = TRUE}
example_array <- c(5, 78, -3, 11, 6, 0)
example_array * 5
```     
```{r array_q5-solution}
example_array <- c(5, 78, -3, 11, 6, 0)
example_array / 10
```

<br>

6. We can also perform logical operations in a vectorized context. The code below asks if all values in `example_array` are greater than 0. Run this code and examine its output to be sure you understand it. Then, modify the code to instead ask if every item is *less than or equal to 0** (but be careful not to delete the `example_array` definition when modifying!).

> Note: This type of output would be referred to as a *logical array*.

  
```{r array_q6, exercise = TRUE}
example_array <- c(5, 78, -3, 1, 6, 0)
example_array > 0
``` 
```{r array_q6-solution}
example_array <- c(5, 78, -3, 1, 6, 0)
example_array <= 0
```
<br>

7. In fact, logical values `TRUE` and `FALSE` evaluate to numbers 1 and 0, respectively. Use the function `sum()` (adds up values in an array) to calculate the sum of the output from `example_array > 0`. (Don't forget to define `example_array` again here!).

>  Hint: You can provide the full expression `example_array > 0` to the function `sum()`, or save that logical expression's output to a variable, and use the variable as the argument to `sum()`.


              
```{r array_q7, exercise = TRUE}

```                
```{r array_q7-solution}
example_array <- c(5, 78, -3, 1, 6, 0)
sum(example_array > 0)
```

8. One of the most POWERFUL logical opertators is `%in%` - it's literally the word `in` enclosed by `%`. You can use it to ask if a given value is present *in* an array. The code below shows an example of using this operator. Run the code and examine its output to see how it works. Then, modify the code to instead ask of the value `"marmoset"` is in the array. *For an added challenge*, defined a variable containing the value `"marmoset"` and use that variable in your code!

```{r array_q8, exercise = TRUE}
animals <- c("goat", "turtle", "axolotl", "hyena", "coral")
"goat" %in% animals
```  

```{r array_q8-solution}
animals <- c("goat", "turtle", "axolotl", "hyena", "coral")
new_animal <- "marmomset"
new_animal %in% animals
```  

9. Use the `%in%` operator to ask if the value -25 is in an array you define to contain all numbers 50 - 500. Use variables in your comparison - first define the value and the array as variables and then compare them with `%in%`.

> Hint: Remember the colon `:` to define ordered arrays!

```{r array_q9, exercise = TRUE}

```  
```{r array_q9-solution}
many_numbers <- 50:500
test_number <- -25
test_number %in% many_numbers
```  

## Set 6: Data frames


This set of exercises will use a *data frame* (aka *tibble*- we will learn the distinction between these terms!) called `msleep`, which has been pre-loaded for you. This dataset contains information about some species of mammals as follows:

+ `name`: The common name
+ `genus`: The taxonomic genus
+ `vore`: Dietary strategy (carnivore, omnivore or herbivore)
+ `order`: The taxonomic order
+ `conservation`: The species' conservation status
  + `cd` --> conservation-dependent
  + `en` --> endangered
  + `lc` --> least concern
  + `nt` --> near-threatened
  + `vu` --> vulnerable    
+ `sleep_total`: Total amount of sleep, in hours, the species tends to experience
+ `sleep_rem`: Total amount of REM sleep, in hours, the species tends to experience
+ `sleep_cycle`: Length of sleep cycle, in hours, the species tends to experience
+ `awake`: Amount of time spent awake, in hours, the species tends to experience
+ `brainwt`: Average brain weight, in kilograms, for this species
+ `bodywt`: Average body weight, in kilograms, for this species


First, in the code chunk below, examine the data frame using functions we have seen `head()`, `nrow()`, and `ncol()` to get an overall sense of the data frame itself. Then,  answer the following multiple choice questions about this data frame.
```{r examine-msleep, exercise=TRUE}
msleep ## Reveals the whole data frame

## Examine below with head(), nrow(), ncol()


## Bonus! Try using the function str() to examine msleep


```

<br>

1. 

```{r msleep_q1, echo=FALSE}
question("How many rows are in this data frame?",
         answer("6"),
         answer("83", correct = TRUE),
         answer("11"),
         answer("We can't tell"),
         allow_retry = TRUE
         )
```

<br>

2. 

```{r msleep_q2, echo=FALSE}
question("How many columns are in this data frame?",
         answer("6"),
         answer("83"),
         answer("11", correct = TRUE),
         answer("We can't tell"),
         allow_retry = TRUE
         )
```

<br>

3. 

```{r msleep_q3, echo=FALSE}
question("How many columns in the data frame contain *numeric* values?",
         answer("6", correct = TRUE),
         answer("5"),
         answer("All columns are numeric"),
         answer("No columns are numeric"),
         allow_retry = TRUE
         )
```

<br>

4. 

```{r msleep_q4, echo=FALSE}
question("How many columns in the data frame contain *character* (string) values?",
         answer("6"),
         answer("5", correct = TRUE),
         answer("All columns are numeric"),
         answer("No columns are numeric")
         )
```

<br>

5. 

```{r msleep_q5, echo=FALSE}
question("How many columns in the data frame contain *logical* (TRUE or FALSE) values?",
         answer("6"),
         answer("5"),
         answer("All columns are logical"),
         answer("No columns are logical", correct = TRUE)
         )
```


<br>

6. Technically, columns in data frames are actually *named arrays*. We can access them to work with directly using a dollar sign `$`. The code below extracts the `name` variable from the data frame. Run and understand this code, and then modify it to extract the `awake` column.

> Hint:  Make sure to *not* use any quotes, i.e. do **not** run `msleep$"name"`


```{r msleep_q6, exercise = TRUE}
msleep$name
``` 

``` {r msleep_q6-solution}
msleep$awake
```


<br>

7. Finally, let's get to everyone's favorite part - statistics! Below shows you how to calculate the *average* (mean!) hours that these mammals are awake for. Try out these other functions to calculate other summary statistics: `median()`, `range()`, `min()`, `max()`, `sd()` (standard deviation) on the `awake` column. 

```{r msleep_q7, exercise = TRUE}
mean(msleep$awake)
``` 


``` {r msleep_q7-solution}
median(msleep$awake)
range(msleep$awake)
min(msleep$awake)
max(msleep$awake)
sd(msleep$awake)
```

<br>

8. Using your new skills, calculate the _median_ hours of REM sleep (columns `sleep_rem`) these mammals get. The output might not look as expected!! To start to understand why, view the full contents of this column (i.e. `msleep$sleep_rem`). You will see some values "NA", meaning there is *missing data*. Missing data is very common - real life science (especially biology!) is tricky, and sometimes we can't gather as much data as we want. 

```{r msleep_q8, exercise = TRUE}

``` 

``` {r msleep_q8-solution}
median(msleep$sleep_rem)
```
<br>

9. So, how do we calculate the median from the question above? We need to include the *second argument* `na.rm=TRUE` which tells median to ignore ("rm" for remove) NA's when performing calculations. See this code in action below, and then modify it to calculate the _mean_ of this column instead. 

> Note: `na.rm` is NOT A FUNCTION!!! It is an ARGUMENT we can use with SOME functions, such as `mean()` and `median()`.

```{r msleep_q9, exercise = TRUE}
 median(msleep$sleep_rem, na.rm = TRUE)
``` 

``` {r msleep_q9-solution}
mean(msleep$sleep_rem, na.rm = TRUE)
```
<br>

10. The column `vore` is a *character* column - it contains strings rather than numbers. Let's examine this column using the function `unique()`, which returns all unique values in a given array - What are the unique vores in this dataset?

> Hint: Remember, you can always look up how to use a function! Try running `?unique` if you're stuck.


```{r msleep_q10, exercise = TRUE}

``` 
``` {r msleep_q10-solution}
unique(msleep$vore)
```
<br>

11. We can use the function `summary()` to summarize arrays, and even entire data frames. Run this function twice: first on the `awake` column and then on `msleep` itself (without indexing any columns), and behold the information!

```{r msleep_q11, exercise = TRUE}

``` 
``` {r msleep_q11-solution}
summary(msleep$awake)
summary(msleep)
```

<br>

### An important learning moment!

You'll notice in the full `summary()` output for `msleep`, there isn't much useful information for the character columns. You may have expected all the *categories* in these columns to be revealed. In fact, R does not treat character variables this way. To force R to consider a vector specifically as containing *categories*, you need to convert it into a new type of data called **factor**, which is a special type of variable in R for, you guessed it, working with categories. The code below shows how the column `vore` would be summarized when converted to a factor with the function `as.factor()`. 

Learn how factors in action by running the code below LINE BY LINE. Examine and understand the output each time! Remember, your code will build on itself in R's "memory."

```{r msleep_factors, exercise=TRUE, exercise.lines = 20}
## Treat it regularly
summary(msleep$vore)

## Treat it as a factor. The column is NOT redefined permanently!!
## Spaces are used for clarity, but they are not needed
summary(  as.factor(msleep$vore) )

## Redefine the column permanently like this:
msleep$vore <- as.factor(msleep$vore)

## We can confirm it is now a factor like this. It outputs "factor" not "character"!
class(msleep$vore)
class(msleep$vore) == "factor" ## Another way to confirm - this will be TRUE if we converted to factor properly

## Now, run summary on just vore column. It's treated forever as factor since we REDFINED it
summary(msleep$vore)

## The function levels() can be used on a factor to see its categories.
# This function will only work on a factor. If you run `levels()` on a variable that is NOT type factor, the function will return `NULL`. 
levels(msleep$vore)
```


<br>

12. *Redefine* the column `conservation` to be a factor, and reveal its categories using `levels()`. Use code from the demo above to help you! 

```{r msleep_q12, exercise=TRUE}

```

``` {r msleep_q12-solution}
msleep$conservation <- as.factor(msleep$conservation)
levels(msleep$conservation)
```
